### 2017-08-17 E1/Code/R_AnalysisCode/Perfect_dependent_t-test.Rmd
initial commit

---
title: "Dependent t-test"
author: "Daniel Lakens, D.Lakens@tue.nl"
output: word_document
---
```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
####################################
#######  How to use ################
####################################

# 1) Install the packages and software specified below (consider also updating all installed packages by chosing Update on the Packages tab). 
# 2) Define the variable names, change the default settings, if desired
# 3) Hit Knit Word (or Knit HTML)! (Output can take while due to calculating the Bayesian Highest Density Iinterval)

####################################
######  FIRST TIME ONLY ############
####################################

#Remove # in front of the line below and run the code. Replace the # after installing the packages, otherwise the R markdown script will give errors.

#install.packages(c("MASS", "akima", "robustbase", "cobs", "robust", "mgcv", "scatterplot3d", "quantreg", "rrcov", "lars", "pwr", "trimcluster", "mc2d", "psych", "Rfit","MBESS", "BayesFactor", "PoweR", "ggplot2", "reshape2", "plyr", "devtools","rmarkdown","gmodels", "car", "gridExtra", "bootES", "BEST"))

#The version of HLMdiag on CRAN gives an error - run the four lines of code below to install the latest version of HLMdiag. Install all required software.
#install.packages("devtools")
#install.packages("stringi")
#library(devtools)
#install_github("aloy/HLMdiag")

#Installation of the robust statistics package: Remove # in front of each of 4 lines below and run the code. Replace the # after installing the packages, otherwise the R markdown script will give errors.

#install.packages("devtools")
#library("devtools")
#install_github("mrxiaohe/WRScpp")
#install_github("nicebread/WRS", subdir="pkg")

#Download and install JAGS to calculate Bayesian HDI: http://sourceforge.net/projects/mcmc-jags/files/JAGS/

###################################
## Define variables names below ###
###################################

#Copy paste your data into a spreadsheet. Create 3 columns. On the first row, type the name of the variables. One column should contain the subject identifer (e.g., PPNR). The other two columns should contain the data from the two within-subject conditions. Make sure there is no missing data (or emtpy cells underneath the data - check in .txt file).
#Save your data as a .txt file - select Text (tab delimited)(*.txt).

#Make sure the data file and the R markdown file are in the same folder.

# Backwards Letters - Study 2 - Experiment 1
# This script loads the parameters (efficacy, latency, and precision) 
# from the matlab data file--output from mixture modelling code. 
# Creates a seperate data frame for each parameter in both wide and long format.

library(R.matlab)

E1.data <- readMat('/Users/admin/Documents/MATLAB/BackwardsLetters/Study_2/Exp1_Random_0and180degs/Data/Exp1_InvertedLtrs_ParamsForAnalysis.mat')
Subject <- c('AA','AB', 'AC', 'AD', 'AH',  'AI', 'AJ', 'AK', 'AL', 'AO', 'AU', 'AV', 'AW', 'AX', 'AZ', 'BA', 'BC', 'BD', 'BE', 'BF', 'BG', 'BI', 'BJ', 'BK', 'BN', 'BO', 'BP', 'BQ', 'BR', 'BS')

# Create dataFrame for efficacy
LeftCanon <- E1.data$E1.allEstMinusExcluded[1,1,,1]*100 # [Condition, Stream, Participant, Parameter]
RightCanon <- E1.data$E1.allEstMinusExcluded[1,2,,1]*100 # Muliply by 100 to convert to percentage
LeftInvert <- E1.data$E1.allEstMinusExcluded[2,2,,1]*100 # Left and right stream data are flipped in the
RightInvert <- E1.data$E1.allEstMinusExcluded[2,1,,1]*100 # inverted condition so we unflip them here
efficacy.wide <- data.frame(Subject, LeftCanon, RightCanon, LeftInvert, RightInvert)
remove(LeftCanon, RightCanon, LeftInvert, RightInvert)

# Create dataFrame for latency
LeftCanon <- E1.data$E1.allEstMinusExcluded[1,1,,2]*(1000/10) # [Condition, Stream, Participant, Parameter]
RightCanon <- E1.data$E1.allEstMinusExcluded[1,2,,2]*(1000/10) # Mutiply by 1000msec divided by item rate (10)
LeftInvert <- E1.data$E1.allEstMinusExcluded[2,2,,2]*(1000/10) # Left and right stream data are flipped in the
RightInvert <- E1.data$E1.allEstMinusExcluded[2,1,,2]*(1000/10) # inverted condition so we unflip them here
latency.wide <- data.frame(Subject, LeftCanon, RightCanon, LeftInvert, RightInvert)
remove(LeftCanon, RightCanon, LeftInvert, RightInvert)

# Create dataFrame for precision
LeftCanon <- E1.data$E1.allEstMinusExcluded[1,1,,3]*(1000/10) # [Condition, Stream, Participant, Parameter]
RightCanon <- E1.data$E1.allEstMinusExcluded[1,2,,3]*(1000/10) # Mutiply by 1000msec divided by item rate  (10)
LeftInvert <- E1.data$E1.allEstMinusExcluded[2,2,,3]*(1000/10) # Left and right stream data are flipped in the
RightInvert <- E1.data$E1.allEstMinusExcluded[2,1,,3]*(1000/10) # inverted condition so we unflip them here
precision.wide <- data.frame(Subject, LeftCanon, RightCanon, LeftInvert, RightInvert)
remove(LeftCanon, RightCanon, LeftInvert, RightInvert)

# Convert efficacy to long format
library(reshape2)
efficacy.long <- melt(data=efficacy.wide, id.var="Subject",
                  measure.vars=c("LeftCanon", "RightCanon", "LeftInvert", "RightInvert"),
                  variable.name="Condition")
names(efficacy.long)[names(efficacy.long)=="value"] <- "Efficacy"

# Split Condition column into Stream and Orientation
efficacy.long$Stream <- NA
efficacy.long$Stream[grepl("^Left",  efficacy.long$Condition)] <- "Left"
efficacy.long$Stream[grepl("^Right", efficacy.long$Condition)] <- "Right"
efficacy.long$Stream <- factor(efficacy.long$Stream)

efficacy.long$Orientation <- NA
efficacy.long$Orientation[grepl("Canon$",  efficacy.long$Condition)] <- "Canonical"
efficacy.long$Orientation[grepl("Invert$", efficacy.long$Condition)] <- "Inverted"
efficacy.long$Orientation <- factor(efficacy.long$Orientation, levels=c("Canonical","Inverted"))

# Remove the Condition column now
efficacy.long$Condition <- NULL

# Convert latency to long format
library(reshape2)
latency.long <- melt(data=latency.wide, id.var="Subject",
                    measure.vars=c("LeftCanon", "RightCanon", "LeftInvert", "RightInvert"),
                    variable.name="Condition")
names(latency.long)[names(latency.long)=="value"] <- "Latency"

# Split Condition column into Stream and Orientation
latency.long$Stream <- NA
latency.long$Stream[grepl("^Left",  latency.long$Condition)] <- "Left"
latency.long$Stream[grepl("^Right", latency.long$Condition)] <- "Right"
latency.long$Stream <- factor(latency.long$Stream)

latency.long$Orientation <- NA
latency.long$Orientation[grepl("Canon$",  latency.long$Condition)] <- "Canonical"
latency.long$Orientation[grepl("Invert$", latency.long$Condition)] <- "Inverted"
latency.long$Orientation <- factor(latency.long$Orientation, levels=c("Canonical","Inverted"))

# Remove the Condition column now
latency.long$Condition <- NULL

# Convert precision to long format
library(reshape2)
precision.long <- melt(data=precision.wide, id.var="Subject",
                      measure.vars=c("LeftCanon", "RightCanon", "LeftInvert", "RightInvert"),
                      variable.name="Condition")
names(precision.long)[names(precision.long)=="value"] <- "Precision"

# Split Condition column into Stream and Orientation
precision.long$Stream <- NA
precision.long$Stream[grepl("^Left",  precision.long$Condition)] <- "Left"
precision.long$Stream[grepl("^Right", precision.long$Condition)] <- "Right"
precision.long$Stream <- factor(precision.long$Stream)

precision.long$Orientation <- NA
precision.long$Orientation[grepl("Canon$",  precision.long$Condition)] <- "Canonical"
precision.long$Orientation[grepl("Invert$", precision.long$Condition)] <- "Inverted"
precision.long$Orientation <- factor(precision.long$Orientation, levels=c("Canonical","Inverted"))

# Remove the Condition column now
precision.long$Condition <- NULL

alldata <- data.frame("Subject" =latency.wide$Subject, 
                  "LeftCanon" = latency.wide$LeftCanon, 
                  "RightCanon" = latency.wide$RightCanon)

#<- read.delim("demo_dependent.txt", header=TRUE, dec=".") #Define the name of the tab delimited txt file (or .csv file) that contains at least 3 columns: a subject identifier and 2 columns of the data in the two within-subject conditions. Specify the decimal symbol (e.g., your text file stores numbers either as 3.14 or 3,14)

#alldata<-na.omit(alldata)#rows with missing data are removed

#IMPORTANT: Do not use spaces, hyphens, or other symbols in names. If variable names in your file include spaces, change the names.

#Define names of the two columns in alldata you want to compare. Difference is computed as x-y (so reverse labels as desired) 
xlabel<-"LeftCanon"  #name dependent variable 1 - needs to match the datafile!
ylabel<-"RightCanon"  #name dependent variable 2 - needs to match the datafile!

#Define name of individual subject identifier variable (e.g., the participantnumber)
subject<-"Subject"  #needs to match the datafile!

#Define name of the factor that describes the difference between x and y (e.g., condition, time).
factorlabel<-"stream"

#If you want to perform a dependent t-test on a subgroup of your data, change this:
subgrouplabel<-"" #Specify header of the addition column specifying the subgroups (e.g., "Year") (needs to match the datafile!). To analyze all data, leave empty (i.e., "")
subgrouptokeep<-"" #Specify the identifier of the group you want to analyze (e.g., "2013")

#Define name of the measure that describes the values of x and y (e.g., reaction times, self-reported happiness)
measurelabel<-"latency"

#Below names for axis are used. These CAN include spaces.
xlabelstring<-"stream" #define variables to be used for axis (can be replaced by "Any Label")
ylabelstring<-"latency"  #define variables to be used for axis

#Set your alpha level and confidence interval
alpha<-0.05
ConfInt<-0.95

#For the Bayes Factor, specify which effect you expect. Standard to 0.5 (small effect). Change to 0.707 (medium effect) or 1 (large effect).
BFrscale<-0.5

#Alternative hypothesis: specify "two.sided" (for x<>y), "less" (for x<y) or "greater" (for x>y)
H1<-"two.sided"

#Calculating the Bayesian HDI can take quite some time (e.g., 30 minutes) in large sample sizes. If you are in a hurry, set the variable below to "YES", if you want the HDI, set it to "NO"
InAHurry<-"NO"

#############################################################
### Changed the information above? Then hit 'Knit Word' #####
#############################################################
#################### Know your way around R? ################ 
############ Feel free to change the script below ###########
#############################################################

options(scipen=20) #disable scientific notation for numbers smaller than x (i.e., 10) digits (e.g., 4.312e+22)

#Remove other conditions in your datafile (only keep groups specified above)
if (subgrouplabel!=""){
  alldata<-subset(alldata, alldata[[subgrouplabel]]==subgrouptokeep)
}

#create two variables (x and y) that contain values of two datasets to be compared
x<-alldata[[xlabel]]
y<-alldata[[ylabel]]
diff<-x-y #difference scores

#Add difference to alldata dataframe for plotting.
alldata[["diff"]] <- diff

# Convert alldata to long format
library(reshape2)
alldata.long <- melt(alldata, id.vars = subject, measure.vars = c(xlabel,ylabel), variable.name = factorlabel,value.name = measurelabel)

#######################################################################
#######################################################################
########### Calculate CI for within and between #######################
################ Scripts from Baguley, 2012 ###########################
#######################################################################
#######################################################################

cm.ci <- function(data.frame, conf.level = ConfInt, difference = TRUE) {
  #cousineau-morey within-subject CIs
  k = ncol(data.frame)
  if (difference == TRUE) 
    diff.factor = 2^0.5/2
  else diff.factor = 1
  n <- nrow(data.frame)
  df.stack <- stack(data.frame)
  index <- rep(1:n, k)
  p.means <- tapply(df.stack$values, index, mean)
  norm.df <- data.frame - p.means + (sum(data.frame)/(n * k))
  t.mat <- matrix(, k, 1)
  mean.mat <- matrix(, k, 1)
  for (i in 1:k) t.mat[i, ] <- t.test(norm.df[i])$statistic[1]
  for (i in 1:k) mean.mat[i, ] <- colMeans(norm.df[i])
  c.factor <- (k/(k - 1))^0.5
  moe.mat <- mean.mat/t.mat * qt(1 - (1 - conf.level)/2, n - 1) * c.factor * 
    diff.factor
  ci.mat <- matrix(, k, 2)
  dimnames(ci.mat) <- list(names(data.frame), c("lower", "upper"))
  for (i in 1:k) {
    ci.mat[i, 1] <- mean.mat[i] - moe.mat[i]
    ci.mat[i, 2] <- mean.mat[i] + moe.mat[i]
  }
  ci.mat
}

bs.ci <- function(data.frame, conf.level = ConfInt, difference = FALSE) {
  # between-subject CIs
  k = ncol(data.frame)
  n <- nrow(data.frame)
  df.stack <- stack(data.frame)
  group.means <- colMeans(data.frame, na.rm = TRUE)
  if (difference == TRUE) 
    ci.mat <- (confint(lm(values ~ 0 + ind, df.stack)) - group.means) * 
    2^0.5/2 + group.means
  else ci.mat <- confint(lm(values ~ 0 + ind, df.stack))
  dimnames(ci.mat) <- list(names(data.frame), c("lower", "upper"))
  ci.mat
}

#change matrix output from functions to dataframe, add CI from between, add labels and means 
#order of matrix is flipped around for bs.ci, which returns alphabetically ordered rows
ci.sum<-as.data.frame(cm.ci(alldata[2:3]))
ordernames<-rownames(ci.sum)#store order of dataframe rows
ifelse(rownames(ci.sum)[1]!=xlabel,ci.sum<-ci.sum[rev(rownames(ci.sum)),],NA)#flip order around 
ci.sum[[factorlabel]] <- c(xlabel,ylabel)
ci.sum[[measurelabel]] <- c(mean(x),mean(y))
ci.sum <- ci.sum[order(ci.sum[[factorlabel]]),]
ci.sum[["lower.between"]] <- as.data.frame(bs.ci(alldata[2:3]))$lower
ci.sum[["upper.between"]] <- as.data.frame(bs.ci(alldata[2:3]))$upper
ci.sum <- ci.sum[order(ordernames),]#flip order around

##################################################################
##################################################################
######## PLOT DATA AND CHECK FOR OUTLIERS AND NORMALITY ##########
##################################################################
##################################################################

#Test normality

require(PoweR)
normalityrejections<-(statcompute(21, diff, levels = c(0.05))$decision + statcompute(6, diff, levels = c(0.05))$decision + statcompute(2, diff, levels = c(0.05))$decision + statcompute(7, diff, levels = c(0.05))$decision)

#Testing equality of variances

require(car)
pvalueLevene<-leveneTest(alldata.long[[measurelabel]] ~ as.factor((alldata.long[[factorlabel]])))$"Pr(>F)"[1:1]
if (pvalueLevene < 0.05){equalvar<-"the assumption that variances are equal is rejected (consider reporting robust statistics)."}
if (pvalueLevene >= 0.05){equalvar<-"the assumption that variances are equal is not rejected."}
cat("Levene's test for equality of variances (p = ", round(pvalueLevene, digits=2),") indicates that ",equalvar,sep="")

#######################################################
#######################################################
###Calculate CI around Cohen's d for within designs ###
#######################################################
#######################################################

#Calculate based on data
r<-cor(x, y) #correlation between dependent measures
m_diff<-mean(diff) #mean of difference scores
s_diff<-sd(diff) #standard deviation of difference scores
sd1<-sd(x) #standard deviation of group 1
sd2<-sd(y) #standard deviation of group 2
N <- length(x) #number of pairs

ttestresult<-t.test(x, y, alternative = H1, paired = TRUE, var.equal = TRUE, conf.level = ConfInt)
tvalue<-ttestresult$statistic #store t-value from dependent t-test
pvalue<-ttestresult$p.value #store p-value from dependent t-test
CI_diff<-ttestresult$conf.int #store confidence interval of mean difference
s_av <- sqrt((sd1^2+sd2^2)/2) #calculate average standard deviation for effect size calculation

#Specify direction of difference
if (mean(x)>mean(y)){direction<-"greater than"}
if(mean(x)<mean(y)){direction<-"smaller than"}
if(pvalue < alpha){surprising<-"surprising"}
if(pvalue >= alpha){surprising<-" not surprising"}

#Effect sizes and 95% CI
#Cohen's d_av, using s_av as standardizer

require(MBESS) # to calculate non-central t limits

d_av <- m_diff/s_av
d_unb <- (1-(3/(4*(N-1)-1)))*d_av #note this is approximation of correction for Hedges'g - ESCI uses accurate correction, so should we.
nct_limits <- conf.limits.nct(t.value = tvalue, df=N-1, conf.level = ConfInt)
ci_l_d_av <- nct_limits$Lower.Limit*s_diff/(s_av*sqrt(N))
ci_u_d_av <- nct_limits$Upper.Limit*s_diff/(s_av*sqrt(N))

#Interpret size of effect (last resort - use only if effect size cannot be compared to other relevant effects in the literature)
if (abs(d_av) < 0.2){effectsize<-"tiny"}
if (0.2 <= abs(d_av) && abs(d_av) < 0.5){effectsize<-"small"}
if (0.5 <= abs(d_av) && abs(d_av) < 0.8){effectsize<-"medium"}
if (abs(d_av) >= 0.8){effectsize<-"large"}

#Cohen's d_z, using s_diff as standardizer
d_z <- tvalue/sqrt(N)
ci_l_d_z <- nct_limits$Lower.Limit/sqrt(N) #Not sure about this formula, but gives same results as Wuensch's files
ci_u_d_z <- nct_limits$Upper.Limit/sqrt(N) #Not sure about this formula

#Common Langaue Effect Size (McGraw & Wong, 1992)
CL<-pnorm(abs(m_diff/s_diff))

#Robust Statistics (if normality is violated)

require(WRS)

yuend<-yuendv2(x,y, tr=0.2,alpha=ConfInt)
if(yuend$est1>yuend$est2){direction2<-"greater than"}
if(yuend$est1<yuend$est2){direction2<-"smaller than"}
if(yuend$p.value < alpha){surprising2<-"surprising"}
if(yuend$p.value >= alpha){surprising2<-" not surprising"}

#Interpret size of effect (last resort - use only if effect size cannot be compared to other relevant effects in the literature)

if (abs(yuend$Effect.Size) < 0.15){effectsize2<-"tiny"}
if (0.15 <= abs(yuend$Effect.Size) && abs(yuend$Effect.Size) < 0.35){effectsize2<-"small"}
if (0.35 <= abs(yuend$Effect.Size) && abs(yuend$Effect.Size) < 0.5){effectsize2<-"medium"}
if (abs(yuend$Effect.Size) >= 0.5){effectsize2<-"large"}

#BayesFactor
require(BayesFactor) #To calculate Bayes Factor

if(H1 == "two.sided"){
  BF <- ttest.tstat(t = tvalue, n1 = N, rscale = BFrscale, simple=TRUE)
} else if (H1 == "greater"){
  BF <- ttest.tstat(t = tvalue, n1 = N, nullInterval = c(0, Inf), rscale = BFrscale, simple=TRUE)
} else if (H1 == "less"){
  BF <- ttest.tstat(t = tvalue, n1 = N, nullInterval = c(-Inf, 0), rscale = BFrscale, simple=TRUE)
}

if(BF!=Inf){round(BF, digits=2)}
if(BF==Inf){BF<-"practically infinitely high"}

#Download and install JAGS: http://sourceforge.net/projects/mcmc-jags/files/JAGS/

require(BEST) #To calculate HIB
if(InAHurry!="YES"){BESTout1g<-BESTmcmc(diff)}
ifelse((InAHurry!="YES"),mu<-summary(BESTout1g)[1,1],mu<-"NOT CALCULATED DUE TO TIME CONSTRAINTS")
if(InAHurry!="YES"){BESTHDI<-hdi(BESTout1g$mu, credMass = ConfInt)}
ifelse((InAHurry!="YES"),HDI_l<-BESTHDI[1],HDI_l<-"NOT CALCULATED")
ifelse((InAHurry!="YES"),HDI_u<-BESTHDI[2],HDI_u<-"NOT CALCULATED")
if(InAHurry!="YES"){Rhat<-attr(BESTout1g, "Rhat")}
if(InAHurry!="YES"){neff<-attr(BESTout1g, "n.eff")}
if(InAHurry!="YES"){ifelse(Rhat[1]<1.1 && neff[1]>10000, BESTacceptable<-"acceptable", BESTacceptable<-"not acceptable - check the HDI calculation")}

#Interpret strength of evidence of Bayes Factor following Jeffreys (1961)
if (0.33 < BF && BF <= 1){evidence<-"anecdotal evidence for H0"}
if (0.1 < BF && BF <=0.33){evidence<-"moderate evidence for H0"}
if (0.03 < BF && BF <= 0.1){evidence<-"strong evidence for H0"}
if (0.01 < BF && BF <= 0.03){evidence<-"very strong evidence for H0"}
if (BF <=0.01){evidence<-"decisive evidence for H0"}
if (1 < BF && BF <= 3){evidence<-"anecdotal evidence for H1"}
if (3 < BF && BF <=10){evidence<-"moderate evidence for H1"}
if (10 < BF && BF <= 30){evidence<-"strong evidence for H1"}
if (30 < BF && BF <= 100){evidence<-"very strong evidence for H1"}
if (BF > 100){evidence<-"decisive evidence for H1"}

```

This document summarizes a comparison between two independent groups, comparing `r ylabelstring` between the `r xlabel` and `r ylabel` conditions. This script can help to facilitate the analysis of data, and the word-output might prevent copy-paste errors when transferring results to a manuscript.

Researchers can base their statistical inferences on Frequentist or robust statistics, as well as on Bayesian statistics. Effect sizes and their confidence intervals are provided, thus inviting researchers to interpret their data from multiple perspectives.

Checking for outliers, normality, equality of variances.
------

###Outliers

Boxplots can be used to identify outliers. Boxplots give the median (thick line), and 25% of the data above and below the median (box). End of whiskers are the maximum and minimum value when excluding outliers (which are indicated by dots). Code adapted from [Sandy Muspratt](https://github.com/SandyMuspratt/ScatterBoxPlot/blob/master/mtcars%20marginal%20boxplots.R).

```{r, echo=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.heigth=4, dpi=300}
#Draw scatterplot with 2 boxplots
#Code adapted from: https://github.com/SandyMuspratt/ScatterBoxPlot/blob/master/mtcars%20marginal%20boxplots.R by Sandy Muspratt

require(ggplot2)
require(gtable)
require(grid)

# Main scatterplot
p1 <- ggplot(alldata, aes(x=eval(parse(text=paste(xlabel))), y=eval(parse(text=paste(ylabel))))) + 
  geom_point(alpha = 0.35) +
  scale_x_continuous(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0)) + 
  expand_limits(y = c(min(alldata[[ylabel]]) - 0.1 * diff(range(alldata[[ylabel]])), 
                      max(alldata[[ylabel]]) + 0.1 * diff(range(alldata[[ylabel]])))) + 
  expand_limits(x = c(min(alldata[[xlabel]]) - 0.1 * diff(range(alldata[[xlabel]])), 
                      max(alldata[[xlabel]]) + 0.1 * diff(range(alldata[[xlabel]])))) +
  theme_bw(base_size=14) +
  ylab(ylabel)  + xlab(xlabel) +
  theme(plot.margin= unit(c(0, 0, 0.5, 0.5), "lines"))

# To remove all axis labelling and marks from the two marginal plots
theme_remove_all <- theme(axis.text = element_blank(),
                          axis.title = element_blank(),
                          axis.ticks =  element_blank(),
                          axis.ticks.margin = unit(0, "lines"),
                          axis.ticks.length = unit(0, "cm"))

# Horizontal marginal boxplot - to appear at the top of the chart
p2 <- ggplot(alldata, aes(x = factor(1), y = eval(parse(text=paste(xlabel))))) + 
  geom_boxplot() +
#  geom_jitter(position = position_jitter(width = 0.05)) +
  scale_y_continuous(expand = c(0, 0)) + 
  expand_limits(y = c(min(alldata[[xlabel]]) - 0.1 * diff(range(alldata[[xlabel]])), 
                      max(alldata[[xlabel]]) + 0.1 * diff(range(alldata[[xlabel]])))) + 
  coord_flip() +
  theme_bw(base_size=14) +
  theme_remove_all +
  theme(plot.margin= unit(c(0.5, 0, 0, 0.5), "lines"))

# Vertical marginal boxplot - to appear at the right of the chart
p3 <- ggplot(alldata, aes(x = factor(1), y = eval(parse(text=paste(ylabel))))) + 
  geom_boxplot() +
#  geom_jitter(position = position_jitter(width = 0.05)) +
  scale_y_continuous(expand = c(0, 0)) + 
  expand_limits(y = c(min(alldata[[ylabel]]) - 0.1 * diff(range(alldata[[ylabel]])), 
                      max(alldata[[ylabel]]) + 0.1 * diff(range(alldata[[ylabel]])))) + 
  theme_bw(base_size=14) +
  theme_remove_all +
  theme(plot.margin= unit(c(0, 0.5, 0.5, 0), "lines"))

# Get the gtables
gt1 <- ggplot_gtable(ggplot_build(p1))
gt2 <- ggplot_gtable(ggplot_build(p2))
gt3 <- ggplot_gtable(ggplot_build(p3))

# Get maximum widths and heights for x-axis and y-axis title and text
maxWidth = unit.pmax(gt1$widths[2:3], gt2$widths[2:3])
maxHeight = unit.pmax(gt1$heights[4:5], gt3$heights[4:5])

# Set the maximums in the gtables for gt1, gt2 and gt3
gt1$widths[2:3] <- as.list(maxWidth)
gt2$widths[2:3] <- as.list(maxWidth)

gt1$heights[4:5] <- as.list(maxHeight)
gt3$heights[4:5] <- as.list(maxHeight)

# Combine the scatterplot with the two marginal boxplots
# Create a new gtable
gt <- gtable(widths = unit(c(7, 1), "null"), height = unit(c(1, 7), "null"))

# Instert gt1, gt2 and gt3 into the new gtable
gt <- gtable_add_grob(gt, gt1, 2, 1)
gt <- gtable_add_grob(gt, gt2, 1, 1)
gt <- gtable_add_grob(gt, gt3, 2, 2)

# And render the plot
grid.newpage()
grid.draw(gt)

```

##Normality assumption

The dependent *t*-test assumes that *difference* scores are normally distributed and that the variances of the two groups are equal. It does *not* assume the data within each measurement (so within the `r xlabel` and `r ylabel` condition) are normally distributed. If the normality assumption is violated, the Type 1 error rate of the test is no longer controlled, and can substantially increase beyond the chosen significance level. Formally, a normality test based on the data is incorrect, and the normality assumption should be tested on additional (e.g., pilot) data. Nevertheless, a two-step procedure (testing the data for normality, and using alternatives for the traditional *t*-test if normality is violated) works well (see [Rochon, Gondan, & Kieser, 2012](http://www.biomedcentral.com/1471-2288/12/81)).

###Tests for normality

[Yap and Sim (2011, p. 2153)](http://www.tandfonline.com/doi/pdf/10.1080/00949655.2010.520163) recommend: "If the distribution is symmetric with low kurtosis values (i.e. symmetric short-tailed distribution), then the D'Agostino-Pearson and Shapiro-Wilkes tests have good power. For symmetric distribution with high sample kurtosis (symmetric long-tailed), the researcher can use the JB, Shapiro-Wilkes, or Anderson-Darling test." The Kolmogorov-Smirnov (K-S) test is often used, but no longer recommended, and not included here.

If a normality test rejects the assumptions that the data is normally distributed (with *p* < .05) non-parametric or robust statistics have to be used (robust analyses are provided below).

**The normality assumption was rejected in `r normalityrejections` out of 4 normality tests (Anderson-Darling, D'Agostino-Pearson, and Shapiro-Wilk).**

Test Name  | *p*-value 
------------- | -------------
Shapiro-Wilk  | *p* `r ifelse(statcompute(21, diff, levels = c(0.05))$pvalue>=0.001," = ", " < ")` `r ifelse(statcompute(21, diff, levels = c(0.05))$pvalue>=0.001, round(statcompute(21, diff, levels = c(0.05))$pvalue, digits=3), "0.001")`
D'Agostino-Pearson  | *p* `r ifelse(statcompute(6, diff, levels = c(0.05))$pvalue>0.001," = ", " < ")` `r ifelse(statcompute(6, diff, levels = c(0.05))$pvalue>0.001, round(statcompute(6, diff, levels = c(0.05))$pvalue, digits=3), "0.001")`
Anderson-Darling  | *p* `r ifelse(statcompute(2, diff, levels = c(0.05))$pvalue>0.001," = ", " < ")` `r ifelse(statcompute(2, diff, levels = c(0.05))$pvalue>0.001, round(statcompute(2, diff, levels = c(0.05))$pvalue, digits=3), "0.001")`
Jarque-Berra  | *p* `r ifelse(statcompute(7, diff, levels = c(0.05))$pvalue>0.001," = ", " < ")` `r ifelse(statcompute(7, diff, levels = c(0.05))$pvalue>0.001, round(statcompute(7, diff, levels = c(0.05))$pvalue, digits=3), "0.001")`

In very large samples (when the test for normality has close to 100% power) tests for normality can result in significant results even when data is normally distributed, based on minor deviations from normality. In very small samples (e.g., n = 10), deviations from normality might not be detected, but this does not mean the data is normally distributed.  Always look at a plot of the data in addition to the test results.

###Histogram, kernel density plot (black line) and normal distribution (red line) of difference scores

The density (or proportion of the observations) is plotted on the y-axis. The grey bars are a histogram of the difference scores. Judging whether data is normally distributed on the basis of a histogram depends too much on the number of bins (or bars) in the graph. A kernel density plot (a non-parametric technique for density estimation) provides an easier way to check the normality of the data by comparing the shape of the density plot (the black line) with a normal distribution (the red dotted line, based on the observed mean and standard deviation). For dependent t-tests, the main DV is the *difference score*, and therefore the difference score should be normally distributed.

```{r, echo=FALSE, message=FALSE, fig.width=6.3, dpi=300}

#density plot with normal distribution (red) and kernel desity plot
ggplot(alldata, aes(x=diff))  + 
  geom_histogram(colour="black", fill="grey", aes(y = ..density..)) +
  stat_function(fun = dnorm, args = c(mean=mean(alldata$diff), sd=sd(alldata$diff)), size = 1, color = "red", lty=2) +
  geom_density(fill=NA, colour="black", size = 1) +
  xlab(ylabelstring) + ggtitle("Difference scores") + theme_bw(base_size=14) + 
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())

require(gridExtra)
#density plot with normal distribution (red) and kernel desity plot
p1<-ggplot(alldata, aes(x=eval(parse(text=paste(xlabel)))))  + 
  geom_histogram(colour="black", fill="grey", aes(y = ..density..)) +
  stat_function(fun = dnorm, args = c(mean=mean(x), sd=sd(x)), size = 1, color = "red", lty=2) +
  geom_density(fill=NA, colour="black", size = 1) +
  xlab(ylabelstring)  + ggtitle(xlabel)+ theme_bw(base_size=14) + 
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())

#density plot with normal distribution (red) and kernel desity plot
p2<-ggplot(alldata, aes(x=eval(parse(text=paste(ylabel)))))  + 
  geom_histogram(colour="black", fill="grey", aes(y = ..density..)) +
  stat_function(fun = dnorm, args = c(mean=mean(y), sd=sd(y)), size = 1, color = "red", lty=2) +
  geom_density(fill=NA, colour="black", size = 1) +
  xlab(ylabelstring) + ggtitle(ylabel) + theme_bw(base_size=14) + 
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())

  grid.arrange(p1, p2, nrow=2)

```

Q-Q-plot
------

In the Q-Q plot for the difference scores the points should fall on the line. Deviations from the line in the upper and lower quartiles indicates the tails of the distributions are thicker or thinner than in the normal distribution. An S-shaped curve with a dip in the middle indicates data is left-skewed (more values to the right of the distribution), while a bump in the middle indicates data is right-skewed (more values to the left of the distribution). For interpretation examples, see [here](http://emp.byui.edu/BrownD/Stats-intro/dscrptv/graphs/qq-plot_egs.htm).

```{r, echo=FALSE, message=FALSE, fig.width=6.3, fig.height=3.6, dpi=300}
require(HLMdiag)
#Q-Q plot
ggplot_qqnorm(diff, line = "quantile") + ggtitle("Difference scores") + theme_bw(base_size=14)  + 
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())

qq1<-ggplot_qqnorm(x, line = "quantile") + ggtitle(xlabel) + theme_bw(base_size=14)  + 
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())
qq2<-ggplot_qqnorm(y, line = "quantile") + ggtitle(ylabel) + theme_bw(base_size=14)  + 
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())
grid.arrange(qq1, qq2, ncol=2)

```

##Equal variances assumption

In addition to the normality assumption, a second assumption of the *t*-test is that variances in both groups are equal. The variance is the standard deviation, squared, and the assumption is thus that the variance in the `r xlabel` condition (`r round(sd1^2, digits = 2)`) equals that in the `r ylabel` condition (`r round(sd2^2, digits = 2)`). [Markowski & Markowski (1990)](http://www.jstor.org/stable/2684360) show that if sample sizes are equal, violations of the equal variance assumption do not lead to unsatisfactory performance (defined as actual significance levels falling outside a 0.03-0.07 boundary for a nominal alpha level of 0.05).

###Levene's test

This equality of variances assumption is typically examined with Levene's test, although in small samples, Levene's test can have low power, and thus fail to reject the null-hypothesis that variances are equal, even when they are unequal. Levene's test for equality of variances (*p* `r ifelse(pvalueLevene>0.001," = ", " < ")` `r ifelse(pvalueLevene>0.001, round(pvalueLevene, digits=3), "0.001")`) indicates that `r equalvar`

Comparing the two sets of data
------

Before looking at the results of the Frequentist statistics and the Robust statistics, decide which of these answer the question you are interested in. Choosing between these two options depending on the outcome of the statistical test inflates the Type 1 error rate. You can always report Bayesian statistics.

###Frequentist statistics

A *p*-value is the probability of obtaining the observed result, or a more extreme result, assuming the null-hypothesis is true. It is not the probability that the null-hypothesis or the alternative hypothesis is true (for such inferences, see Bayesian statistics below). In repeated sampling, `r 100*ConfInt`% of future `r 100*ConfInt`% confidence intervals can be expected to contain the true population parameters (e.g, the mean difference or the effect size). Confidence intervals are not a statement about the probability that a single confidence interval contains the true population parameter, but a statement about the probability that future confidence intervals will contain the true population parameter. Hedges' *g* (also referred to as *d*~unbiased~, see Borenstein, Hedges, Higgins, & Rothstein, 2009) is provided as best estimate of Cohen's *d*, but the best estimate of the confidence interval is based on *d*~av~ (as recommended by Cumming, 2012). Hedges's *g* and the `r 100*ConfInt`% CI around the effect size are calculated using the MBESS package by ([Kelley (2007](http://dx.doi.org/10.3758/BF03192993)). The common language effect size expresses the probability that in any random pairing of two observations from both groups, the observation from one group is higher than the observation from the other group, see [McGraw & Wong, 1992](http://dx.doi.org/10.1037/0033-2909.111.2.361). In a dependent *t*-test, the effect size Cohen's *d* can be calculated by using a standardizer that controls for the correlation between observations (*d*~av~) or not (*d*~z~). Both are provided, but *d*~av~ (or actually it's unbiased estimate, *g*~av~) is recommended. For a discussion, see [Lakens, 2013](http://journal.frontiersin.org/Journal/10.3389/fpsyg.2013.00863/full). Default interpretations of the size of an effect as provided here should only be used as a last resort, and it is preferable to interpret the size of the effect in relation to other effects in the literature, or in terms of its practical significance.

####Results

The mean `r ylabelstring` of participants in the `r xlabel` condition (*M* = `r round(mean(x), digits = 2)`, *SD* = `r round(sd1, digits = 2)`)  was `r direction` the mean of participants in the `r ylabel` condition (*M* = `r round(mean(y), digits = 2)`, *SD* = `r round(sd2,digits=2)`, *r* = `r round(r, digits = 2)`). The difference between measurements (*M* = `r round(m_diff, digits=2)`, *SD* = `r round(s_diff, digits=2)`, `r 100*ConfInt`% CI = [`r round(CI_diff[1:1], digits=2)`;`r round(CI_diff[2:2],digits=2)`]) was analyzed with a dependent *t*-test, *t*(`r round(ttestresult$parameter, digits=2)`) = `r round(tvalue, digits=2)`, *p* `r ifelse(pvalue>0.001," = ", " < ")` `r ifelse(pvalue>0.001, formatC(round(pvalue, digits=3),digits=3, format="f"), "0.001")`, Hedges' *g* = `r round(d_unb, digits=2)`, `r 100*ConfInt`% CI [`r round(ci_l_d_av, digits=2)`;`r round(ci_u_d_av, digits=2)`] (or *d*~z~ = `r round(d_z, digits=2)`, `r 100*ConfInt`% CI [`r round(ci_l_d_z, digits=2)`;`r round(ci_u_d_z, digits=2)`]). This can be considered a `r effectsize` effect. The observed data is `r surprising` under the assumption that the null-hypothesis is true. The Common Language effect size (McGraw & Wong, 1992) indicates that after controlling for individual differences, the likelihood that a persons `r ylabelstring` in the `r xlabel` condition is `r direction` the `r ylabelstring` in the `r ylabel` condition is `r round(100*CL, digits=0)`%.

###Bayesian statistics

Bayesian statistics can quantify the relative evidence in the data for either the alternative hypothesis or the null hypothesis. Bayesian statistics require priors to be defined. In the Bayes Factor calculation reported below, a non-informative Jeffreys prior is placed on the variance of the normal population, while a Cauchy prior is placed on the standardized effect size (for details, [see Morey & Rouder, 2011](http://drsmorey.org/bibtex/upload/Morey:Rouder:2011.pdf)). Calculations are performed using the [BayesFactor package](http://cran.r-project.org/web/packages/BayesFactor/BayesFactor.pdf). Default interpretations of the strength of the evidence are provided but should not distract from the fact that strength of evidence is a continuous function of the Bayes Factor. A second popular Bayesian approach relies on estimation, and the mean posterior and `r 100*ConfInt`% higest density intervals (HDI) are calculated following recommendations by [Kruschke, (2013)](http://www.indiana.edu/~kruschke/BEST/BEST.pdf) based on vague priors. According to Kruschke (2010, p. 34): 'The HDI indicates which points of a distribution we believe in most strongly. The width of the HDI is another way of measuring uncertainty of beliefs. If the HDI is wide, then beliefs are uncertain. If the HDI is narrow, then beliefs are fairly certain.' To check the convergence and fit of the HDI simulations, the Brooks-Gelman-Rubin scale reduction factor for the difference score should be smaller than 1.1 (it is `r ifelse (InAHurry!="YES",Rhat[1], "NOT CALCULATED")`) and the effective sample size should be larger than 10000 (it is `r ifelse (InAHurry!="YES",round(neff[1]), "NOT CALCULATED")`). Thus, the HDI simulation is `r ifelse (InAHurry!="YES",BESTacceptable, "NOT CALCULATED")`.

####Results

The JZS BF~10~ (with r scale = `r BFrscale`) = `r round(BF, digits=2)`. This indicates the data are `r round(BF, digits = 2)` (or log~e~ BF =`r round(log(BF), digits=2)`) times more probable under the alternative hypothesis, than under the null hypothesis. This data provides `r evidence`. The posterior mean difference is `r ifelse(InAHurry!="YES",round(mu, digits=2), "NOT CALCULATED")`, `r 100*ConfInt`% HDI = [`r ifelse(InAHurry!="YES",round(HDI_l, digits=2), "NOT CALCULATED")`; `r ifelse(InAHurry!="YES",round(HDI_u, digits=2), "NOT CALCULATED")`].

###Robust statistics

Values in the tails of the distribution can have a strong influence on the mean. If values in the tails differ from a normal distribution, the power of a test is reduced and the effect size estimates are biased, even under slight deviations from normality (Wilcox, 2012). One way to deal with this problem is to remove the tails in the analysis by using *trimmed means*. A recommended percentage of trimming is 20% from both tails (Wilcox, 2012), which means inferences are based on the 60% of the data in the middle of the distribution. Yuen's method can be used to compare trimmed means (when the percentage of trimming is 0%, Yuen's method reduces to Welch's *t*-test). The equivalent of Cohen's *d* for within designs is not yet available, so the explanatory effect size is reported ([Wilcox & Tian, 2011](http://dx.doi.org/10.1080/02664763.2010.498507)). Explanatory power (Xi, replace in the output below by the Greek lowercase Xi symbol) is the robust equivalent of omega squared (unbiased eta squared, or *r* squared), and thus related to *d*~z~ in size, not to *d*~av~. The effect size convention of small, medium, and large corresponds approximately to Xi = 0.15, 0.35 and 0.50.

####Results

The 20% trimmed mean `r ylabelstring` of participants in the `r xlabel` condition (*M* = `r round(yuend$est1, digits = 2)`)  was `r direction` the 20% trimmed mean of participants in the `r ylabel` condition (*M* = `r round(yuend$est2, digits = 2)`). The difference in `r ylabelstring` between the conditions (*M* = `r round(yuend$dif, digits = 2)`, `r 100*ConfInt`% CI [`r round(yuend$ci[1], digits = 2)`;`r round(yuend$ci[2], digits = 2)`]) was analyzed using the Yuen-Welch test for 20% trimmed means, *t*(`r yuend$df`) = `r round(yuend$teststat, digits = 2)`, *p* `r ifelse(yuend$p.value>0.001," = ", " < ")` `r ifelse(yuend$p.value>0.001, formatC(round(yuend$p.value, digits = 3), digits=3, format="f"), "0.001")`, Xi = `r round(yuend$Effect.Size, digits=2)`. The observed data is `r surprising2` under the assumption that the null-hypothesis is true. This can be considered a `r effectsize2` effect.

###*Figure 1*. Means, violin plot, and two-tiered `r 100*ConfInt`% within (crossbars) and between (endpoints of lines) confidence intervals following Morey (2008) and Baguley (2012).

```{r, echo=FALSE,  fig.width=3, dpi=300}
#Example 1: means and two-tiered 95% CI (within and between) suggested by Baguley 
ggplot(ci.sum, aes(x=eval(parse(text=paste(factorlabel))), y=eval(parse(text=paste(measurelabel))), group=1)) +
#  geom_bar(position=position_dodge(.9), colour="black", stat="identity", fill="white") +
  geom_errorbar(width=.25, size=0.5, aes(ymin=lower, ymax=upper)) +
  geom_errorbar(width=0, size=1, aes(ymin=lower.between, ymax=upper.between)) +
  geom_point(size=2) +
#  geom_point(data=alldata.long) +
  geom_violin(data=alldata.long, aes(group=eval(parse(text=paste(factorlabel)))), alpha=0) +
  ylab(ylabelstring)  + xlab(xlabelstring) + theme_bw(base_size=14) + 
  theme(panel.grid.major.x = element_blank())
```

###*Figure 2*. Means, datapoints, and `r 100*ConfInt`% CI (between & within)

```{r, echo=FALSE,  fig.width=3, dpi=300}
#Example 2: bar chart with individual data point and 95% CI (between)
ggplot(ci.sum, aes(x=eval(parse(text=paste(factorlabel))), y=eval(parse(text=paste(measurelabel))), group=1)) +
  geom_bar(position=position_dodge(.9), colour="black", stat="identity", fill="white") +
  geom_errorbar(width=.5, size=0.5, aes(ymin=lower, ymax=upper)) +
  geom_errorbar(width=.3, size=0.5, aes(ymin=lower.between, ymax=upper.between)) +
#  geom_point(size=4) +
#  geom_violin(data=alldata.long, aes(group=eval(parse(text=paste(factorlabel)))), alpha=0) +
   geom_point(data=alldata.long, alpha=0.25) +
   ylab(ylabelstring)  + xlab(xlabelstring) + theme_bw(base_size=14) + 
  theme(panel.grid.major.x = element_blank())
```

###*Figure 3*. Bar chart displaying means and `r 100*ConfInt`% CI (between and within)

```{r, echo=FALSE,  fig.width=3, dpi=300}
#Example 2: bar chart with individual data point and 95% CI (between)
ggplot(ci.sum, aes(x=eval(parse(text=paste(factorlabel))), y=eval(parse(text=paste(measurelabel))), group=1)) +
  geom_bar(position=position_dodge(.9), colour="black", stat="identity", fill="white") +
  geom_errorbar(width=.25, size=0.5, aes(ymin=lower, ymax=upper)) +
  geom_errorbar(width=0, size=1.1, aes(ymin=lower.between, ymax=upper.between)) +
#  geom_point(size=4) +
#  geom_point(data=alldata.long) +
  ylab(ylabelstring)  + xlab(xlabelstring) + theme_bw(base_size=14) + 
  theme(panel.grid.major.x = element_blank())
```

###References

This script uses the *reshape2* package to convert data from wide to long format, the *PoweR* package to perform the normality tests, *HLMdiag* to create the QQplots, *ggplot2* for all plots, *gtable* and *gridExtra* to combine multiple plots into one, *car* to perform Levene's test, *MBESS* to calculate effect sizes and their confidence intervals, *WRS* for the robust statistics, *BayesFactor* for the bayes factor, and *BEST* to calculate the Bayesian highest density interval.

Auguie, B. (2012). *gridExtra: functions in Grid graphics*. R package version 0.9.1, URL: http://CRAN.R-project.org/package=gridExtra.

Baguley, T. (2012). Calculating and graphing within-subject confidence intervals for ANOVA. *Behavior research methods*, *44*, 158-175.

Borenstein, M., Hedges, L. V., Higgins, J. P., & Rothstein, H. R. (2009). *Introduction to meta-analysis*. Hoboken, NJ: Wiley.

Box, G. E. P. (1953). Non-normality and tests on variance. *Biometrika*, *40*, 318-335.

Cumming, G. (2012). *Understanding the new statistics: Effect sizes, confidence intervals, and meta-analysis*. New York: Routledge.

Cohen, J. (1988). *Statistical power analysis for the behavioral sciences (2nd ed.)*. Hillsdale, NJ: Erlbaum.

Fox, J. & Weisberg, S. (2011). *An R Companion to Applied Regression, Second edition*. Sage, Thousand Oaks CA. URL: http://socserv.socsci.mcmaster.ca/jfox/Books/Companion.

Kelley, K. (2005). The effects of nonnormal distributions on confidence intervals around the standardized mean difference: Bootstrap and parametric confidence intervals. *Educational and Psychological Measurement*, *65*, 51-69.

Kelley, K. (2007). Confidence intervals for standardized effect sizes: Theory, application, and implementation. *Journal of Statistical Software*, *20*, 1-24.

Kelley, K. & Lai, K. (2012). *MBESS. R package version 3.3.3*, URL:
http://CRAN.R-project.org/package=MBESS.

Kruschke, J. (2010). *Doing Bayesian data analysis: A tutorial introduction with R*. Academic Press.

Kruschke, J. K. (2013). Bayesian estimation supersedes the t-test. *Journal of Experimental Psychology: General*, *142*, 573-603.

Kruschke, J. K., & Meredith, M. (2014). *BEST: Bayesian Estimation Supersedes the t-test*. R package version 0.2.2, URL: http://CRAN.R-project.org/package=BEST.

Lakens, D. (2013). Calculating and reporting effect sizes to facilitate cumulative science: a practical primer for t-tests and ANOVAs. *Frontiers in psychology*, *4*.

Loy, A., & Hofmann, H. (2014). HLMdiag: A Suite of Diagnostics for Hierarchical Linear Models. R. Journal of Statistical Software, 56, pp. 1-28. URL: http://www.jstatsoft.org/v56/i05/.

McGraw, K. O., & Wong, S. P. (1992). A common language effect size statistic. *Psychological Bulletin*, *111*, 361-365.

Micheaux, PLd. & Tran, V. (2012). PoweR. URL: http://www.biostatisticien.eu/PoweR/.

Morey, R. D. (2008). Confidence intervals from normalized data: A correction to Cousineau (2005). *Tutorial in Quantitative Methods for Psychology*, *4*, 61-64.

Morey, R. D. & Rouder, J. N. (2011). Bayes Factor Approaches for Testing Interval Null Hypotheses. *Psychological Methods*, *16*, 406-419

Morey R and Rouder J (2015). *BayesFactor: Computation of Bayes Factors for Common Designs*. R package version 0.9.11-1, URL: http://CRAN.R-project.org/package=BayesFactor.

Rochon, J., Gondan, M., & Kieser, M. (2012). To test or not to test: Preliminary assessment of normality when comparing two independent samples. *BMC Medical Research Methodology*, 12:81.

Rouder, J. N., Speckman, P. L., Sun, D., Morey, R. D., & Iverson, G. (2009). Bayesian t-tests for accepting and rejecting the null hypothesis. *Psychonomic Bulletin & Review*, *16*, 752-760

Ruxton, G. D. (2006). The unequal variance t-test is an underused alternative to Student's t-test and the Mann-Whitney U test. *Behavioral Ecology*, *17*, 688-690.

Wickham, H. (2007). Reshaping Data with the reshape Package. *Journal of Statistical Software*, *21*, pp. 1-20. URL: http://www.jstatsoft.org/v21/i12/.

Wickham, H. (2009). *ggplot2: elegant graphics for data analysis*. Springer New York. ISBN 978-0-387-98140-6, URL: http://had.co.nz/ggplot2/book.

Wickham, H. (2012). *gtable: Arrange grobs in tables*. R package version 0.1.2, URL: http://CRAN.R-project.org/package=gtable.

Wilcox, R. R. (2012). *Introduction to robust estimation and hypothesis testing*. Academic Press.

Wilcox, R. R., & Sch?nbrodt, F. D. (2015). *The WRS package for robust statistics in R (version 0.27.5)*. URL: https://github.com/nicebread/WRS.

Wilcox, R. R., & Tian, T. S. (2011). Measuring effect size: a robust heteroscedastic approach for two or more groups. *Journal of Applied Statistics*, *38*, 1359-1368.

Yap, B. W., & Sim, C. H. (2011). Comparisons of various types of normality tests. *Journal of Statistical Computation and Simulation*, *81*, 2141-2155.

###Apendix A: Data & Session Information

```{r}
alldata

sessionInfo()
```

Copyright ? 2015 Daniel Lakens

Lakens, D. (2015). The perfect *t*-test. Retrieved
from https://github.com/Lakens/perfect-t-test. doi:10.5281/zenodo.17603

This program is free software: you can redistribute it and/or modify it under the terms of the GNU Affero General Public License as published by the Free Software Foundation, either version 3 of the License, or (at your option) any later version.

This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. For more information, see the [GNU Affero General Public License](http://www.gnu.org/licenses/)

### 2017-08-17 E1/Code/ExperimentCode/DualRSVP_0and180degsRandom.m

@@ -1,0 +1,1185 @@
+% Dual RSVP stream demo code
+% PTG 11/07/12
+Screen('CloseAll');
+clear all; 
+
+addpath(genpath('/Users/supalab/Documents/MATLAB/BackwardsLetters/'));
+
+AssertOpenGL;
+
+% Initialise experiment parameters
+saveDirectory = '/Users/supalab/Documents/MATLAB/BackwardsLetters/Exp1_Random_0and180degs/Data/';
+userDirectory = '/Users/supalab/Docu  ments/MATLAB/BackwardsLetters/Exp1_Random_0and180degs/UserData/';
+nTrials = 100;            % 100, 50 (per condition)
+nPracticeTrials = 15;    % 15, 10
+nBlocks = 4;            % 4
+nConditions = 2;
+nSessions = 1;          % 2
+praticeOrder = 1; % 1 = canonical practice first, 2 = inverted practice first.
+
+% Randomise
+randomSeed = sum(100*clock);
+thisStream = RandStream.create('mt19937ar', 'seed', randomSeed);
+RandStream.setGlobalStream(thisStream); 
+
+
+while true
+    participantID = upper(input('Enter two-letter participant initials: ', 's'));
+    if length(participantID)==2
+        break
+    end
+end
+
+%{
+while true
+    thisConditionOrder = input('Enter condition order 1-6 (odd Canonical 0 degs start, even Inverted 180 degs start, 0 for random): ');
+    if any(thisConditionOrder == 1:6)
+        break
+    elseif thisConditionOrder == 0
+            thisConditionOrder = randi(6)
+        break
+    end
+end
+%}
+
+% Get practise order
+
+while true
+    praticeOrder = input('Enter first practise block condition 1 or 2 (1 = Canonical, 2 = Inverted): ');
+    if any(praticeOrder == 1:2)
+        break
+    end
+end
+
+
+% Update practice order and intructions
+
+if praticeOrder == 1
+    firstPracticeTxt =  ' in their normal orientation.\n\n';
+    nextPracticeTxt =  ' rotated 180 degrees.\n\n';
+    
+else
+    firstPracticeTxt = ' rotated 180 degrees.\n\n';  
+    nextPracticeTxt =  ' in their normal orientation.\n\n';
+    
+end
+
+
+oldDirectory = cd(userDirectory);
+userDataFile = [participantID '_RotatedDualRSVP_Exp1_UserData.mat'];
+
+if ~exist(userDataFile,'file')
[…]

### 2017-08-17 E1/Code/ModelingCode/E1_InvertedLetters_MixModCode.m

@@ -1,0 +1,256 @@
+clear all;
+close all;
+
+% SET ---------------------------------------------------------------------
+
+% Add directories
+
+addpath(genpath('/Users/admin/Documents/MATLAB/'));
+cd('/Users/admin/Documents/MATLAB/BackwardsLetters/Study_2/Exp1_Random_0and180degs/Data');
+
+% Task parameters
+
+letterArray = [char(65:66) char(68:85) char(87:90)];      % A to Z
+nLeadTrailItems = 6;            % Nuumber of items in stream at beginning and end with no target
+nConditions = 2;
+nStreams = 2;
+nTrials = 100;                   % Per participant, per condition, per session
+nSessions = 1;
+nBlocks = 4;
+itemRate = 10;
+% !!!!!!!!!!!!!!
+
+% Participant details
+
+%allParticipants =  {'AD', 'AE', 'AF', 'AG', 'AH', 'AI', 'AO', 'AP', 'AQ', 'AR', 'AU', 'AW', 'AX', 'BC', 'BD'};
+allParticipants = {'AA', 'AB', 'AC', 'AD', 'AE', 'AF', 'AG', 'AH', 'AI', 'AJ', 'AK', 'AL', 'AM', 'AN', 'AO', 'AP', 'AQ', 'AR', 'AS', 'AT', 'AU', 'AV', 'AW', 'AX', 'AY', 'AZ', 'BA', 'BB', 'BC', 'BD', 'BE', 'BF', 'BG', 'BH', 'BI', 'BJ', 'BK', 'BL', 'BM' ,'BN' ,'BO' ,'BP' ,'BQ' ,'BR', 'BS'};
+
+% Model fitting parameters
+
+nFreeParameters = 3;
+pdf_normmixture = @pdf_Mixture_Exp1;
+pdf_uniformonly = @pdf_Uniform_Exp1;
+nReplicates = 100;
+pCrit = .05;
+smallNonZeroNumber = 10^-10;
+fitMaxIter = 10^4;
+fitMaxFunEvals = 10^4;
+
+
+% CALCULATE ---------------------------------------------------------------
+
+% Task parameters
+totalTrials = (nTrials*nConditions)/2;
+nLetters = length(letterArray); % Number of possible letters
+maxError = nLetters-(nLeadTrailItems+1);
+errorValues = -maxError:1:maxError;
+offsetFactor = maxError+1;
+nErrorValues = numel(errorValues);
+
+% Participant details
+
+nParticipants = numel(allParticipants);
+
+
[…]

### 2017-08-17 E1/Code/ModelingCode/pdf_Uniform_Exp1.m

@@ -1,0 +1,16 @@
+function result = pdf_Uniform_Exp1(x,p)
+    
+    %fprintf('\np %0.2f, mu1 %2.2f, sigma1 %2.2f', p, mu1, sigma1);
+    % Need to taper the normal distribution with the pseudouniform
+    
+    pseudo_uniform = [0:1:11 12*ones(1,13) 11:-1:0];
+    normFactor_uniform = sum(pseudo_uniform);
+    
+    if normFactor_uniform == 0
+        normFactor_uniform = 0.00001;
+    end
+    
+    uniResultTemp = interp1(-18:1:18,pseudo_uniform,x);
+    result = p*(uniResultTemp/normFactor_uniform);
+
+end

### 2017-08-17 E1/Code/ModelingCode/removeSubjectsWithZeroEfficacy.m

@@ -1,0 +1,17 @@
+clear all;
+close all;
+
+cd('/Users/admin/Documents/MATLAB/BackwardsLetters/Study_2/Exp1_Random_0and180degs/Data');
+load('Exp1_Random_0and180degs_ModelRotated.mat');
+
+allParticipants = {'AA', 'AB', 'AC', 'AD', 'AE', 'AF', 'AG', 'AH', 'AI', 'AJ', 'AK', 'AL', 'AM', 'AN', 'AO', 'AP', 'AQ', 'AR', 'AS', 'AT', 'AU', 'AV', 'AW', 'AX', 'AY', 'AZ', 'BA', 'BB', 'BC', 'BD', 'BE', 'BF', 'BG', 'BH', 'BI', 'BJ', 'BK', 'BL', 'BM' ,'BN' ,'BO' ,'BP' ,'BQ' ,'BR', 'BS'};
+participantsToExclude = [5 6 7 13 14 16 17 18 19 20 25 28 34 38 39];
+
+subjectsForAnalysis = allParticipants;
+subjectsForAnalysis(participantsToExclude) = [];
+
+E1_allEstMinusExcluded = allEstimates_byParticipant;
+E1_allEstMinusExcluded(:,:,participantsToExclude,:) = [];
+
+cd('/Users/admin/Documents/MATLAB/BackwardsLetters/Study_2/Exp1_Random_0and180degs/Data');
+save('Exp1_InvertedLtrs_ParamsForAnalysis.mat','E1_allEstMinusExcluded','subjectsForAnalysis');

### 2017-08-17 E1/Code/R_AnalysisCode/GraphParams.R

@@ -1,0 +1,51 @@
+# Backwards Letters - Study 2 - Experiment 1
+
+setwd("/Users/admin/Desktop/WorkInProgress/BackwardsLetterStudy-2/E1/code/Rcode/")
+source("StatsFunctions.R")
+
+# Make efficacy graph
+efficacySummary <- summarySEwithin(efficacy.long, measurevar="Efficacy", withinvars=c("Stream","Orientation"), idvar="Subject")
+
+library(ggplot2)
+ggplot(efficacySummary, aes(x=Orientation, y=Efficacy, fill=Stream)) +
+  geom_bar(position=position_dodge(.9), colour="black", stat="identity") +
+  geom_errorbar(position=position_dodge(.9), width=.25, aes(ymin=Efficacy-ci, ymax=Efficacy+ci)) +
+  coord_cartesian(ylim=c(0,100)) +
+  scale_fill_manual(values=c("#CCCCCC","#FFFFFF")) +
+  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
+  theme_bw() + 
+  guides(fill = guide_legend(override.aes = list(colour = NULL))) +
+  theme(legend.key = element_rect(colour = "black"))
+
+
+# Make latency graph
+latencySummary <- summarySEwithin(latency.long, measurevar="Latency", withinvars=c("Stream","Orientation"), idvar="Subject")
+
+library(ggplot2)
+ggplot(latencySummary, aes(x=Orientation, y=Latency, fill=Stream)) +
+  ylab("Latency (ms)") +
+  geom_bar(position=position_dodge(.9), colour="black", stat="identity") +
+  geom_errorbar(position=position_dodge(.9), width=.25, aes(ymin=Latency-ci, ymax=Latency+ci)) +
+  coord_cartesian(ylim=c(0,150)) +
+  scale_fill_manual(values=c("#CCCCCC","#FFFFFF")) +
+  scale_y_continuous(breaks = seq(0, 150, by = 25)) +
+  theme_bw() + 
+  guides(fill = guide_legend(override.aes = list(colour = NULL))) +
+  theme(legend.key = element_rect(colour = "black"))
+
+
+# Make precsion graph
+precisionSummary <- summarySEwithin(precision.long, measurevar="Precision", withinvars=c("Stream","Orientation"), idvar="Subject")
+
+library(ggplot2)
[…]

### 2017-08-17 E1/Code/R_AnalysisCode/LoadParams.R

@@ -1,0 +1,106 @@
+# Backwards Letters - Study 2 - Experiment 1
+# This script loads the parameters (efficacy, latency, and precision) 
+# from the matlab data file--output from mixture modelling code. 
+# Creates a seperate data frame for each parameter in both wide and long format. 
+
+library(R.matlab)
+
+E1.data <- readMat('/Users/admin/Documents/MATLAB/BackwardsLetters/Study_2/Exp1_Random_0and180degs/Data/Exp1_InvertedLtrs_ParamsForAnalysis.mat')
+Subject <- c('AA','AB', 'AC', 'AD', 'AH',  'AI', 'AJ', 'AK', 'AL', 'AO', 'AU', 'AV', 'AW', 'AX', 'AZ', 'BA', 'BC', 'BD', 'BE', 'BF', 'BG', 'BI', 'BJ', 'BK', 'BN', 'BO', 'BP', 'BQ', 'BR', 'BS')
+
+
+# Create dataFrame for efficacy
+LeftCanon <- E1.data$E1.allEstMinusExcluded[1,1,,1]*100 # [Condition, Stream, Participant, Parameter]
+RightCanon <- E1.data$E1.allEstMinusExcluded[1,2,,1]*100 # Muliply by 100 to convert to percentage
+LeftInvert <- E1.data$E1.allEstMinusExcluded[2,2,,1]*100 # Left and right stream data are flipped in the
+RightInvert <- E1.data$E1.allEstMinusExcluded[2,1,,1]*100 # inverted condition so we unflip them here
+efficacy.wide <- data.frame(Subject, LeftCanon, RightCanon, LeftInvert, RightInvert)
+remove(LeftCanon, RightCanon, LeftInvert, RightInvert)
+
+
+# Create dataFrame for latency
+LeftCanon <- E1.data$E1.allEstMinusExcluded[1,1,,2]*(1000/10) # [Condition, Stream, Participant, Parameter]
+RightCanon <- E1.data$E1.allEstMinusExcluded[1,2,,2]*(1000/10) # Mutiply by 1000msec divided by item rate (10)
+LeftInvert <- E1.data$E1.allEstMinusExcluded[2,2,,2]*(1000/10) # Left and right stream data are flipped in the
+RightInvert <- E1.data$E1.allEstMinusExcluded[2,1,,2]*(1000/10) # inverted condition so we unflip them here
[…]

### 2017-08-17 E1/Code/R_AnalysisCode/StatsFunctions.R

@@ -1,0 +1,137 @@
+## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
+##   data: a data frame.
+##   measurevar: the name of a column that contains the variable to be summariezed
+##   groupvars: a vector containing names of columns that contain grouping variables
+##   na.rm: a boolean that indicates whether to ignore NA's
+##   conf.interval: the percent range of the confidence interval (default is 95%)
+summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
+                      conf.interval=.95, .drop=TRUE) {
+  library(plyr)
+  
+  # New version of length which can handle NA's: if na.rm==T, don't count them
+  length2 <- function (x, na.rm=FALSE) {
+    if (na.rm) sum(!is.na(x))
+    else       length(x)
+  }
+  
+  # This does the summary. For each group's data frame, return a vector with
+  # N, mean, and sd
+  datac <- ddply(data, groupvars, .drop=.drop,
+                 .fun = function(xx, col) {
+                   c(N    = length2(xx[[col]], na.rm=na.rm),
+                     mean = mean   (xx[[col]], na.rm=na.rm),
+                     sd   = sd     (xx[[col]], na.rm=na.rm)
+                   )
+                 },
+                 measurevar
+  )
+  
+  # Rename the "mean" column    
+  datac <- rename(datac, c("mean" = measurevar))
+  
+  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
+  
+  # Confidence interval multiplier for standard error
+  # Calculate t-statistic for confidence interval: 
+  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
+  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
+  datac$ci <- datac$se * ciMult
+  
+  return(datac)
+}
+
+
+## Norms the data within specified groups in a data frame; it normalizes each
+## subject (identified by idvar) so that they have the same mean, within each group
+## specified by betweenvars.
+##   data: a data frame.
+##   idvar: the name of a column that identifies each subject (or matched subjects)
+##   measurevar: the name of a column that contains the variable to be summariezed
+##   betweenvars: a vector containing names of columns that are between-subjects variables
+##   na.rm: a boolean that indicates whether to ignore NA's
[…]

### 2017-08-17 E2/DualRSVP_randomOrientationWithinStreams.m

@@ -1,0 +1,1136 @@
+% Dual RSVP stream demo code
+% PTG 11/07/12
+Screen('CloseAll');
+clear all; 
+
+addpath(genpath('/Users/supalab/Documents/MATLAB/BackwardsLetters/E2/'));
+
+AssertOpenGL;
+
+% Initialise experiment parameters
+saveDirectory = '/Users/supalab/Documents/MATLAB/BackwardsLetters/E2/Data/';
+userDirectory = '/Users/supalab/Documents/MATLAB/BackwardsLetters/E2/UserData/';
+nTrials = 100;            % 100, 50 (per condition)
+nPracticeTrials = 16;    % 15, 10
+nBlocks = 4;            % 4
+nConditions = 2;
+nSessions = 1;          % 2
+praticeOrder = 1; % 1 = canonical practice first, 2 = inverted practice first.
+
+% Randomise
+randomSeed = sum(100*clock);
+thisStream = RandStream.create('mt19937ar', 'seed', randomSeed);
+RandStream.setGlobalStream(thisStream); 
+
+
+while true
+    participantID = upper(input('Enter two-letter participant initials: ', 's'));
+    if length(participantID)==2
+        break
+    end
+end
+
+
+% Get practise order
+
+while true
+    praticeOrder = input('Enter first practise block condition 1 or 2 (1 = Canonical, 2 = Inverted): ');
+    if any(praticeOrder == 1:2)
+        break
+    end
+end
+
+
+% Update practice order and intructions
+
+if praticeOrder == 1
+    firstPracticeTxt =  ' in their normal upright orientation.\n\n';
+    nextPracticeTxt =  ' rotated 180 degrees.\n\n';
+    
+else
+    firstPracticeTxt = ' rotated 180 degrees.\n\n';  
+    nextPracticeTxt =  ' in their normal upright orientation.\n\n';
+    
+end
+
+
+oldDirectory = cd(userDirectory);
+userDataFile = [participantID '_RotatedDualRSVP_Exp1_UserData.mat'];
+
+if ~exist(userDataFile,'file')
+   % Make data file
+   % Check number of existing records
+   allUserDataFiles = what;
+   numUserDataFiles = size(allUserDataFiles.mat,1);
+   
+   % Assign next session
+   sessionOrder = randperm(2);
+   nextSession = 1;
+   totalTime = [];
+   
+   % Save
+   save(userDataFile,'sessionOrder','nextSession','totalTime');
[…]

### 2017-08-17 E2/ModelingCode/E1_InvertedLetters_MixModCode.m

@@ -1,0 +1,256 @@
+clear all;
+close all;
+
+% SET ---------------------------------------------------------------------
+
+% Add directories
+
+addpath(genpath('/Users/admin/Documents/MATLAB/'));
+cd('/Users/admin/Documents/MATLAB/BackwardsLetters/Study_2/Exp1_Random_0and180degs/Data');
+
+% Task parameters
+
+letterArray = [char(65:66) char(68:85) char(87:90)];      % A to Z
+nLeadTrailItems = 6;            % Nuumber of items in stream at beginning and end with no target
+nConditions = 2;
+nStreams = 2;
+nTrials = 100;                   % Per participant, per condition, per session
+nSessions = 1;
+nBlocks = 4;
+itemRate = 8;
+% !!!!!!!!!!!!!!
+
+% Participant details
+
+%allParticipants =  {'AD', 'AE', 'AF', 'AG', 'AH', 'AI', 'AO', 'AP', 'AQ', 'AR', 'AU', 'AW', 'AX', 'BC', 'BD'};
+allParticipants = {'AA', 'AB', 'AC', 'AD', 'AE', 'AF', 'AG', 'AH', 'AI', 'AJ', 'AK', 'AL', 'AM', 'AN', 'AO', 'AP', 'AQ', 'AR', 'AS', 'AT', 'AU', 'AV', 'AW', 'AX', 'AY', 'AZ', 'BA', 'BB', 'BC', 'BD', 'BE', 'BF', 'BG', 'BH', 'BI', 'BJ', 'BK', 'BL', 'BM' ,'BN' ,'BO' ,'BP' ,'BQ' ,'BR', 'BS'};
+
+% Model fitting parameters
+
+nFreeParameters = 3;
+pdf_normmixture = @pdf_Mixture_Exp1;
+pdf_uniformonly = @pdf_Uniform_Exp1;
+nReplicates = 100;
+pCrit = .05;
+smallNonZeroNumber = 10^-10;
+fitMaxIter = 10^4;
+fitMaxFunEvals = 10^4;
+
+
+% CALCULATE ---------------------------------------------------------------
+
+% Task parameters
+totalTrials = (nTrials*nConditions)/2;
+nLetters = length(letterArray); % Number of possible letters
+maxError = nLetters-(nLeadTrailItems+1);
+errorValues = -maxError:1:maxError;
+offsetFactor = maxError+1;
+nErrorValues = numel(errorValues);
+
+% Participant details
+
+nParticipants = numel(allParticipants);
+
+
[…]

### 2017-08-17 E2/ModelingCode/pdf_Mixture_Exp1.m

@@ -1,0 +1,32 @@
+function result = pdf_Mixture_Exp1(x,p,mu1,sigma1)
+    
+    pseudo_uniform = [0:1:11 12*ones(1,13) 11:-1:0];
+    pseudo_normal = normpdf(-18:1:18,mu1,sigma1).*pseudo_uniform;
+    
+    normFactor_uniform = sum(pseudo_uniform);
+    normFactor_normal = sum(pseudo_normal);
+    
+    if normFactor_uniform == 0
+        normFactor_uniform = 0.00001;
+    end
+    
+    if normFactor_normal == 0
+        normFactor_normal = 0.00001;
+    end
+    
+    uniResultTemp = interp1(-18:1:18,pseudo_uniform,x);
+    normResultTemp = normpdf(x,mu1,sigma1).*uniResultTemp;
+    
+    uniResultTemp = uniResultTemp/normFactor_uniform;
+    normResultTemp = normResultTemp/normFactor_normal;
+   
+    normResult = p*normResultTemp;
+    uniResult = (1-p)*uniResultTemp;
+    
+    if sum(size(normResult)==size(uniResult))==2
+        result = normResult+uniResult;
+    else
+        result = normResult+uniResult';
+    end
+
+end

### 2017-08-17 E2/ModelingCode/pdf_Normal_Exp1.m

@@ -1,0 +1,19 @@
+function result = pdf_Normal_Exp1(x,p,mu1,sigma1)
+    
+    pseudo_uniform = [0:1:11 12*ones(1,13) 11:-1:0];
+    pseudo_normal = normpdf(-18:1:18,mu1,sigma1).*pseudo_uniform;
+    
+    normFactor_normal = sum(pseudo_normal);
+    
+    if normFactor_normal == 0
+        normFactor_normal = 0.00001;
+    end
+    
+    uniResultTemp = interp1(-18:1:18,pseudo_uniform,x);
+    normResultTemp = normpdf(x,mu1,sigma1).*uniResultTemp;
+    
+    normResultTemp = normResultTemp/normFactor_normal;
+   
+    result = p*normResultTemp;
+
+end

### 2018-02-27 E2/ModelingCode/E1_InvertedLetters_MixModCode.m
seems this was left uncommitted for many months

@@ -6,12 +6,14 @@ close all;
 % Add directories
 
 addpath(genpath('/Users/admin/Documents/MATLAB/'));
-cd('/Users/admin/Documents/MATLAB/BackwardsLetters/Study_2/Exp1_Random_0and180degs/Data');
+addpath(genpath('~/Google Drive/Backwards paper/secondPaper/E2/ModelingCode'));
+cd('~/Google Drive/Backwards paper/secondPaper/E2/Data');
+
 
 % Task parameters
 
-letterArray = [char(65:66) char(68:85) char(87:90)];      % A to Z
-nLeadTrailItems = 6;            % Nuumber of items in stream at beginning and end with no target
+letterArray = [char(65:71) char(74:76) char(81:82) char(84:86) char(89)];      % A to Z
+nLeadTrailItems = 4;            % Nuumber of items in stream at beginning and end with no target
 nConditions = 2;
 nStreams = 2;
 nTrials = 100;                   % Per participant, per condition, per session
@@ -23,14 +25,14 @@ itemRate = 8;
 % Participant details
 
 %allParticipants =  {'AD', 'AE', 'AF', 'AG', 'AH', 'AI', 'AO', 'AP', 'AQ', 'AR', 'AU', 'AW', 'AX', 'BC', 'BD'};
-allParticipants = {'AA', 'AB', 'AC', 'AD', 'AE', 'AF', 'AG', 'AH', 'AI', 'AJ', 'AK', 'AL', 'AM', 'AN', 'AO', 'AP', 'AQ', 'AR', 'AS', 'AT', 'AU', 'AV', 'AW', 'AX', 'AY', 'AZ', 'BA', 'BB', 'BC', 'BD', 'BE', 'BF', 'BG', 'BH', 'BI', 'BJ', 'BK', 'BL', 'BM' ,'BN' ,'BO' ,'BP' ,'BQ' ,'BR', 'BS'};
+allParticipants = {'CB','CL','PP'};
 
 % Model fitting parameters
 
 nFreeParameters = 3;
 pdf_normmixture = @pdf_Mixture_Exp1;
 pdf_uniformonly = @pdf_Uniform_Exp1;
-nReplicates = 100;
+nReplicates = 300;
 pCrit = .05;
 smallNonZeroNumber = 10^-10;
 fitMaxIter = 10^4;
@@ -212,7 +214,7 @@ for thisCondition = 1:nConditions
[…]

### 2018-02-27 E1/importE1data.Rmd
adding new files to start analysis from scratch (previously it was done during development of mixture modeling before creating package

---
title: "importE2data"
author: "Alex Holcombe"
date: "2/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Modified from backwardsLtrsLoadRawData.R

```{r set up functions}

#Loads raw data from MATLAB files for second backwards paper first experiment
require(R.matlab)

GoogleDrivePath<-"." # "Google\ Drive/Backwards\ paper"

#First experiment of second backwards-letters paper. random orientation each trial.
rawDataPath<- file.path("~",GoogleDrivePath,
                            "E1/Data/RawData/Data")
#raw data path containing .mat file for each subject
rawDataExcludedByMATLABpath<- file.path(rawDataPath, "Excluded")

source(file.path("turnMATintoMeltedDataframe.R"))

readInAllFiles<- function(rawDataPath) {

  files <- dir(path=rawDataPath,pattern='.mat')  #find all data files in this directory
  dfAll<-data.frame()
  for (i in 1:length(files)) { #read in each file
    fileThis<- file.path(rawDataPath,files[i])
    rawDataLoad=tryCatch(
      readMat(fileThis),
      error=function(e) {
        stop( paste0("ERROR reading the file ",fname," :",e) )
      } )
    apparentSubjectName <- strsplit(files[i],split="_")[[1]][1]
    subjectName<- rawDataLoad$participantID[[1]]
    if (apparentSubjectName != subjectName) {
      stop( paste0("WARNING apparentSubjectName",apparentSubjectName," from filename does not match subjectName in data structure",subjectName) )
    }
    rawDataLoad$file <- files[i]

    dfThis<- turnMATintoMeltedDataframe(rawDataLoad)

    tryCatch(
      dfAll<-rbind(dfAll,dfThis), #if fail to bind new with old,
      error=function(e) { #Give feedback about how the error happened
        cat(paste0("Tried to merge but error:",e) )
      } )
  }
  return(dfAll)
}
```

```{r read raw data}

E<- readInAllFiles(rawDataPath)
#Calculate the serial position error
E$SPE<- E$respSP - E$targetSP

sanityVisualCheck<-FALSE
if (sanityVisualCheck) {
  library(ggplot2)
  g=ggplot(E, aes(x=SPE))
  #plot data
  g<-g+geom_histogram(binwidth=1)
  g
}

#Excluded subfolder contains Ss Chris excluded for low efficacy
excludedSsPath<- file.path(rawDataPath,"Excluded")

excluded<- readInAllFiles(excludedSsPath)
#Calculate the serial position error
excluded$SPE<- excluded$respSP - excluded$targetSP
```

```{r sanity visual check}

sanityVisualCheck<-TRUE
if (sanityVisualCheck) {
  library(ggplot2)
  g=ggplot(excluded, aes(x=SPE))
  #plot data
  g<-g+geom_histogram(binwidth=1)
  g
}

```

#Meaning of condition variable
#1 - upright
#2 - inverted

Combine excluded and non-excluded (based on old MATLAB anaylsis) participants
```{r combine}

E$excluded<-FALSE
excluded$excluded<-TRUE

backwards2_E1<-rbind(E,excluded)

#save dataframe into proper library data directory
save(backwards2_E1, file = "data/backwards2_E1.rda", compress="bzip2")

```

### 2018-02-27 E1/importE1data.Rmd
starting to mixture model

@@
 #First experiment of second backwards-letters paper. random orientation each trial.
-rawDataPath<- file.path("~",GoogleDrivePath,
-                            "E1/Data/RawData/Data")
+rawDataPath<- file.path(GoogleDrivePath,
+                            "Data/RawData/Data")
 #raw data path containing .mat file for each subject
@@
       error=function(e) {
-        stop( paste0("ERROR reading the file ",fname," :",e) )
+        stop( paste0("ERROR reading the file ",fileThis," :",e) )
       } )
@@
 
+Sanity visual check that see histogram.
 ```{r sanity visual check}
@@
   g<-g+geom_histogram(binwidth=1)
-  g
+  show(g)
+
+  h<-ggplot(E,aes(x=SPE)) + geom_histogram(binwidth=1)
+  show(h)
 }
@@
 
-#Meaning of condition variable
-#1 - upright
-#2 - inverted
+Meaning of condition variable
+* 1 - upright
+* 2 - inverted
 
@@
 #save dataframe into proper library data directory
-save(backwards2_E1, file = "data/backwards2_E1.rda", compress="bzip2")
-
+#save(backwards2_E1, file = "data/backwards2E1_rawDataFromMAT.rda", compress="bzip2")
+saveRDS(backwards2_E1, file = "Data/backwards2E1_rawDataFromMAT.rda", compress="bzip2")
 ```

### 2018-02-27 E1/mixModelE1data.Rmd
starting to mixture model

---
title: "mixModelE1data"
author: "Alex Holcombe"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE,  comment = NA)
# http://htmlpreview.github.io/?https://github.com/alexholcombe/MixtureModelRSVP/blob/master/compareMATLABtoR.html
```

Load data

```{r load data, echo=FALSE, message=FALSE}

#Import raw data
dataPath<- file.path("Data/")
#Experiment was administered by MATLAB
#.mat file been preprocessed into melted long dataframe by importE1data.Rmd
data<- readRDS( file.path(dataPath, "backwards2E1_rawDataFromMAT.rda") )

#tidy data
library(dplyr)
df<- data
#to work with dplyr, can't have array field like letterSeq
df$letterSeq<- NULL

numItemsInStream<- length( data$letterSeq[1,] )  
minSPE<- -17; maxSPE<- 17

df<-df %>% rename(stream = target, orientation = condition )
df <- df %>% mutate( stream =ifelse(stream==1, "Left","Right") )
#mutate condition to Orientation
df <- df %>% mutate( orientation =ifelse(orientation==1, "Canonical","Inverted") )

```

Fit mixture model with R (if fitting not previously done and cached). 
```{r, echo=FALSE, message=FALSE, cache=TRUE}
#Setting cache = TRUE in hopes won't have to recalculate estimates "if cached results exist and this code chunk has not been changed since last run" 
library(mixRSVP)

#Check whether already have parameter estimates or instead need to do it
calculate<-FALSE
if (!exists("estimates")) { 
  calculate<-TRUE 
}

condtnVariableNames <- c("subject","stream", "orientation") #

if (calculate) {
  estimates<- df %>%   filter(subject=="AA") %>%
    group_by_(.dots = condtnVariableNames) %>%  #.dots needed when you have a variable containing multiple factor names
    do(  analyzeOneCondition(.,numItemsInStream,parameterBounds(),nReplicates=5) )
  estimates<- estimates %>% rename(efficacy = p1, latency = p2, precision = p3)

}
#head(estimates)
```

Unfortunately the differences between MATLAB and R are not trivial. Not substantial either! but we'd prefer trivial.

```{r, echo=FALSE, message=FALSE}
#Compare MATLAB parameter estimates to R
estimates_MATLAB<- estimates_MATLAB %>% rename(efficacy_M = efficacy, latency_M = latency, precision_M = precision, val_M = val)
merged<- merge(estimates,estimates_MATLAB)  
merged<- merged %>% mutate( effDiff = efficacy-efficacy_M, latDiff= latency_M-latency, preDiff= precision_M-precision, negLogL_Diff = val - val_M )

#show differences columns only
diffs<- select(merged, ends_with("Diff"))
```

The mean discrepancy (mean of the absolute value of the difference of the parameter estimates) for
efficacy (effDiff), latency (latDiff), and precision (preDiff) is
```{r, echo=FALSE, message=FALSE}
#calc discrepancies
#Don't include negLogL because absolute value not interpretable
meanDiscrepancy<- diffs %>% select(-negLogL_Diff) 
meanDiscrepancy<- summarise_all(abs(meanDiscrepancy),mean)
round(meanDiscrepancy,3)
```

Now we report the bias difference:  the mean difference without taking the absolute value. Note that the negLogLikelihood for the MATLAB estimates is calculated by R, because Pat's code doesn't save the negative log likelihoods. Also in R we now use integration of the whole area under the bin, whereas Pat used the height of the density function in the center of the bin.
```{r, echo=FALSE, message=FALSE}
biasDiff<- summarise_all(diffs,mean)
round(biasDiff,3)
```
Negative numbers above mean that MATLAB gives slightly higher efficacy and trivially higher precision.
For likelihood, lower neg log likelihood is better, so the negative number means R did better.

Let's inspect histogram plots and fits. Note that in every case, the negative log likelihood is better (smaller) for R, showing that the R code provides a better fit than the MATLAB code.

* yellow = guessing component 
* light blue = Gaussian component
* green = sum of the guessing and Gaussian components. In other words, the histogram heights predicted by the model
* dark blue = continuous Gaussian. This helps get a sense of the effect of discretising the Gaussian. For instance, it's possible (especially using Pat's method, it seems), for the Gaussian peak to fly high above the bars and still fit the discrete bins (or bin centers, in Pat's method), suggesting an undesirably high estimates of the efficacy (likely accompanied by an undesirably low precision)

```{r, echo=FALSE, message=FALSE, fig.height=360, fig.width=10}
#want fig.height of 10 per subject

source( file.path(mixModelingPath,"calcCurvesDataframes.R") )

#create R curves
df<- df # %>%  dplyr::filter(subject <="BB") #dplyr::filter(subject <= "BD" & subject >="AP")
#dplyr::filter(subject=="AA",orientation=="Canonical")
source( file.path(mixModelingPath,"histogramPlotting.R") ) #for calcFitDataframes
source( file.path(mixModelingPath,"theme_apa.R") ) #for calcFitDataframes

#Add R parameter estimates to dataframe
df<- merge(df,estimates)

curvesR<- df %>% group_by_at(.vars = condtnVariableNames) %>% 
  do(calcCurvesDataframes(.,minSPE,maxSPE,numItemsInStream))

#Calc numObservations to each condition. This is needed only for scaling the fine-grained Gaussian
#Calc the number of observations for each condition, because gaussianScaledforData needs to know.
dfGroups<- df %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
#add nPerCond back to parameter estimates
estimates<- merge(estimates,dfGroups)
grain<-.05
gaussianFineR<- estimates %>% group_by_at(.vars = condtnVariableNames) %>% do(
  gaussianScaledFromDataframe(.,minSPE,maxSPE,grain) )

#create MATLAB curves
#need to rename parameters so that calcCurvesDataframes can find them
estimates_M<- estimates_MATLAB %>% rename(efficacy=efficacy_M,latency=latency_M,precision=precision_M)
#create temporary dataframe with data plus MATLAB estimates to generate curves
dataStrippedOfEstimates<- select(df,-efficacy,-latency,-precision, -val)
dM<-merge( dataStrippedOfEstimates, estimates_M )
dM<- dM %>% rename( val = val_M ) #has to be called val so calcFitDataframes uses it rather than using R to calculate its own
curvesMATLAB<- dM %>% group_by_at(.vars = condtnVariableNames) %>% 
            do(calcCurvesDataframes(.,minSPE,maxSPE,numItemsInStream))

#Calculate fine-grained Gaussian for MATLAB
#Calc the number of observations for each condition, because gaussianScaledforData needs to know.
dNum<- dataStrippedOfEstimates %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
#add nPerCond to MATLAB estimates, which will then be merged with raw data
estimates_M<- merge(estimates_M,dNum)
#merge MATLAB estimates with raw data
dM<-merge( dataStrippedOfEstimates, estimates_M )
gaussianFineM<- estimates_M %>% group_by_at(.vars = condtnVariableNames) %>% do(
  gaussianScaledFromDataframe(.,minSPE,maxSPE,grain) )

#Concatenate R and MATLAB into single dataframe
curvesR$lang<-"R"; curvesMATLAB$lang<-"MATLAB"
curves_R_MATLAB <- rbind( data.frame(curvesR), data.frame(curvesMATLAB) )
#To facilitate plotting,
#make dataframe containing the raw data twice, one for R plots column, one for MATLAB plots column (data identical)
dfBoth<-rbind(   df %>% mutate(lang="R"), df %>% mutate(lang="MATLAB")  )

#Concatenate fine-grained gaussian for R and MATLAB into single dataframe, so can plot both with ggplot
gaussianFineR$lang<-"R"; gaussianFineM$lang<-"MATLAB"; 
#Get the col names to match by only using a few columns
gaussianFineR<- gaussianFineR %>% select(-val,-warnings)
gaussianFineM<- gaussianFineM %>% select(-val_M)
gaussFine_R_M <- rbind( data.frame(gaussianFineR), data.frame(gaussianFineM) )

#PLOT EVERYTHING
g=ggplot(dfBoth, aes(x=SPE)) + facet_grid(subject+stream+orientation~lang,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g +theme_apa() #+theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank())# hide all gridlines.
#g<-g+ theme(line=element_blank(), panel.border = element_blank())
sz=.8
#Plot the underlying Gaussian , not just the discretized Gaussian
g<-g + geom_line(data=gaussFine_R_M,aes(x=x,y=gaussianFreq),color="darkblue",size=1.2)

g<-g+ geom_point(data=curves_R_MATLAB,aes(x=x,y=combinedFitFreq),color="chartreuse3",size=sz*2.5)
g<-g+ geom_line(data=curves_R_MATLAB,aes(x=x,y=guessingFreq),color="yellow",size=sz)
#Discretized Gaussian
g<-g+ geom_line(data=curves_R_MATLAB,aes(x=x,y=gaussianFreq),color="lightblue",size=sz)

numGroups<- length(table(dfBoth$orientation,dfBoth$subject,dfBoth$stream))
fontSz = 400/numGroups
g<-g + geom_text(data=curves_R_MATLAB,aes(x=-9,y=32, label = paste("-logLik==", round(val,1), sep = "")),  parse = TRUE,size=fontSz) +
  geom_text(data=curves_R_MATLAB,aes(x=-7,y=28, label = paste("plain(e)==", round(efficacy,2), sep = "")),  parse = TRUE,size=fontSz) +
  geom_text(data=curves_R_MATLAB,aes(x=-7,y=25, label = paste("mu==", round(latency,2), sep = "")),  parse = TRUE,size=fontSz)+
  geom_text(data=curves_R_MATLAB,aes(x=-7,y=22, label = paste("sigma==", round(precision,2), sep = "")),  parse = TRUE,size=fontSz)
show(g)
```

### 2018-02-27 E1/mixModelE1data.Rmd
mix model initial pilot

@@
 Fit mixture model with R (if fitting not previously done and cached). 
-```{r, echo=FALSE, message=FALSE, cache=TRUE}
+```{r, echo=FALSE, cache=TRUE}
 #Setting cache = TRUE in hopes won't have to recalculate estimates "if cached results exist and this code chunk has not been changed since last run" 
@@
 
-condtnVariableNames <- c("subject","stream", "orientation") #
+condtnVariableNames <- c("subject","stream", "orientation")
 
@@
     group_by_(.dots = condtnVariableNames) %>%  #.dots needed when you have a variable containing multiple factor names
-    do(  analyzeOneCondition(.,numItemsInStream,parameterBounds(),nReplicates=5) )
+    do(  analyzeOneConditionDF(.,numItemsInStream,parameterBounds(),nReplicates=5) )
   estimates<- estimates %>% rename(efficacy = p1, latency = p2, precision = p3)
@@
 }
-#head(estimates)
+head(estimates)
 ```
 
-Unfortunately the differences between MATLAB and R are not trivial. Not substantial either! but we'd prefer trivial.
+Save the param estimates to a file
+```{Save param estimates, echo=FALSE}
 
-```{r, echo=FALSE, message=FALSE}
-#Compare MATLAB parameter estimates to R
-estimates_MATLAB<- estimates_MATLAB %>% rename(efficacy_M = efficacy, latency_M = latency, precision_M = precision, val_M = val)
-merged<- merge(estimates,estimates_MATLAB)  
-merged<- merged %>% mutate( effDiff = efficacy-efficacy_M, latDiff= latency_M-latency, preDiff= precision_M-precision, negLogL_Diff = val - val_M )
+write.csv(estimates, file = "Data/backwards2E1_paramEstimates")
 
-#show differences columns only
-diffs<- select(merged, ends_with("Diff"))
 ```
-
-The mean discrepancy (mean of the absolute value of the difference of the parameter estimates) for
-efficacy (effDiff), latency (latDiff), and precision (preDiff) is
-```{r, echo=FALSE, message=FALSE}
-#calc discrepancies
-#Don't include negLogL because absolute value not interpretable
-meanDiscrepancy<- diffs %>% select(-negLogL_Diff) 
-meanDiscrepancy<- summarise_all(abs(meanDiscrepancy),mean)
-round(meanDiscrepancy,3)
-```
-
-Now we report the bias difference:  the mean difference without taking the absolute value.
-Note that the negLogLikelihood for the MATLAB estimates is calculated by R, because Pat's code doesn't save the negative log likelihoods.
-Also in R we now use integration of the whole area under the bin, whereas Pat used the height of the density function in the center of the bin.
-```{r, echo=FALSE, message=FALSE}
-biasDiff<- summarise_all(diffs,mean)
-round(biasDiff,3)
-```
-Negative numbers above mean that MATLAB gives slightly higher efficacy and trivially higher precision.
-For likelihood, lower neg log likelihood is better, so the negative number means R did better.
 

### 2018-02-28 E1/importE1data.Rmd
looking good

@@
 
+```
+
+Make condition names more sensible - stream (Left/Right) and orientation (Canonical/Inverted)
+```{r}
+names(backwards2_E1)[names(backwards2_E1) == 'target'] <-'stream'
+names(backwards2_E1)[names(backwards2_E1) == 'condition'] <-'orientation'
+backwards2_E1$stream[ backwards2_E1$stream==1 ] <- "Left"
+backwards2_E1$stream[ backwards2_E1$stream==2 ] <- "Right"
+#mutate condition to Orientation
+backwards2_E1$orientation[ backwards2_E1$orientation==1 ] <- "Canonical"
+backwards2_E1$orientation[ backwards2_E1$orientation==2 ] <- "Inverted"
+
+```
+
+Save data
+```{r}
 #save dataframe into proper library data directory

### 2018-02-28 E1/mixModelE1data.Rmd
looking good

@@
 
-df<-df %>% rename(stream = target, orientation = condition )
-df <- df %>% mutate( stream =ifelse(stream==1, "Left","Right") )
-#mutate condition to Orientation
-df <- df %>% mutate( orientation =ifelse(orientation==1, "Canonical","Inverted") )
+```
 
+Sanity-check some data by graphing.
+
+Investigate whether the pathological identicalness of inverted/canonical is already at this point
+```{r}
+dg<- df %>% filter(subject < "AC")
+
+library(ggplot2)
+g=ggplot(dg, aes(x=SPE)) + facet_grid(subject+orientation~stream) #,  scales="free_y")
+g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
+g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
+show(g)
 ```
@@
 Fit mixture model with R (if fitting not previously done and cached). 
-```{r, echo=FALSE, cache=TRUE}
-#Setting cache = TRUE in hopes won't have to recalculate estimates "if cached results exist and this code chunk has not been changed since last run" 
+```{r, echo=FALSE}
+
 library(mixRSVP)
 
-#Check whether already have parameter estimates or instead need to do it
-calculate<-FALSE
-if (!exists("estimates")) { 
-  calculate<-TRUE 
-}
+#Bizarrely, when knitting this often quits after only 6 parameter estimates. Don't know if there's some function fitting error catch problem
 
@@
 
-if (calculate) {
-  estimates<- df %>%   filter(subject=="AA") %>%
+estimates<- df  %>%  #  filter(subject=="AA") %>%
     group_by_(.dots = condtnVariableNames) %>%  #.dots needed when you have a variable containing multiple factor names
     do(  analyzeOneConditionDF(.,numItemsInStream,parameterBounds(),nReplicates=5) )
-  estimates<- estimates %>% rename(efficacy = p1, latency = p2, precision = p3)
 
-}
 head(estimates)
@@
 
-Save the param estimates to a file
+Save the param estimates to a file.
 ```{Save param estimates, echo=FALSE}
 
-write.csv(estimates, file = "Data/backwards2E1_paramEstimates")
+write.csv(estimates, file = "Results/backwards2E1_paramEstimates.csv")
+#saveRDS(estimates, file = "Data/backwards2E1_paramEstimates.rda", compress="bzip2")
 
 ```
-
-Let's inspect histogram plots and fits. Note that in every case, the negative log likelihood is better (smaller) for R, showing that the R code provides a better fit than the MATLAB code.
-
-* yellow = guessing component 
-* light blue = Gaussian component
-* green = sum of the guessing and Gaussian components. In other words, the histogram heights predicted by the model
-* dark blue = continuous Gaussian.
-This helps get a sense of the effect of discretising the Gaussian.
-For instance, it's possible (especially using Pat's method, it seems), for the Gaussian peak to fly high above the bars and still fit the discrete bins (or bin centers, in Pat's method), suggesting an undesirably high estimates of the efficacy (likely accompanied by an undesirably low precision)
-
-```{r, echo=FALSE, message=FALSE, fig.height=360, fig.width=10}
-#want fig.height of 10 per subject
-
-source( file.path(mixModelingPath,"calcCurvesDataframes.R") )
-
-#create R curves
-df<- df # %>%  dplyr::filter(subject <="BB") #dplyr::filter(subject <= "BD" & subject >="AP")
-#dplyr::filter(subject=="AA",orientation=="Canonical")
-source( file.path(mixModelingPath,"histogramPlotting.R") ) #for calcFitDataframes
-source( file.path(mixModelingPath,"theme_apa.R") ) #for calcFitDataframes
-
-#Add R parameter estimates to dataframe
-df<- merge(df,estimates)
-
-curvesR<- df %>% group_by_at(.vars = condtnVariableNames) %>% 
-  do(calcCurvesDataframes(.,minSPE,maxSPE,numItemsInStream))
-
-#Calc numObservations to each condition. This is needed only for scaling the fine-grained Gaussian
-#Calc the number of observations for each condition, because gaussianScaledforData needs to know.
-dfGroups<- df %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
-#add nPerCond back to parameter estimates
-estimates<- merge(estimates,dfGroups)
-grain<-.05
-gaussianFineR<- estimates %>% group_by_at(.vars = condtnVariableNames) %>% do(
-  gaussianScaledFromDataframe(.,minSPE,maxSPE,grain) )
-
-#create MATLAB curves
-#need to rename parameters so that calcCurvesDataframes can find them
-estimates_M<- estimates_MATLAB %>% rename(efficacy=efficacy_M,latency=latency_M,precision=precision_M)
-#create temporary dataframe with data plus MATLAB estimates to generate curves
-dataStrippedOfEstimates<- select(df,-efficacy,-latency,-precision, -val)
-dM<-merge( dataStrippedOfEstimates, estimates_M )
-dM<- dM %>% rename( val = val_M ) #has to be called val so calcFitDataframes uses it rather than using R to calculate its own
-curvesMATLAB<- dM %>% group_by_at(.vars = condtnVariableNames) %>% 
-            do(calcCurvesDataframes(.,minSPE,maxSPE,numItemsInStream))
-
-#Calculate fine-grained Gaussian for MATLAB
-#Calc the number of observations for each condition, because gaussianScaledforData needs to know.
-dNum<- dataStrippedOfEstimates %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
-#add nPerCond to MATLAB estimates, which will then be merged with raw data
-estimates_M<- merge(estimates_M,dNum)
-#merge MATLAB estimates with raw data
-dM<-merge( dataStrippedOfEstimates, estimates_M )
-gaussianFineM<- estimates_M %>% group_by_at(.vars = condtnVariableNames) %>% do(
-  gaussianScaledFromDataframe(.,minSPE,maxSPE,grain) )
-
-#Concatenate R and MATLAB into single dataframe
-curvesR$lang<-"R"; curvesMATLAB$lang<-"MATLAB"
-curves_R_MATLAB <- rbind( data.frame(curvesR), data.frame(curvesMATLAB) )
-#To facilitate plotting,
-#make dataframe containing the raw data twice, one for R plots column, one for MATLAB plots column (data identical)
-dfBoth<-rbind(   df %>% mutate(lang="R"), df %>% mutate(lang="MATLAB")  )
-
-#Concatenate fine-grained gaussian for R and MATLAB into single dataframe, so can plot both with ggplot
-gaussianFineR$lang<-"R"; gaussianFineM$lang<-"MATLAB"; 
-#Get the col names to match by only using a few columns
-gaussianFineR<- gaussianFineR %>% select(-val,-warnings)
-gaussianFineM<- gaussianFineM %>% select(-val_M)
-gaussFine_R_M <- rbind( data.frame(gaussianFineR), data.frame(gaussianFineM) )
-
-#PLOT EVERYTHING
-g=ggplot(dfBoth, aes(x=SPE)) + facet_grid(subject+stream+orientation~lang,  scales="free_y")
-g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
-g<-g +theme_apa() #+theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank())# hide all gridlines.
-#g<-g+ theme(line=element_blank(), panel.border = element_blank())
-sz=.8
-#Plot the underlying Gaussian , not just the discretized Gaussian
-g<-g + geom_line(data=gaussFine_R_M,aes(x=x,y=gaussianFreq),color="darkblue",size=1.2)
-
-g<-g+ geom_point(data=curves_R_MATLAB,aes(x=x,y=combinedFitFreq),color="chartreuse3",size=sz*2.5)
-g<-g+ geom_line(data=curves_R_MATLAB,aes(x=x,y=guessingFreq),color="yellow",size=sz)
-#Discretized Gaussian
-g<-g+ geom_line(data=curves_R_MATLAB,aes(x=x,y=gaussianFreq),color="lightblue",size=sz)
-
-numGroups<- length(table(dfBoth$orientation,dfBoth$subject,dfBoth$stream))
-fontSz = 400/numGroups
-g<-g + geom_text(data=curves_R_MATLAB,aes(x=-9,y=32, label = paste("-logLik==", round(val,1), sep = "")),  parse = TRUE,size=fontSz) +
-  geom_text(data=curves_R_MATLAB,aes(x=-7,y=28, label = paste("plain(e)==", round(efficacy,2), sep = "")),  parse = TRUE,size=fontSz) +
-  geom_text(data=curves_R_MATLAB,aes(x=-7,y=25, label = paste("mu==", round(latency,2), sep = "")),  parse = TRUE,size=fontSz)+
-  geom_text(data=curves_R_MATLAB,aes(x=-7,y=22, label = paste("sigma==", round(precision,2), sep = "")),  parse = TRUE,size=fontSz)
-show(g)
-```
+Done.

### 2018-03-01 E1/doInferentialStats.Rmd
long to wide was a nightmare

---
title: "InspectHistogramsE1"
author: "Alex Holcombe"
date: "2/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in parameter estimates

```{r}

estimates<- read.csv(file="Results/backwards2E1_paramEstimates.csv", header=TRUE, sep=",")
estimates$X<- NULL #Not sure where X comes from but it's just a row number
```

Exclude Ss who failed likelihood ratio test in all four cases.
Create new column nPassed for in how many conditions they passed (out of 4).

```{r}
library(dplyr)

estimates<- estimates %>% mutate(mixModelFitsBetter = pLRtest<.05)

condtnVariableNames <- c("subject","stream", "orientation")

e<- estimates %>%  group_by(subject) %>%
  summarise(nPassed = sum(mixModelFitsBetter)) 
estimates<- merge(e,estimates)

estimates<- estimates %>% mutate( exclude = (nPassed==0)  )
```

Assess the size of the effect to inform Bayesian test of next experiment (probably use raw units)

```{r, echo=FALSE, message=FALSE}

gg<-ggplot(estimates, aes(x=stream, y=efficacy)) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
show(gg)

library(tidyr)

data.table::dcast(df, orientation~stream,value.var="efficacy")

data.table::dcast(df, stream~orientation,value.var="efficacy")

temp<-estimates[,1:5]
ok<-data.table::dcast(temp, orientation+subject~stream,value.var="efficacy")
#Good except how do you preserve nPassed? Well, nPassed is the same for every factor combintation
data.table::dcast(temp, orientation+subject+nPassed~stream,value.var="efficacy")
#Good so now try with two DVs. 
temp<-estimates[,1:6]
data.table::dcast(temp, orientation+subject+nPassed~stream,
                  value.var=c("efficacy","latency")) #doesn't work
data.table::dcast(temp, orientation+subject+nPassed~stream,value.var="latency")

#The below works but drops many variables, so think about merging later
eff<- data.table::dcast(temp, orientation+subject+nPassed~stream,value.var="efficacy")
lat<-data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="latency") 
pre<-data.table::dcast(temp, orientation+subject+nPassed~stream,value.var="precision")

#So now scale up to full dataframe

```

```{r}

temp<-estimates[1:10,1:5]
temp$nPassed<-NULL
data.table::dcast(temp, subject+orientation~stream,value.var="efficacy")

temp %>%
  gather(key, value, -subject, -orientation, -nPassed, -stream) %>%
  unite(col, key, stream) %>%
  spread(col, value)

data.table::dcast(temp, subject+orientation+nPassed~stream,value.var="efficacy")

foo = data.frame(subject=c("AA","AA","AA","AA","AB","AB","AB","AB"),stream=c("Left","Left","Right","Right","Left","Left","Right","Right"),orientation=c("Canonical","Inverted","Canonical","Inverted","Canonical","Inverted","Canonical","Inverted"),efficacy=c(0.8,0.4,0.5,0.7,0.84,.37,.36,.69))
data.table::dcast(foo, subject+orientation~stream,value.var="efficacy")
foo.short<- foo[,1:4] #,c(2,3,4)]
f<-foo[1:4,2:4]
data.table::dcast(f, orientation~stream,value.var="efficacy")

tempp<-estimates[,4:6]
temppp<- tempp[,c(1,3)]

library(reshape2)

#data.table functions change things in place
require(data.table)
setDT(temp)
data.table::dcast(temp, subject+orientation+X+nPassed~stream,value.var="efficacy")

dcast(foo,group+num_users~times,value.var=c("action_rate","action_rate_c95"))

dcast(temp, subject+nPassed+X+orientation ~stream, value.var=efficacy)
?ddcast(tempp, 
temppp$id<- 1:nrow(temppp)
dcast(temppp, id~stream)
dcast(temppp, efficacy~stream+id)

wide <-  estimates %>%
  tidyr::spread(stream,efficacy)

df$

leftBias<- estimates %>% group_by(subject) %>%

deficitCanonical <- estimates %>% filter( stream=="Left" ) -

```

Do t tests.

Do Bayesian t-test. That is, take the likelihood of the left bias difference observed under the prior and divide it by the likelihood of the left bias difference observed under the null.

http://jeffrouder.blogspot.com.au/2016/01/what-priors-should-i-use-part-i.html
For the null model, the bias is just zero (delta function).

For our model, support is zero to infinity.

```{r, echo=FALSE, message=FALSE}
#want fig.height of 10 per subject

condtnVariableNames <- c("subject","stream", "orientation")

#Add R parameter estimates to dataframe
df<- merge(df,estimates)

```

### 2018-03-01 E1/doInferentialStats.Rmd
good

@@
 ```{r, echo=FALSE, message=FALSE}
-
+require(ggplot2)
 gg<-ggplot(estimates, aes(x=stream, y=efficacy)) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
@@
 
-library(tidyr)
-
-data.table::dcast(df, orientation~stream,value.var="efficacy")
-
-data.table::dcast(df, stream~orientation,value.var="efficacy")
-
-temp<-estimates[,1:5]
-ok<-data.table::dcast(temp, orientation+subject~stream,value.var="efficacy")
-#Good except how do you preserve nPassed? Well, nPassed is the same for every factor combintation
-data.table::dcast(temp, orientation+subject+nPassed~stream,value.var="efficacy")
-#Good so now try with two DVs. 
-temp<-estimates[,1:6]
-data.table::dcast(temp, orientation+subject+nPassed~stream,
-                  value.var=c("efficacy","latency")) #doesn't work
-data.table::dcast(temp, orientation+subject+nPassed~stream,value.var="latency")
+library(data.table)
 
 #The below works but drops many variables, so think about merging later
-eff<- data.table::dcast(temp, orientation+subject+nPassed~stream,value.var="efficacy")
+eff<- data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="efficacy")
+eff$dv<-"efficacy"
 lat<-data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="latency") 
-pre<-data.table::dcast(temp, orientation+subject+nPassed~stream,value.var="precision")
+lat$dv<-"latency"
+pre<-data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="precision") 
+pre$dv<-"precision"
+all<-rbind(eff,lat,pre)
+#I tried to do the above with a single command but makes an error even though StackOverflow example works
+#https://stackoverflow.com/questions/37332076/from-long-to-wide-data-with-multiple-columns/37332345
+# data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var=c("efficacy","latency"))
+head(eff)
+```
 
-#So now scale up to full dataframe
+Calculate bias
+```{r , fig.height=40, fig.width=5}
+all$leftBias<- all$Left - all$Right
+hh<-ggplot(all, aes(x=orientation, y=leftBias)) + facet_grid(dv~.) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
+hh<-hh+stat_summary(fun.data="mean_cl_boot",geom="errorbar",width=.3,conf.int=.95,
+                  width=5,size=1) 
+show(hh)
+```
+
+Do conventional stats.
+```{r}
+
+require(ez)
+
+aa <- ezANOVA(data=filter(all,dv=="efficacy"), dv=leftBias, within=orientation, wid=subject)
+cat("F=", aa$ANOVA$F, " p=", aa$ANOVA$p, "\n", sep="")
 
 ```
-
-```{r}
-
-temp<-estimates[1:10,1:5]
-temp$nPassed<-NULL
-data.table::dcast(temp, subject+orientation~stream,value.var="efficacy")
-
-temp %>%
-  gather(key, value, -subject, -orientation, -nPassed, -stream) %>%
-  unite(col, key, stream) %>%
-  spread(col, value)
-
-data.table::dcast(temp, subject+orientation+nPassed~stream,value.var="efficacy")
-
-foo = data.frame(subject=c("AA","AA","AA","AA","AB","AB","AB","AB"),stream=c("Left","Left","Right","Right","Left","Left","Right","Right"),orientation=c("Canonical","Inverted","Canonical","Inverted","Canonical","Inverted","Canonical","Inverted"),efficacy=c(0.8,0.4,0.5,0.7,0.84,.37,.36,.69))
-data.table::dcast(foo, subject+orientation~stream,value.var="efficacy")
-foo.short<- foo[,1:4] #,c(2,3,4)]
-f<-foo[1:4,2:4]
-data.table::dcast(f, orientation~stream,value.var="efficacy")
-
-tempp<-estimates[,4:6]
-temppp<- tempp[,c(1,3)]
-
-library(reshape2)
-
-#data.table functions change things in place
-require(data.table)
-setDT(temp)
-data.table::dcast(temp, subject+orientation+X+nPassed~stream,value.var="efficacy")
-
-dcast(foo,group+num_users~times,value.var=c("action_rate","action_rate_c95"))
-
-dcast(temp, subject+nPassed+X+orientation ~stream, value.var=efficacy)
-?ddcast(tempp, 
-temppp$id<- 1:nrow(temppp)
-dcast(temppp, id~stream)
-dcast(temppp, efficacy~stream+id)
-
-wide <-  estimates %>%
-  tidyr::spread(stream,efficacy)
-
-df$
-
-leftBias<- estimates %>% group_by(subject) %>%
-
-deficitCanonical <- estimates %>% filter( stream=="Left" ) -
-
-```
-
-Do t tests.
 
@@
 For our model, support is zero to infinity.
-
-```{r, echo=FALSE, message=FALSE}
-#want fig.height of 10 per subject
-
-condtnVariableNames <- c("subject","stream", "orientation")
-
-#Add R parameter estimates to dataframe
-df<- merge(df,estimates)
-
-```

### 2018-03-01 E1/doInferentialStats.Rmd
good

@@
 ---
-title: "InspectHistogramsE1"
+title: "InferentialStats"
 author: "Alex Holcombe"
@@
 Calculate bias
-```{r , fig.height=40, fig.width=5}
+```{r , fig.height=20, fig.width=5}
 all$leftBias<- all$Left - all$Right
-hh<-ggplot(all, aes(x=orientation, y=leftBias)) + facet_grid(dv~.) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
+hh<-ggplot(all, aes(x=orientation, y=leftBias)) + facet_grid(.~dv) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
 hh<-hh+stat_summary(fun.data="mean_cl_boot",geom="errorbar",width=.3,conf.int=.95,
@@
 ```
+
+Print effect size
+```{r}
+leftBiasCanonical <- mean(   filter(all,dv=="efficacy" & orientation=="Canonical")$leftBias  )
+leftBiasInverted<- mean(   filter(all,dv=="efficacy" & orientation=="Inverted")$leftBias  )
+leftBiasDiff<- leftBiasCanonical - leftBiasInverted
+```
+
+The leftBiasDiff for this experiment is `r leftBiasDiff`, very healthy.
+
+```{r, echo=FALSE}
+P1E1canonical=.23
+P1E1reversed= .062
+P1E2canonical=.218 #sd=.034
+P1E2inverted=.017 #sd=.049
+```
+
+What was it for the first paper? For E1,`r P1E1canonical`  Canonical and 
+`r P1E1reversed` Mirror-reversed. Difference is `r P1E1canonical-P1E1reversed`
+
+For E2, very similar at `r P1E2canonical - P1E2inverted`.
 
@@
 For our model, support is zero to infinity. 
+```{r}
+
+#Specify alternative hypothesis (prior)
+
+#Specify Alternative (up to constant of proportionality)
+lo=0 #lower bound of support
+hi=Inf #upper bound of support
+altDens=function(delta) {
+  y= dnorm(delta,.5,.3)
+  y=y*as.integer(delta>lo)*as.integer(delta<hi) #multiply by zero if outside of support
+}
+
+#Normalize alternative density in case user does not, 
+K=1/integrate(altDens,lower=lo,upper=hi)$value
+f=function(delta) K*altDens(delta)
+
+delta=seq(-2,3,.01)
+
+#Plot Alternative as a density and Null as a point arrow
+maxAlt=max(f(delta))
+plot(delta,f(delta),typ='n',xlab="Effect Size Parameter Delta",ylab="Density",ylim=c(0,1.4*maxAlt),main="Models")
+arrows(0,0,0,1.3*maxAlt,col='darkblue',lwd=2)
+lines(delta,f(delta),col='darkgreen',lwd=2)
+legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
+
+```
+
+Do the correlogram and correlation

### 2018-03-01 E1/doInferentialStats.Rmd
great

@@
 P1E2inverted=.017 #sd=.049
+P1E2vertCanon=.188 #sd=.043
+P1E2vertUpright= -.07 #sd=.049
+P1E1leftBiasDiff<- P1E1canonical-P1E1reversed
+P1E2leftBiasDiff<- P1E2canonical - P1E2inverted
+P1E2vertBiasDiff<- P1E2vertCanon - P1E2vertUpright
 ```
@@
 `r P1E1reversed` Mirror-reversed. Difference is `r P1E1canonical-P1E1reversed`
-
-For E2, very similar at `r P1E2canonical - P1E2inverted`.
+. For E2, bias difference very similar at `r P1E2canonical - P1E2inverted`.
+For E2 vertical arrangement, `r P1E2vertCanon - P1E2vertUpright`
 
@@
 altDens=function(delta) {
-  y= dnorm(delta,.5,.3)
+  y= dnorm(delta,.3,.1)
   y=y*as.integer(delta>lo)*as.integer(delta<hi) #multiply by zero if outside of support
@@
 
-delta=seq(-2,3,.01)
+delta=seq(-1,1.2,.01)
 
@@
 maxAlt=max(f(delta))
-plot(delta,f(delta),typ='n',xlab="Effect Size Parameter Delta",ylab="Density",ylim=c(0,1.4*maxAlt),main="Models")
+plot(delta,f(delta),typ='n',xlab="Left bias difference",ylab="Density",ylim=c(0,1.4*maxAlt),main="Models")
 arrows(0,0,0,1.3*maxAlt,col='darkblue',lwd=2)
-lines(delta,f(delta),col='darkgreen',lwd=2)
-legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
+lines(delta,f(delta),col='green',lwd=2)
+legend("topright",legend=c("Null","Alternative","Previous findings"),col=c('darkblue','green','black'),lwd=2)
+
+points(P1E1leftBiasDiff,.1,pch=19)
+points(P1E2leftBiasDiff,.1,pch=19)
+points(P1E2vertBiasDiff,.1,pch=22)
+points(leftBiasDiff,.1,pch=23)
 

### 2018-03-06 E1/doInferentialStats.Rmd
right

@@
 ```{r}
-leftBiasCanonical <- mean(   filter(all,dv=="efficacy" & orientation=="Canonical")$leftBias  )
-leftBiasInverted<- mean(   filter(all,dv=="efficacy" & orientation=="Inverted")$leftBias  )
+leftBiasesCanonical<- filter(all,dv=="efficacy" & orientation=="Canonical")$leftBias
+leftBiasCanonical <- mean( leftBiasesCanonical )
+leftBiasesInverted<- filter(all,dv=="efficacy" & orientation=="Inverted")$leftBias
+leftBiasInverted<- mean(   leftBiasesInverted  )
 leftBiasDiff<- leftBiasCanonical - leftBiasInverted
+leftBias_sd <- sd(   leftBiasesCanonical - leftBiasesInverted )
 ```
 
-The leftBiasDiff for this experiment is `r leftBiasDiff`, very healthy.
+The leftBiasDiff for this experiment is `r leftBiasDiff` or in Cohen's d `leftBiasDiff / leftBias_sd`, very healthy.
 
 ```{r, echo=FALSE}
-P1E1canonical=.23
-P1E1reversed= .062
-P1E2canonical=.218 #sd=.034
-P1E2inverted=.017 #sd=.049
-P1E2vertCanon=.188 #sd=.043
-P1E2vertUpright= -.07 #sd=.049
+E1n<- 16
+#Raw effect size plus standard error and sd
+P1E1canonical=.23;  P1E1c_se = .058; P1E1c_sd = P1E1c_se * sqrt(E1n)
+P1E1reversed= .062;  P1E1r_se = .07; P1E1r_sd = P1E1r_se * sqrt(E1n)
+#To calculate sd of the bias diff, consider that the variances sum,
+#therefore first sum the variances then convert back to standard deviation
+#Although only true for *independent* random variables and these unlikely to be totally independent
+P1E1diff_sdCalculated = sqrt( (P1E1c_sd^2 + P1E1r_sd^2) )  #.257
+
+E2n<- 24
+P1E2canonical=.218; P1E2c_se =.034; P1E2c_sd= P1E2c_se * sqrt(E2n)
+P1E2inverted=.017; P1E2i_se =.049; P1E2i_sd = P1E2i_se * sqrt(E2n)
+P1E2diff_sdCalculated = sqrt( (P1E2c_sd^2 + P1E2i_sd^2) ) #.2066
+
+P1E2vertUpright=.188; P1E2vu_se=.043; P1E2vu_sd = P1E2vu_se * sqrt(E2n)
+P1E2vertInverted= -.07; P1E2vi_se=.049; P1E2vi_sd = P1E2vi_se * sqrt(E2n)
+P1E2vdiff_sdCalculated = sqrt( (P1E2vi_sd^2 + P1E2vu_sd^2) )
+
 P1E1leftBiasDiff<- P1E1canonical-P1E1reversed
 P1E2leftBiasDiff<- P1E2canonical - P1E2inverted
-P1E2vertBiasDiff<- P1E2vertCanon - P1E2vertUpright
+P1E2vertBiasDiff<- P1E2vertUpright - P1E2vertInverted
+#reported as  .258 +/- .047 in the paper
+
+#The diff standard errors are also available from the paper directly. But still need to 
+#convert to SD to get Cohen's d.
+
+P1E1diff_se<-.07; P1E1diff_sd<- P1E1diff_se*sqrt(E2n) 
+P1E2diff_se<-.038; P1E2diff_sd<- P1E2diff_se*sqrt(E2n) 
+P1E2vdiff_se<- .047; P1E2vdiff_sd<- P1E2vdiff_se*sqrt(E2n)
+
 ```
@@
 `r P1E1reversed` Mirror-reversed. Difference is `r P1E1canonical-P1E1reversed`
-. For E2, bias difference very similar at `r P1E2canonical - P1E2inverted`.
-For E2 vertical arrangement, `r P1E2vertCanon - P1E2vertUpright`
+. For E2, bias difference very similar at `r P1E2leftBiasDiff`.
+For E2 vertical arrangement, `r P1E2vertBiasDiff`
 
@@
 http://jeffrouder.blogspot.com.au/2016/01/what-priors-should-i-use-part-i.html
-For the null model, the bias is just zero (delta function).
+For the null model, the bias is just zero (delta function).For our model, support is zero to infinity.
 
-For our model, support is zero to infinity. 
+Look at it first in raw units.
 ```{r}
@@
 altDens=function(delta) {
-  y= dnorm(delta,.3,.1)
+  y= dnorm(delta,.3,.13)
   y=y*as.integer(delta>lo)*as.integer(delta<hi) #multiply by zero if outside of support
@@
 
-delta=seq(-1,1.2,.01)
+delta=seq(-.2,1.2,.01)
 
@@
 
+Probably better to do it in standardized units (Cohen's d) because otherwise need to create a model of the variance separately (for the data model) so that can calculate the probability of the data given e.g.
+the null hypothesis.
+
+```{r}
+
+P2E1_cohensD <- leftBiasDiff / leftBias_sd
+cohensDs<-c(P1E1leftBiasDiff / P1E1diff_sd,
+            P1E2leftBiasDiff / P1E2diff_sd,
+            P1E2vertBiasDiff / P1E2vdiff_sd,
+            P2E1_cohensD ) 
+labels<-c("P1E1","P1E2","P1E2v","P2E1")
+meanCohensDs<- mean(cohensDs)
+
+#Specify Alternative (up to constant of proportionality)
+lo=-.6 #lower bound of support
+hi=Inf #upper bound of support
+
+altHypothesisPrior=function(delta) {
+  mean= meanCohensDs
+  sigma= sd(cohensDs)
+  y= dnorm(delta,mean,sigma)
+  y=y*as.integer(delta>lo)*as.integer(delta<hi) #multiply by zero if outside of support
+}
+#Normalize alternative density in case is not already
+K=1/integrate(altHypothesisPrior,lower=lo,upper=hi)$value
+f=function(delta) K*altHypothesisPrior(delta)
+
+dDomain=seq(-.3,2,.01)
+
+#Plot Alternative as a density and Null as a point arrow
+maxAlt=max(f(delta))
+plot(dDomain,f(dDomain),typ='n',xlab="Left/top bias difference (Cohen's d)",ylab="Density",ylim=c(0,1.4*maxAlt),main="Models")
+arrows(0,0,0,1.3*maxAlt,col='darkblue',lwd=2)
+lines(dDomain,f(dDomain),col='green',lwd=2)
+legend("topright",legend=c("Null","Alternative","Previous findings"),col=c('darkblue','green','black'),lwd=2)
+
+points(cohensDs, rep(.1,4),pch=19) #plot previous effect sizes
+text(cohensDs, c(.25,.35,.25,.25), labels,cex=.75)
+
+```
+
+Assume left bias difference found for any particular subject reflect independent normal draws with mean μ and variance σ2.
+Set variance across Ss based on previous experiments.
+Take something like the mean variance of those previous experiments.
+Then use that to generate the probability of any particular result.
+
+Let's say we have a sample size of 50. 
+```{r}
+
+nullPredF=function(obs,N) dt(sqrt(N)*obs,N-1)
+
+#compute the probability of the data (predicted density) for any observed effect size or for all of them.
+The following code does it for a reasonable range of effect sizes for a sample size of 30.
+You can change N, the sample size, as needed.
+obsEffectSizes=seq(-2,3,.01)
+N=50
+nullPred=nullPredF(obsEffectSizes,N)
+
+#Getting the probabilites of the data for the alternative is a bit harder.
+For each nonzero effect size parameter, the distribution of the observed effect follows a noncentral t distribution.
+Hence, to obtain predictions across all nonzero effect sizeS, we need to integrate the alternative model against the noncentral t distribution.
+Here is the code with a simple loop:
+
+altPredIntegrand=function(delta,obs,N) { 
+  dt(sqrt(N)*obs,N-1,sqrt(N)*delta)*f(delta)
+}
+
+altPredF=function(obsEffectSizes,N) {
+  suppressWarnings( integrate(altPredIntegrand,lower=lo,upper=hi,obs=obsEffectSizes,N=N)$value
+  )
+}
+
+#For each effect size, compute density
+I=length(obsEffectSizes)
+altPred=1:I
+for (i in 1:I) {
+  altPred[i] = altPredF(obsEffectSizes[i],N)
+}
+
+#Now we can plot the predictions for all observed effect sizes:
+top=max(altPred,nullPred)
+
+plot(type='l',obsEffectSizes,nullPred,ylim=c(0,top),xlab="Observed Effect Size",ylab="Density",main="P(data|H)",col='darkblue',lwd=2)
+lines(obsEffectSizes,altPred,col='darkgreen',lwd=2)
+legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
+text(-1.8,.35,"n=50")
+```
+
+BTW get this warning: https://stackoverflow.com/questions/39183938/rs-t-distribution-says-full-precision-may-not-have-been-achieved
+
+```{r}  
+smaller=.3333
+my.es= meanCohensDs * smaller
+
+```
+
+Assume effect size is only `r smaller` as big as original `r meanCohensDs`, yielding `r smaller`
+ . So, both set prior to half as big but with wider variance and show likelihood functions for whole range but highlight that effect size, for a few different sample sizes.
+```{r alternativePrior}
+
+altPrior=function(delta) {
+  mean= my.es
+  sigma= sd(cohensDs)
+  y= dnorm(delta,mean,sigma)
+  y=y*as.integer(delta>lo)*as.integer(delta<hi) #multiply by zero if outside of support
+}
+
+#Normalize alternative density in case user does not, 
+K=1/integrate(altPrior,lower=lo,upper=hi)$value
+altPriorNormalized=function(delta) K*altPrior(delta)
+
+N=75
+```
+Have set alternative prior as smaller effect.
+
+Now visualize likelihoods for N=`r N`
+```{r }
+
+# to obtain predictions across all nonzero effect sizeS, we need to integrate the alternative model against the noncentral t distribution. Here is the code with a simple loop:
+
+altPredIntegrand=function(delta,obs,N) { 
+  dt(sqrt(N)*obs,N-1,sqrt(N)*delta)*altPriorNormalized(delta)
+}
+
+altPredF=function(obsEffectSizes,N) {
+  suppressWarnings( integrate(altPredIntegrand,lower=lo,upper=hi,obs=obsEffectSizes,N=N)$value
+  )
+}
+
+#For each effect size, compute density
+obsEffectSizes=seq(-1,1.5,.01)
+
+I=length(obsEffectSizes)
+altPred=1:I
+for (i in 1:I) {
+  altPred[i] = altPredF(obsEffectSizes[i],N)
+}
+
+nullPred=nullPredF(obsEffectSizes,N)
+
+#Now we can plot the predictions for all observed effect sizes:
+top=max(altPred,nullPred)
+
+plot(type='l',obsEffectSizes,nullPred,ylim=c(0,top),xlab="Observed Effect Size",ylab="Density",main="P(data|H)",col='darkblue',lwd=2)
+lines(obsEffectSizes,altPred,col='darkgreen',lwd=2)
+legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
+text(-.9,.35,paste0("n=",N))
+#plot example effect size points on curve
+abline(v=my.es,lty=2,lwd=2,col='red')
+valNull=nullPredF(my.es,N)
+valAlt=altPredF(my.es,N)
+points(pch=19,c(my.es,my.es),c(valNull,valAlt),col="red")
+text(my.es+.2,valAlt+.1,paste0("BF=",round(valAlt/valNull,2)),col="red")
+```
+And for n=100
+```{r subjexts100}
+#For each effect size, compute density
+obsEffectSizes=seq(-1,1.5,.01)
+
+I=length(obsEffectSizes)
+altPred=1:I
+N=100
+for (i in 1:I) {
+  altPred[i] = altPredF(obsEffectSizes[i],N)
+}
+
+nullPred=nullPredF(obsEffectSizes,N)
+
+#Now we can plot the predictions for all observed effect sizes:
+top=max(altPred,nullPred)
+
+plot(type='l',obsEffectSizes,nullPred,ylim=c(0,top),xlab="Observed Effect Size",ylab="Density",main="P(data|H)",col='darkblue',lwd=2)
+lines(obsEffectSizes,altPred,col='darkgreen',lwd=2)
+legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
+text(-.9,.35,paste0("n=",N))
+
+#plot example effect size points on curve
+abline(v=my.es,lty=2,lwd=2,col='red')
+valNull=nullPredF(my.es,N)
+valAlt=altPredF(my.es,N)
+points(pch=19,c(my.es,my.es),c(valNull,valAlt),col="red")
+text(my.es+.2,valAlt+.1,paste0("BF=",round(valAlt/valNull,2)),col="red")
+```
+
 Do the correlogram and correlation

### 2018-03-07 E1/doInferentialStats.Rmd
good
moved prereg over to E2 folder

--- 2018-03-01 E1/doInferentialStats.Rmd
@@
 ```{r}
-leftBiasCanonical <- mean(   filter(all,dv=="efficacy" & orientation=="Canonical")$leftBias  )
-leftBiasInverted<- mean(   filter(all,dv=="efficacy" & orientation=="Inverted")$leftBias  )
+leftBiasesCanonical<- filter(all,dv=="efficacy" & orientation=="Canonical")$leftBias
+leftBiasCanonical <- mean( leftBiasesCanonical )
+leftBiasesInverted<- filter(all,dv=="efficacy" & orientation=="Inverted")$leftBias
+leftBiasInverted<- mean(   leftBiasesInverted  )
 leftBiasDiff<- leftBiasCanonical - leftBiasInverted
+leftBias_sd <- sd(   leftBiasesCanonical - leftBiasesInverted )
 ```
 
-The leftBiasDiff for this experiment is `r leftBiasDiff`, very healthy.
+The leftBiasDiff for this experiment is `r leftBiasDiff` or in Cohen's d `leftBiasDiff / leftBias_sd`, very healthy.
 
 ```{r, echo=FALSE}
-P1E1canonical=.23
-P1E1reversed= .062
-P1E2canonical=.218 #sd=.034
-P1E2inverted=.017 #sd=.049
+E1n<- 16
+#Raw effect size plus standard error and sd
+P1E1canonical=.23;  P1E1c_se = .058; P1E1c_sd = P1E1c_se * sqrt(E1n)
+P1E1reversed= .062;  P1E1r_se = .07; P1E1r_sd = P1E1r_se * sqrt(E1n)
+#To calculate sd of the bias diff, consider that the variances sum,
+#therefore first sum the variances then convert back to standard deviation
+#Although only true for *independent* random variables and these unlikely to be totally independent
+P1E1diff_sdCalculated = sqrt( (P1E1c_sd^2 + P1E1r_sd^2) )  #.257
+
+E2n<- 24
+P1E2canonical=.218; P1E2c_se =.034; P1E2c_sd= P1E2c_se * sqrt(E2n)
+P1E2inverted=.017; P1E2i_se =.049; P1E2i_sd = P1E2i_se * sqrt(E2n)
+P1E2diff_sdCalculated = sqrt( (P1E2c_sd^2 + P1E2i_sd^2) ) #.2066
+
+P1E2vertUpright=.188; P1E2vu_se=.043; P1E2vu_sd = P1E2vu_se * sqrt(E2n)
+P1E2vertInverted= -.07; P1E2vi_se=.049; P1E2vi_sd = P1E2vi_se * sqrt(E2n)
+P1E2vdiff_sdCalculated = sqrt( (P1E2vi_sd^2 + P1E2vu_sd^2) )
+
+P1E1leftBiasDiff<- P1E1canonical-P1E1reversed
+P1E2leftBiasDiff<- P1E2canonical - P1E2inverted
+P1E2vertBiasDiff<- P1E2vertUpright - P1E2vertInverted
+#reported as  .258 +/- .047 in the paper
+
+#The diff standard errors are also available from the paper directly. But still need to 
+#convert to SD to get Cohen's d.
+
+P1E1diff_se<-.07; P1E1diff_sd<- P1E1diff_se*sqrt(E2n) 
+P1E2diff_se<-.038; P1E2diff_sd<- P1E2diff_se*sqrt(E2n) 
+P1E2vdiff_se<- .047; P1E2vdiff_sd<- P1E2vdiff_se*sqrt(E2n)
+
 ```
@@
 `r P1E1reversed` Mirror-reversed. Difference is `r P1E1canonical-P1E1reversed`
-
-For E2, very similar at `r P1E2canonical - P1E2inverted`.
+. For E2, bias difference very similar at `r P1E2leftBiasDiff`.
+For E2 vertical arrangement, `r P1E2vertBiasDiff`
 
@@
 
-http://jeffrouder.blogspot.com.au/2016/01/what-priors-should-i-use-part-i.html
-For the null model, the bias is just zero (delta function).
+Use the default Bayes test (but for E2 will use a custom prior)
 
-For our model, support is zero to infinity. 
-```{r}
+```{r, echo=FALSE}
+if(!require(BayesFactor)){install.packages('BayesFactor')}
+library(BayesFactor)
 
-#Specify alternative hypothesis (prior)
+leftBiasDifferences<- leftBiasesCanonical - leftBiasesInverted
 
-#Specify Alternative (up to constant of proportionality)
-lo=0 #lower bound of support
-hi=Inf #upper bound of support
-altDens=function(delta) {
-  y= dnorm(delta,.5,.3)
-  y=y*as.integer(delta>lo)*as.integer(delta<hi) #multiply by zero if outside of support
-}
-
-#Normalize alternative density in case user does not, 
-K=1/integrate(altDens,lower=lo,upper=hi)$value
-f=function(delta) K*altDens(delta)
-
-delta=seq(-2,3,.01)
-
-#Plot Alternative as a density and Null as a point arrow
-maxAlt=max(f(delta))
-plot(delta,f(delta),typ='n',xlab="Effect Size Parameter Delta",ylab="Density",ylim=c(0,1.4*maxAlt),main="Models")
-arrows(0,0,0,1.3*maxAlt,col='darkblue',lwd=2)
-lines(delta,f(delta),col='darkgreen',lwd=2)
-legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
+ttestResults<- ttestBF( leftBiasDifferences )
 
@@
 
+The Bayes factor from a one-sample default Bayes t-test for the leftBiasDifferences is `r ttestResults@bayesFactor$bf`
+
+The full results are 
+```r 
+print(ttestResults)
+```
+
+A custom prior would provide more power, as detailed in bayesianAnalysisForPrereg.Rmd in E2 folder
+
 Do the correlogram and correlation

### 2018-03-07 E2/analysis/bayesianAnalysisForPrereg.Rmd
moved prereg over to E2 folder

---
title: "BayesianAnalysisForPrereg"
author: "Alex Holcombe"
date: "2/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Illustrate the Bayesian analysis to be done, using the results from E1 as a dummy dataset and showing how the prior is informed by the results of our previous experiments.

Read in parameter estimates

```{r}

estimates<- read.csv(file="../../E1/Results/backwards2E1_paramEstimates.csv", header=TRUE, sep=",")
estimates$X<- NULL #Not sure where X came from but it's just a row number
```

Exclude Ss who failed likelihood ratio test in all four cases.
Create new column nPassed for in how many conditions they passed (out of 4).

```{r}
library(dplyr)

estimates<- estimates %>% mutate(mixModelFitsBetter = pLRtest<.05)

condtnVariableNames <- c("subject","stream", "orientation")

e<- estimates %>%  group_by(subject) %>%
  summarise(nPassed = sum(mixModelFitsBetter)) 
estimates<- merge(e,estimates)

estimates<- estimates %>% mutate( exclude = (nPassed==0)  )
```

Assess the size of the effect to inform Bayesian test of next experiment (probably use raw units)

```{r, echo=FALSE, message=FALSE}
require(ggplot2)
gg<-ggplot(estimates, aes(x=stream, y=efficacy)) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
show(gg)

library(data.table)

#The below works but drops many variables, so think about merging later
eff<- data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="efficacy")
eff$dv<-"efficacy"
lat<-data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="latency") 
lat$dv<-"latency"
pre<-data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="precision") 
pre$dv<-"precision"
all<-rbind(eff,lat,pre)
#I tried to do the above with a single command but makes an error even though StackOverflow example works
#https://stackoverflow.com/questions/37332076/from-long-to-wide-data-with-multiple-columns/37332345
# data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var=c("efficacy","latency"))
head(eff)
```

Calculate bias
```{r , fig.height=20, fig.width=5}
all$leftBias<- all$Left - all$Right
hh<-ggplot(all, aes(x=orientation, y=leftBias)) + facet_grid(.~dv) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
hh<-hh+stat_summary(fun.data="mean_cl_boot",geom="errorbar",width=.3,conf.int=.95,
                  width=5,size=1) 
show(hh)
```

Print effect size
```{r}
leftBiasesCanonical<- filter(all,dv=="efficacy" & orientation=="Canonical")$leftBias
leftBiasCanonical <- mean( leftBiasesCanonical )
leftBiasesInverted<- filter(all,dv=="efficacy" & orientation=="Inverted")$leftBias
leftBiasInverted<- mean(   leftBiasesInverted  )
leftBiasDiff<- leftBiasCanonical - leftBiasInverted
leftBias_sd <- sd(   leftBiasesCanonical - leftBiasesInverted )
```

The leftBiasDiff for this experiment is `r leftBiasDiff` or in Cohen's d `leftBiasDiff / leftBias_sd`, very healthy.

```{r, echo=FALSE}
E1n<- 16
#Raw effect size plus standard error and sd
P1E1canonical=.23;  P1E1c_se = .058; P1E1c_sd = P1E1c_se * sqrt(E1n)
P1E1reversed= .062;  P1E1r_se = .07; P1E1r_sd = P1E1r_se * sqrt(E1n)
#To calculate sd of the bias diff, consider that the variances sum,
#therefore first sum the variances then convert back to standard deviation
#Although only true for *independent* random variables and these unlikely to be totally independent
P1E1diff_sdCalculated = sqrt( (P1E1c_sd^2 + P1E1r_sd^2) )  #.257

E2n<- 24
P1E2canonical=.218; P1E2c_se =.034; P1E2c_sd= P1E2c_se * sqrt(E2n)
P1E2inverted=.017; P1E2i_se =.049; P1E2i_sd = P1E2i_se * sqrt(E2n)
P1E2diff_sdCalculated = sqrt( (P1E2c_sd^2 + P1E2i_sd^2) ) #.2066

P1E2vertUpright=.188; P1E2vu_se=.043; P1E2vu_sd = P1E2vu_se * sqrt(E2n)
P1E2vertInverted= -.07; P1E2vi_se=.049; P1E2vi_sd = P1E2vi_se * sqrt(E2n)
P1E2vdiff_sdCalculated = sqrt( (P1E2vi_sd^2 + P1E2vu_sd^2) )

P1E1leftBiasDiff<- P1E1canonical-P1E1reversed
P1E2leftBiasDiff<- P1E2canonical - P1E2inverted
P1E2vertBiasDiff<- P1E2vertUpright - P1E2vertInverted
#reported as  .258 +/- .047 in the paper

#The diff standard errors are also available from the paper directly. But still need to 
#convert to SD to get Cohen's d.

P1E1diff_se<-.07; P1E1diff_sd<- P1E1diff_se*sqrt(E2n) 
P1E2diff_se<-.038; P1E2diff_sd<- P1E2diff_se*sqrt(E2n) 
P1E2vdiff_se<- .047; P1E2vdiff_sd<- P1E2vdiff_se*sqrt(E2n)

```

What was it for the first paper? For E1,`r P1E1canonical`  Canonical and 
`r P1E1reversed` Mirror-reversed. Difference is `r P1E1canonical-P1E1reversed`
. For E2, bias difference very similar at `r P1E2leftBiasDiff`.
For E2 vertical arrangement, `r P1E2vertBiasDiff`

Do conventional stats.
```{r}

require(ez)

aa <- ezANOVA(data=filter(all,dv=="efficacy"), dv=leftBias, within=orientation, wid=subject)
cat("F=", aa$ANOVA$F, " p=", aa$ANOVA$p, "\n", sep="")

```

Do Bayesian t-test. That is, take the likelihood of the left bias difference observed under the prior and divide it by the likelihood of the left bias difference observed under the null.

http://jeffrouder.blogspot.com.au/2016/01/what-priors-should-i-use-part-i.html
For the null model, the bias is just zero (delta function).For our model, support is zero to infinity.

Look at it first in raw units.
```{r}

#Specify alternative hypothesis (prior)

#Specify Alternative (up to constant of proportionality)
lo=0 #lower bound of support
hi=Inf #upper bound of support
altDens=function(delta) {
  y= dnorm(delta,.3,.13)
  y=y*as.integer(delta>lo)*as.integer(delta<hi) #multiply by zero if outside of support
}

#Normalize alternative density in case user does not, 
K=1/integrate(altDens,lower=lo,upper=hi)$value
f=function(delta) K*altDens(delta)

delta=seq(-.2,1.2,.01)

#Plot Alternative as a density and Null as a point arrow
maxAlt=max(f(delta))
plot(delta,f(delta),typ='n',xlab="Left bias difference",ylab="Density",ylim=c(0,1.4*maxAlt),main="Models")
arrows(0,0,0,1.3*maxAlt,col='darkblue',lwd=2)
lines(delta,f(delta),col='green',lwd=2)
legend("topright",legend=c("Null","Alternative","Previous findings"),col=c('darkblue','green','black'),lwd=2)

points(P1E1leftBiasDiff,.1,pch=19)
points(P1E2leftBiasDiff,.1,pch=19)
points(P1E2vertBiasDiff,.1,pch=22)
points(leftBiasDiff,.1,pch=23)

```

Probably better to do it in standardized units (Cohen's d) because otherwise need to create a model of the variance separately (for the data model) so that can calculate the probability of the data given e.g. the null hypothesis.

```{r}

P2E1_cohensD <- leftBiasDiff / leftBias_sd
cohensDs<-c(P1E1leftBiasDiff / P1E1diff_sd,
            P1E2leftBiasDiff / P1E2diff_sd,
            P1E2vertBiasDiff / P1E2vdiff_sd,
            P2E1_cohensD ) 
labels<-c("P1E1","P1E2","P1E2v","P2E1")
meanCohensDs<- mean(cohensDs)

```

Our informed prior will be a Gaussian distribution with mean our mean observed effect, `r meanCohensDs` and sigma `r sd(cohensDs)`.

```{r}
#Specify Alternative (up to constant of proportionality)
lo=0 #lower bound of support
hi=Inf #upper bound of support

altHypothesisPrior=function(delta) {
  mean= meanCohensDs
  sigma= sd(cohensDs)
  y= dnorm(delta,mean,sigma)
  y=y*as.integer(delta>lo)*as.integer(delta<hi) #multiply by zero if outside of support
}
#Normalize alternative density in case is not already
K=1/integrate(altHypothesisPrior,lower=lo,upper=hi)$value
f=function(delta) K*altHypothesisPrior(delta)

dDomain=seq(-.3,2,.01)

#Plot Alternative as a density and Null as a point arrow
maxAlt=max(f(delta))
plot(dDomain,f(dDomain),typ='n',xlab="Left/top bias difference (Cohen's d)",ylab="Density",ylim=c(0,1.4*maxAlt),main="Models")
arrows(0,0,0,1.3*maxAlt,col='darkblue',lwd=2)
lines(dDomain,f(dDomain),col='green',lwd=2)
legend("topright",legend=c("Null","Alternative","Previous findings"),col=c('darkblue','green','black'),lwd=2)

points(cohensDs, rep(.1,4),pch=19) #plot previous effect sizes
text(cohensDs, c(.25,.35,.25,.25), labels,cex=.75)

```

Assume left bias difference found for any particular subject reflect independent normal draws with mean μ and variance σ2. Set variance across Ss based on previous experiments. Take something like the mean variance of those previous experiments. Then use that to generate the probability of any particular result.

Let's say we have a sample size of 50. 
```{r}

nullPredF=function(obs,N) dt(sqrt(N)*obs,N-1)

#compute the probability of the data (predicted density) for any observed effect size or for all of them. The following code does it for a reasonable range of effect sizes for a sample size of 30. You can change N, the sample size, as needed.
obsEffectSizes=seq(-2,3,.01)
N=50
nullPred=nullPredF(obsEffectSizes,N)

#Getting the probabilites of the data for the alternative is a bit harder. For each nonzero effect size parameter, the distribution of the observed effect follows a noncentral t distribution. Hence, to obtain predictions across all nonzero effect sizeS, we need to integrate the alternative model against the noncentral t distribution. Here is the code with a simple loop:

altPredIntegrand=function(delta,obs,N) { 
  dt(sqrt(N)*obs,N-1,sqrt(N)*delta)*f(delta)
}

altPredF=function(obsEffectSizes,N) {
  suppressWarnings( integrate(altPredIntegrand,lower=lo,upper=hi,obs=obsEffectSizes,N=N)$value
  )
}

#For each effect size, compute density
I=length(obsEffectSizes)
altPred=1:I
for (i in 1:I) {
  altPred[i] = altPredF(obsEffectSizes[i],N)
}

#Now we can plot the predictions for all observed effect sizes:
top=max(altPred,nullPred)

plot(type='l',obsEffectSizes,nullPred,ylim=c(0,top),xlab="Observed Effect Size",ylab="Density",main="P(data|H)",col='darkblue',lwd=2)
lines(obsEffectSizes,altPred,col='darkgreen',lwd=2)
legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
text(-1.8,.35,"n=50")
```

BTW get this warning: https://stackoverflow.com/questions/39183938/rs-t-distribution-says-full-precision-may-not-have-been-achieved

```{r}  
smaller=.3333
my.es= meanCohensDs * smaller

```

Assume effect size is only `r smaller` as big as original `r meanCohensDs`, yielding `r smaller`
 . So, both set prior to half as big but with wider variance and show likelihood functions for whole range but highlight that effect size, for a few different sample sizes.
```{r alternativePrior}

altPrior=function(delta) {
  mean= my.es
  sigma= sd(cohensDs)
  y= dnorm(delta,mean,sigma)
  y=y*as.integer(delta>lo)*as.integer(delta<hi) #multiply by zero if outside of support
}

#Normalize alternative density in case user does not, 
K=1/integrate(altPrior,lower=lo,upper=hi)$value
altPriorNormalized=function(delta) K*altPrior(delta)

N=75
```
Have set alternative prior as smaller effect.

Now visualize likelihoods for N=`r N`
```{r }

# to obtain predictions across all nonzero effect sizeS, we need to integrate the alternative model against the noncentral t distribution. Here is the code with a simple loop:

altPredIntegrand=function(delta,obs,N) { 
  dt(sqrt(N)*obs,N-1,sqrt(N)*delta)*altPriorNormalized(delta)
}

altPredF=function(obsEffectSizes,N) {
  suppressWarnings( integrate(altPredIntegrand,lower=lo,upper=hi,obs=obsEffectSizes,N=N)$value
  )
}

#For each effect size, compute density
obsEffectSizes=seq(-1,1.5,.01)

I=length(obsEffectSizes)
altPred=1:I
for (i in 1:I) {
  altPred[i] = altPredF(obsEffectSizes[i],N)
}

nullPred=nullPredF(obsEffectSizes,N)

#Now we can plot the predictions for all observed effect sizes:
top=max(altPred,nullPred)

plot(type='l',obsEffectSizes,nullPred,ylim=c(0,top),xlab="Observed Effect Size",ylab="Density",main="P(data|H)",col='darkblue',lwd=2)
lines(obsEffectSizes,altPred,col='darkgreen',lwd=2)
legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
text(-.9,.35,paste0("n=",N))
#plot example small effect size points on curve
abline(v=my.es,lty=2,lwd=2,col='red')
valNull=nullPredF(my.es,N)
valAlt=altPredF(my.es,N)
points(pch=19,c(my.es,my.es),c(valNull,valAlt),col="red")
text(my.es+.1,valAlt+.1,paste0("BF=",round(valAlt/valNull,2)),col="red")

#plot example large effect size points on curve
large.es<-1
abline(v=large.es,lty=2,lwd=2,col='red')
valNull=nullPredF(large.es,N)
valAlt=altPredF(large.es,N)
points(pch=19,c(large.es,large.es),c(valNull,valAlt),col="red")
text(large.es+.3,valAlt+.1,paste0("BF=",round(valAlt/valNull,2)),col="red")
```
And for n=100
```{r subjexts100}
#For each effect size, compute density
obsEffectSizes=seq(-1,1.5,.01)

I=length(obsEffectSizes)
altPred=1:I
N=100
for (i in 1:I) {
  altPred[i] = altPredF(obsEffectSizes[i],N)
}

nullPred=nullPredF(obsEffectSizes,N)

#Now we can plot the predictions for all observed effect sizes:
top=max(altPred,nullPred)

plot(type='l',obsEffectSizes,nullPred,ylim=c(0,top),xlab="Observed Effect Size",ylab="Density",main="P(data|H)",col='darkblue',lwd=2)
lines(obsEffectSizes,altPred,col='darkgreen',lwd=2)
legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
text(-.9,.35,paste0("n=",N))

#plot example effect size points on curve
abline(v=my.es,lty=2,lwd=2,col='red')
valNull=nullPredF(my.es,N)
valAlt=altPredF(my.es,N)
points(pch=19,c(my.es,my.es),c(valNull,valAlt),col="red")
text(my.es+.2,valAlt+.1,paste0("BF=",round(valAlt/valNull,2)),col="red")
```

To do a true Bayesian power analysis, could follow Lakens here: https://daniellakens.blogspot.com.au/2016/01/power-analysis-for-default-bayesian-t.html

Do the correlogram and correlation

### 2018-04-10 E2/analysis/bayesianAnalysisForPrereg.Rmd
picking it up after computer broke, luckily all on Google Drive

@@
 
-Assess the size of the effect to inform Bayesian test of next experiment (probably use raw units)
+Assess the size of the effect to inform Bayesian test of next experiment (originally tried this with raw units, but then realized that makes the stats difficult, I forget why)
 
@@
 pre$dv<-"precision"
-all<-rbind(eff,lat,pre)
+allP2E1<-rbind(eff,lat,pre)
 #I tried to do the above with a single command but makes an error even though StackOverflow example works
@@
 ```{r , fig.height=20, fig.width=5}
-all$leftBias<- all$Left - all$Right
-hh<-ggplot(all, aes(x=orientation, y=leftBias)) + facet_grid(.~dv) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
+allP2E1$leftBias<- allP2E1$Left - allP2E1$Right
+hh<-ggplot(allP2E1, aes(x=orientation, y=leftBias)) + facet_grid(.~dv) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
 hh<-hh+stat_summary(fun.data="mean_cl_boot",geom="errorbar",width=.3,conf.int=.95,
@@
 ```{r}
-leftBiasesCanonical<- filter(all,dv=="efficacy" & orientation=="Canonical")$leftBias
+leftBiasesCanonical<- filter(allP2E1,dv=="efficacy" & orientation=="Canonical")$leftBias
 leftBiasCanonical <- mean( leftBiasesCanonical )
-leftBiasesInverted<- filter(all,dv=="efficacy" & orientation=="Inverted")$leftBias
+leftBiasesInverted<- filter(allP2E1,dv=="efficacy" & orientation=="Inverted")$leftBias
 leftBiasInverted<- mean(   leftBiasesInverted  )
@@
 
-The leftBiasDiff for this experiment is `r leftBiasDiff` or in Cohen's d `leftBiasDiff / leftBias_sd`, very healthy.
+The leftBiasDiff for this experiment is `r leftBiasDiff` or in Cohen's d `r leftBiasDiff / leftBias_sd`, very healthy.
 
@@
 
-Do conventional stats.
+Do conventional stats on P2E1 efficacy leftBias.
 ```{r}
@@
 
-aa <- ezANOVA(data=filter(all,dv=="efficacy"), dv=leftBias, within=orientation, wid=subject)
+aa <- ezANOVA(data=filter(allP2E1,dv=="efficacy"), dv=leftBias, within=orientation, wid=subject)
 cat("F=", aa$ANOVA$F, " p=", aa$ANOVA$p, "\n", sep="")
@@
 
-Look at it first in raw units.
+First, a quick look at it with a custom prior in raw units.
 ```{r}
@@
             P2E1_cohensD ) 
+print(cohensDs)
 labels<-c("P1E1","P1E2","P1E2v","P2E1")
-meanCohensDs<- mean(cohensDs)
-
+meanCohensD<- mean(cohensDs)
 ```
 
-Our informed prior will be a Gaussian distribution with mean our mean observed effect, `r meanCohensDs` and sigma `r sd(cohensDs)`.
+Our informed prior will be a Gaussian distribution with mean our mean observed effect (as a Cohen's d), `r meanCohensD` and sigma `r sd(cohensDs)`.
 
@@
 altHypothesisPrior=function(delta) {
-  mean= meanCohensDs
+  mean= meanCohensD
   sigma= sd(cohensDs)
@@
 
+To be honest it is hard to know how to set the sigma (width of the prior in the above graph) because how much to think of the previous estimates as estimating the same thing versus being actually different effects?
+
+To do the custom NEED STILL TO CHANGE THIS TO COHEN'S D?
 Assume left bias difference found for any particular subject reflect independent normal draws with mean μ and variance σ2.
@@
 smaller=.3333
-my.es= meanCohensDs * smaller
+my.es= meanCohensD * smaller
 
@@
 
-Assume effect size is only `r smaller` as big as original `r meanCohensDs`, yielding `r smaller`
+Assume effect size is only `r smaller` as big as original `r meanCohensD`, yielding `r smaller`
  . So, both set prior to half as big but with wider variance and show likelihood functions for whole range but highlight that effect size, for a few different sample sizes.

### 2018-04-11 E2/analysis/bayesianAnalysisForPrereg.Rmd
improving prereg bayesian doc

@@
 To be honest it is hard to know how to set the sigma (width of the prior in the above graph) because how much to think of the previous estimates as estimating the same thing versus being actually different effects?
+Setting it to the sd of the previous results looks reasonable.
 
-To do the custom NEED STILL TO CHANGE THIS TO COHEN'S D?
-Assume left bias difference found for any particular subject reflect independent normal draws with mean μ and variance σ2.
-Set variance across Ss based on previous experiments.
-Take something like the mean variance of those previous experiments.
-Then use that to generate the probability of any particular result.
+To calculate Bayes factors, first
+generate the probability of any particular result according to the null hypothesis and according to the alternative hypothesis.
+For the alternative, integrate over the whole prior distribution of effect sizes, weighting by the prior.
 
@@
 
+#The predicted density of data for the null is below; it's related to the central t distribution.
 nullPredF=function(obs,N) dt(sqrt(N)*obs,N-1)
@@
 You can change N, the sample size, as needed.
-obsEffectSizes=seq(-2,3,.01)
+obsEffectSizes=seq(-2,3,.01) #a list of thousands of possible Cohen's ds we might observe
 N=50
 nullPred=nullPredF(obsEffectSizes,N)
+#nullPred is the height of the density function under the null hypothesis at each cohen's d obsEffectSize
 
@@
 For each nonzero effect size parameter, the distribution of the observed effect follows a noncentral t distribution.
-Hence, to obtain predictions across all nonzero effect sizeS, we need to integrate the alternative model against the noncentral t distribution.
-Here is the code with a simple loop:
+Hence, to obtain predictions across all nonzero effect sizes, we need to integrate the alternative model (with, in the prior, its different height in the density for each possible effect size) against the noncentral t distribution.
 
 altPredIntegrand=function(delta,obs,N) {
+  #call the student-t distribution, with delta the effect size 
   dt(sqrt(N)*obs,N-1,sqrt(N)*delta)*f(delta)
@@
 
-altPredF=function(obsEffectSizes,N) {
-  suppressWarnings( integrate(altPredIntegrand,lower=lo,upper=hi,obs=obsEffectSizes,N=N)$value
+altPredF=function(obs,N) {
+  #integrate delta from lo to hi, for one particular possible observed effect size. Because calculating the 
+  # probability of the observed effect size for each imagined true effect size.
+  suppressWarnings( integrate(altPredIntegrand,lower=lo,upper=hi,obs=obs,N=N)$value
   )
@@
 
-#For each effect size, compute density
+#For each observed effect size, compute density of the alternative - essentially the probability of that 
+#outcome according to the alternative hypothesis.
 I=length(obsEffectSizes)
@@
 my.es= meanCohensD * smaller
-
+my.sd = sd(cohensDs)
 ```
@@
 Assume effect size is only `r smaller` as big as original `r meanCohensD`, yielding `r smaller`
- . So, both set prior to half as big but with wider variance and show likelihood functions for whole range but highlight that effect size, for a few different sample sizes.
+ . So, both set prior to half as big and show likelihood functions for whole range but highlight that effect size, for a few different sample sizes.
+For standard deviation of prior  will use `r my.sd`
 ```{r alternativePrior}
@@
   mean= my.es
-  sigma= sd(cohensDs)
+  sigma= my.sd
   y= dnorm(delta,mean,sigma)

### 2018-04-11 E2/analysis/bayesianAnalysisForPrereg.Rmd
good

@@
 K=1/integrate(altHypothesisPrior,lower=lo,upper=hi)$value
-f=function(delta) K*altHypothesisPrior(delta)
+prior=function(delta) K*altHypothesisPrior(delta)
 
@@
 #Plot Alternative as a density and Null as a point arrow
-maxAlt=max(f(delta))
-plot(dDomain,f(dDomain),typ='n',xlab="Left/top bias difference (Cohen's d)",ylab="Density",ylim=c(0,1.4*maxAlt),main="Models")
+maxAlt=max(prior(delta))
+plot(dDomain,prior(dDomain),typ='n',xlab="Left/top bias difference (Cohen's d)",ylab="Density",ylim=c(0,1.4*maxAlt),main="Models")
 arrows(0,0,0,1.3*maxAlt,col='darkblue',lwd=2)
-lines(dDomain,f(dDomain),col='green',lwd=2)
+lines(dDomain,prior(dDomain),col='green',lwd=2)
 legend("topright",legend=c("Null","Alternative","Previous findings"),col=c('darkblue','green','black'),lwd=2)
@@
   #call the student-t distribution, with delta the effect size 
-  dt(sqrt(N)*obs,N-1,sqrt(N)*delta)*f(delta)
+  dt(sqrt(N)*obs,N-1,sqrt(N)*delta)*prior(delta)
 }
@@
 my.es= meanCohensD * smaller
-my.sd = sd(cohensDs)
+my.sd = sd(cohensDs) * 1.5
 ```
 
-Assume effect size is only `r smaller` as big as original `r meanCohensD`, yielding `r smaller`
+Assume effect size is only `r smaller` as big as original `r meanCohensD`, yielding `r my.es`
  . So, both set prior to half as big and show likelihood functions for whole range but highlight that effect size, for a few different sample sizes.
@@
 
+#Plot this new smaller-effect prior
+dDomain=seq(-.3,2,.01)
+
+#Plot Alternative as a density and Null as a point arrow
+maxAlt=max(altPriorNormalized(delta))
+plot(dDomain,altPriorNormalized(dDomain),typ='n',xlab="Left/top bias difference (Cohen's d)",ylab="Density",ylim=c(0,1.4*maxAlt),main="New prior putting more density over smaller effect sizes")
+arrows(0,0,0,1.3*maxAlt,col='darkblue',lwd=2)
+lines(dDomain,altPriorNormalized(dDomain),col='green',lwd=2)
+legend("topright",legend=c("Null","Alternative","Previous findings"),col=c('darkblue','green','black'),lwd=2)
+
+points(cohensDs, rep(.1,4),pch=19) #plot previous effect sizes
+text(cohensDs, c(.25,.35,.25,.25), labels,cex=.75)
+```
+
+``` {r}
 N=75
 ```
-Have set alternative prior as smaller effect.
 
@@
 
-altPredF=function(obsEffectSizes,N) {
-  suppressWarnings( integrate(altPredIntegrand,lower=lo,upper=hi,obs=obsEffectSizes,N=N)$value
+altPredF=function(obs,N) {
+  suppressWarnings( integrate(altPredIntegrand,lower=lo,upper=hi,obs=obs,N=N)$value
   )
@@
 text(-.9,.35,paste0("n=",N))
-#plot example small effect size points on curve
+#plot example small effect size point on curve
 abline(v=my.es,lty=2,lwd=2,col='red')
@@
 points(pch=19,c(my.es,my.es),c(valNull,valAlt),col="red")
-text(my.es+.1,valAlt+.1,paste0("BF=",round(valAlt/valNull,2)),col="red")
+text(my.es+.1,valAlt+.1,paste0("BF=",round(valAlt/valNull,1)),col="red")
+
+#plot example even smaller effect size points on curve
+abline(v=my.es-.1,lty=2,lwd=2,col='red')
+valNull=nullPredF(my.es-.1,N)
+valAlt=altPredF(my.es-.1,N)
+points(pch=19,c(my.es-.1,my.es-.1),c(valNull,valAlt),col="red")
+text(my.es-.2,valAlt+.1,paste0("BF=",round(valAlt/valNull,1)),col="red")
 
@@
 points(pch=19,c(large.es,large.es),c(valNull,valAlt),col="red")
-text(large.es+.3,valAlt+.1,paste0("BF=",round(valAlt/valNull,2)),col="red")
+text(large.es+.3,valAlt-.1,paste0("BF=",round(valAlt/valNull,1)),col="red")
 ```
-And for n=100
-```{r subjexts100}
+
+```{r echo=FALSE, message=FALSE}
+N=100
+```
+
+And for n=`r N`
+
+```{r subjexts100, echo=FALSE}
 #For each effect size, compute density
@@
 altPred=1:I
-N=100
 for (i in 1:I) {
@@
 points(pch=19,c(my.es,my.es),c(valNull,valAlt),col="red")
-text(my.es+.2,valAlt+.1,paste0("BF=",round(valAlt/valNull,2)),col="red")
+text(my.es+.2,valAlt+.1,paste0("BF=",round(valAlt/valNull,1)),col="red")
+
+#plot example even smaller effect size points on curve
+abline(v=my.es-.1,lty=2,lwd=2,col='red')
+valNull=nullPredF(my.es-.1,N)
+valAlt=altPredF(my.es-.1,N)
+points(pch=19,c(my.es-.1,my.es-.1),c(valNull,valAlt),col="red")
+text(my.es-.15,valAlt,paste0("BF=",round(valAlt/valNull,1)),col="red")
+
 ```
 
+So in summary, the above two graphs with associated blocks of code show how the Bayes factor will be calculated for whatever observed effect size turns up, and is thus suitable for citing in the prereg document.
+To boil it down, the prior we'll be using is simply a Gaussian truncated at zero (bright green curve several plots ago with mean Cohen's d = `r my.es` and sigma = `r my.sd` but it does mean that even with n=100, an observed effect size of `r round(my.es-.1,2)` means the evidence is very ambiguous (Bayes factor close to 1, specifically `r round(valAlt/valNull,1)` )
+
 To do a true Bayesian power analysis, could follow Lakens here: https://daniellakens.blogspot.com.au/2016/01/power-analysis-for-default-bayesian-t.html
-
-Do the correlogram and correlation

### 2018-08-16 E2/analysis/importE2data.Rmd
after partially migrated from GoogleDrive

---
title: "importE2data"
author: "Alex Holcombe"
date: "2/27/2018"
output: html_document
---

1/3 of trials have 1 target (numTargets==1)
targetSide ==0  means targets on both sides

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Modified from importE1data.R

```{r set up functions}

#Loads raw data from MATLAB files for second backwards paper second experiment
require(R.matlab)

mainPath<-".." # to main E2 folder

#Second experiment of second backwards-letters paper. random orientation each frame.
rawDataPath<- file.path(mainPath,
                            "Data/RawData/Data")
#raw data path containing .mat file for each subject

source(file.path("turnMATintoMeltedDataframe.R"))
#cat("Here are the files in the raw data directory")
#dir(rawDataPath)
```

```{r}
library(dplyr)
readInAllFiles<- function(rawDataPath) {

  files <- dir(path=rawDataPath,pattern='.mat')  #find all data files in this directory
  dfAll<-data.frame()
  for (i in 1:length(files)) { #read in each file
    fileThis<- file.path(rawDataPath,files[i])
    rawDataLoad=tryCatch(
      readMat(fileThis),
      error=function(e) {
        stop( paste0("ERROR reading the file ",fileThis," :",e) )
      } )
    apparentSubjectName <- strsplit(files[i],split="_")[[1]][1]
    subjectName<- rawDataLoad$participantID[[1]]
    if (apparentSubjectName != subjectName) {
      stop( paste0("WARNING apparentSubjectName",apparentSubjectName," from filename does not match subjectName in data structure",subjectName) )
    }
    else { cat(paste0(' Subject ',apparentSubjectName,' read in')) }
    rawDataLoad$file <- files[i]

    dfThis<- turnMATintoMeltedDataframe(rawDataLoad)

    tryCatch(
      dfAll<-rbind(dfAll,dfThis), #if fail to bind new with old,
      error=function(e) { #Give feedback about how the error happened
        cat(paste0("Tried to merge but error:",e) )
      } )
  }
  return(dfAll)
}
```

```{r read raw data}

E<- readInAllFiles(rawDataPath)
#TT<- as_tibble(E).  Seems to have no effect because weird hybrid type

#Calculate the serial position error
E$SPE<- E$respSP - E$targetSP

#give targetSide values that make  sense, like "dual", "left", and "right"
E<- E %>% mutate(targetSide = case_when(targetSide==0 ~ "both", 
                                       targetSide==1 ~"left", targetSide==2 ~"right",
                                       TRUE ~ "unexpected value"))
```

Next is very TIME CONSUMING.

Break out into target position -3, -2 etc. the targetStreamOrient

```{r veryTimeConsuming}
#Need to decompose into SPE=-3 orientation, SPE=-2 orientation, etc.
#Hm, can I use list apply to go through every row and pull out relevant bits?
#Very time-consuming!
E$orientMinus3<- -99; E$orientMinus2<- -99; E$orientMinus1<- -99; E$orient0<- -99; E$orientPlus1<- -99; E$orientPlus2<- -99;

for (i in 1:nrow(E)) {
  #row<- E[i,]
  #orientsThis <- row$targetStreamOrientations[[1]]
  #targetPos <- row$targetSP
  #If targetSP == NaN, it's a one-target trial and wrong side
  if (is.nan(E[i,]$targetSP)) {
    E[i,]$orientMinus3<- NaN
  } else {
    E[i,]$orientMinus3<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -3    ]
    E[i,]$orientMinus2<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -2    ]
    E[i,]$orientMinus1<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -1    ]
    E[i,]$orient0<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP    ]
    E[i,]$orientPlus1<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP +1    ]
    E[i,]$orientPlus2<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP +2    ]
  }
}

# More explicit way of doing it below but might take more time.
# for (i in 1:nrow(E)) {
#   row<- E[i,]
#   orientsThis <- row$targetStreamOrientations[[1]]
#   targetPos <- row$targetSP
#   #If targetSP == NaN, it's a one-target trial and wrong side
#   if (is.nan(targetPos)) {
#     E[i,]$orientMinus3<- NaN
#   }
#   E[i,]$orientMinus3<- orientsThis[targetPos-3]
#   #E[i,]$orientMinus3<- E[i,]$targetStreamOrientations[[1]][targetPos-3]
#   E[i,]$orientMinus2<- orientsThis[targetPos-2]
#   E[i,]$orientMinus1<- orientsThis[targetPos-1]
#   E[i,]$orient0<- orientsThis[targetPos] #should be same as $targetOrient
#   E[i,]$orientPlus1<- orientsThis[targetPos+1]
#   E[i,]$orientPlus2<- orientsThis[targetPos+2]
# }
```

Saving data.
```{r saveData}
backwards2_E2 <- E
pathToSave <- file.path( "..","Data")
saveRDS(backwards2_E2, file = "../Data/backwards2E1_rawDataFromMAT.rda", compress="bzip2")
```

### 2018-08-16 E2/analysis/inspectHists.Rmd
after partially migrated from GoogleDrive

---
title: "inspect histograms"
author: "Alex Holcombe"
date: "2/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data

```{r load data, echo=FALSE, message=FALSE}

#Import raw data
dataPath<- file.path("../Data/")
#Experiment was administered by MATLAB
#.mat file been preprocessed into melted long dataframe by importE1data.Rmd
data<- readRDS( file.path(dataPath, "backwards2E2_rawDataFromMAT.rda") )

#tidy data
library(dplyr)
df<- data
#to work with dplyr, can't have array field like letterSeq
df$letterSeq<- NULL

```

Set up the experiment-specific parameters

```{r detailsNeeded}

numItemsInStream<- length( data$letterSeq[1,] )  
minSPE<- -17; maxSPE<- 17
annotateIt<-TRUE

numSs<-length(unique(df$subject))
cat( paste0('Total num Ss=',numSs) )
```

Plot sample histograms, make sure plotting working.

```{r}
dg<- df %>% filter(subject < "AC")

library(ggplot2)
g=ggplot(dg, aes(x=SPE)) + facet_grid(subject+orientation~stream) #,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
show(g)

```

Plot all Ss
```{r, fig.height=200, fig.width=10}
g=ggplot(df, aes(x=SPE)) + facet_grid(subject+orientation~stream) #,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times

show(g)
```

CY is bimodal but not outrageously so.

Subject AP looks to both have close to zero efficacy in some conditions and disregards instructions, reporting mostly after the cue.

BJ has essentially zero efficacy and may have tended to report letters from beginning of stream in one condition.

SsWhoLookWeird<-c("CY","AP","BJ")

We'll see whether the statistical comparison of mixture modeling to guessing distribution flags these as not being fit significantly better by mixture model.

### 2018-08-16 E2/analysis/mixModelE2data.Rmd
after partially migrated from GoogleDrive

--- 2018-02-28 E1/mixModelE1data.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE,  comment = NA)
-# http://htmlpreview.github.io/?https://github.com/alexholcombe/MixtureModelRSVP/blob/master/compareMATLABtoR.html
 ```
@@
 
-#Import raw data
-dataPath<- file.path("Data/")
+#Import data that was changed from MATLAB raw to R dataframe by importE2data.Rmd
+dataPath<- file.path("..","Data")
 #Experiment was administered by MATLAB
 #.mat file been preprocessed into melted long dataframe by importE1data.Rmd
-data<- readRDS( file.path(dataPath, "backwards2E1_rawDataFromMAT.rda") )
+data<- readRDS( file.path(dataPath, "backwards2E2_rawDataFromMAT.rda") )
 
@@
 
+Why are there NaNs for targetSP?
+```{r}
+table(is.nan(df$targetSP))
+
+```
+
 Sanity-check some data by graphing.
 
-Investigate whether the pathological identicalness of inverted/canonical is already at this point
+Investigate whether pathological identicalness of inverted/canonical happens, that happened in E1 . Seems it doesn't
+
 ```{r}
@@
 show(g)
+```
+
+Try to analyze one bit of data.
+WHY ARE THERE NANS IN TARGETSP?
+```{r}
+library(mixRSVP)
+dg<-df %>% filter(subject=="AA")
+analyzeOneCondition(dg,numItemsInStream,parameterBounds(),nReplicates=5)
+
+analyzeOneConditionDF(df %>% filter(subject=="AA"),numItemsInStream,parameterBounds(),nReplicates=5)
+
 ```
@@
 
-estimates<- df  %>%  #  filter(subject=="AA") %>%
+estimates<- df  %>% filter(subject=="AA") %>%
     group_by_(.dots = condtnVariableNames) %>%  #.dots needed when you have a variable containing multiple factor names
@@
 
-write.csv(estimates, file = "Results/backwards2E1_paramEstimates.csv")
+write.csv(estimates, file = "Results/backwards2E2_paramEstimates.csv")
 #saveRDS(estimates, file = "Data/backwards2E1_paramEstimates.rda", compress="bzip2")

### 2018-08-16 E2/analysis/validateE2data.Rmd
after partially migrated from GoogleDrive

---
title: "importE2data"
author: "Alex Holcombe"
date: "2/27/2018"
output: html_document
---

1/3 of trials have 1 target (numTargets==1)
targetSide ==0  means targets on both sides

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Modified from importE1data.R

```{r set up functions}

#Loads raw data from MATLAB files for second backwards paper second experiment
require(R.matlab)

GoogleDrivePath<-".." # to main E2 folder

#Second experiment of second backwards-letters paper. random orientation each frame.
rawDataPath<- file.path(GoogleDrivePath,
                            "Data/RawData/Data")
#raw data path containing .mat file for each subject

source(file.path("turnMATintoMeltedDataframe.R"))
cat("Here are the files in the raw data directory")
dir(rawDataPath)
```

```{r}
readInAllFiles<- function(rawDataPath) {

  files <- dir(path=rawDataPath,pattern='.mat')  #find all data files in this directory
  dfAll<-data.frame()
  for (i in 1:length(files)) { #read in each file
    fileThis<- file.path(rawDataPath,files[i])
    rawDataLoad=tryCatch(
      readMat(fileThis),
      error=function(e) {
        stop( paste0("ERROR reading the file ",fileThis," :",e) )
      } )
    apparentSubjectName <- strsplit(files[i],split="_")[[1]][1]
    subjectName<- rawDataLoad$participantID[[1]]
    if (apparentSubjectName != subjectName) {
      stop( paste0("WARNING apparentSubjectName",apparentSubjectName," from filename does not match subjectName in data structure",subjectName) )
    }
    else { cat(paste0(' Subject ',apparentSubjectName,' read in')) }
    rawDataLoad$file <- files[i]

    dfThis<- turnMATintoMeltedDataframe(rawDataLoad)

    tryCatch(
      dfAll<-rbind(dfAll,dfThis), #if fail to bind new with old,
      error=function(e) { #Give feedback about how the error happened
        cat(paste0("Tried to merge but error:",e) )
      } )
  }
  return(dfAll)
}
```

```{r read raw data}

E<- readInAllFiles(rawDataPath)

#Calculate the serial position error
E$SPE<- E$respSP - E$targetSP

```

Why are there NaNs for targetSP? Because that's when only one target (targetSide==1 or ==2)

```{r}
table(E$subject, E$targetSP,useNA="always")
```

Show how targetSP works
```{r}
library(dplyr)
En <- E
En$letterSeq<- NULL #otherwise can't use dplyr
CU <- En %>% filter(subject=="CU")
table(CU$target,CU$targetSide, useNA="always")
#targetSide ==0  means targets on both sides 
table(CU$targetSide, CU$targetSP, useNA="always")
```

Why is it that letterSeq is still in with all its values, but allLetterOrientations etc only has one? As you can see below, you literally can't slice

```{r}
CUall <- E[ ,E$subject=="CU" ]
i<- E$subject == "CU"
E[i,] #doens't work. You can't slice a dataframe that has a list (letterSeq) as a member, yet no error message!
subset(E, subject=="CU")
j<- En$subject == "CU"
EnCU<- En[j,]
```

Load in one file and inspect it
```{r}

source(file.path("turnMATintoMeltedDataframe.R"))

fileThis<- file.path(rawDataPath, "CU_18-05-10_1.mat")

rawDat=tryCatch(
  readMat(fileThis),
  error=function(e) {
    stop( paste0("ERROR reading the file ",fileThis," :",e) )
  } )
str(rawDat)
CUmelted<- turnMATintoMeltedDataframe(rawDat)

```
Somehow, what would seem to be the second trial has NaN for lots of stuff

 $ allLetterOrder         : num [1:216, 1:2, 1:16] 13 6 9 1 10 16 11 1 12 8 ...
 $ allLetterOrientations  : num [1:216, 1:16] 1 0 1 0 0 0 1 0 1 0 ...
 $ allTargets             : num [1:216, 1:2] 5 NaN 6 8 7 7 6 6 6 8 ...
 $ allTargetOrientations  : num [1:216, 1:2] 0 NaN 1 1 0 1 1 1 0 0 ...
 $ allResponses           : num [1:216, 1:2] 9 NaN 1 3 5 11 5 3 8 11 ...
 $ allResponseOrientations: num [1:216, 1:2] 0 NaN 0 0 0 1 0 0 0 0 ...
 $ allResponseOrder       : num [1:216, 1] 2 2 1 2 2 1 2 1 1 1 ...
 $ allRTs                 : num [1:216, 1:2] 12.78 NaN 1.81 6.19 10.19 ...

Inspect another one 
```{r}
fileThis<- file.path(rawDataPath, "BC_18-01-25_1.mat")

rawDat=tryCatch(
  readMat(fileThis),
  error=function(e) {
    stop( paste0("ERROR reading the file ",fileThis," :",e) )
  } )
str(rawDat)

```

Sanity visual check that see histogram.
```{r sanity visual check}

sanityVisualCheck<-TRUE
if (sanityVisualCheck) {
  library(ggplot2)
  g=ggplot(E, aes(x=SPE))
  #plot data
  g<-g+geom_histogram(binwidth=1)
  show(g)

  h<-ggplot(E,aes(x=SPE)) + geom_histogram(binwidth=1)
  show(h)
}

```

Meaning of condition variable
* 1 - upright
* 2 - inverted

Make condition names more sensible - stream (Left/Right) and orientation (Canonical/Inverted)
```{r}
E2<- E
names(E2)[names(E2) == 'target'] <-'stream'
names(E2)[names(E2) == 'condition'] <-'orientation'
E2$stream[ E2$stream==1 ] <- "Left"
E2$stream[ E2$stream==2 ] <- "Right"
#mutate condition to Orientation
E2$orientation[ E2$orientation==1 ] <- "Canonical"
E2$orientation[ E2$orientation==2 ] <- "Inverted"

```

Save data
```{r}
#save dataframe into proper library data directory
backwards2_E2<-E2
saveRDS(backwards2_E2, file = "../Data/backwards2E2_rawDataFromMAT.rda", compress="bzip2")
```

Look at 1-target performance, understand that variable better because didn't exist in previous experiment I think.

The first table printed below could easily fool you into thinking that stream (Left, Right) and targetSide (1,2) are not consistent. We think that's because it records something for both sides (stream variable) even when there is only 1 target (targetSide == 1 or 2 rather than 0, which means two targets).

ALEX TO-DO
Disaggregating and thus preserving the all variables, like allLetterOrientations and allTargetORientations
Also maybe give targetSide values that make more common sense, like "dual", "left", and "right"

But the second table below shows that they make sense, as they are only inconsistent when allResponseOrientations==NaN.  But what the hell are those trials where  allResponseOrientations==NaN?
```{r}
library(dplyr)
E2$letterSeq <- NULL
one <- E2 %>% filter(numTargets==1)

table(one$stream,one$targetSide,useNA="always")

table(E2$stream,E2$targetSide,E2$numTargets, useNA="always")

table(one$stream,one$targetSide,one$orientation,useNA="always")

table(one$stream,one$targetSide,one$allResponseOrientations,useNA="always")
```

### 2018-08-16 E2/ExperimentCode/DualRSVP_randomOrientationWithinStreams.m

@@ -1,0 +1,1109 @@
+% Dual RSVP stream demo code  
+% PTG 11/07/12
+
+Screen('CloseAll')
+clear all; 
+
+%Screen('Preference', 'SkipSyncTests', 1)
+
+addpath(genpath('/Users/supalab/Documents/MATLAB/BackwardsLetters/E2/'));
+
+AssertOpenGL;
+
+% Initialise experiment parameters
+saveDirectory = '/Users/supalab/Documents/MATLAB/BackwardsLetters/E2/Data/';
+userDirectory = '/Users/supalab/Documents/MATLAB/BackwardsLetters/E2/UserData/';
+nTrials = 108;            % Per condition upright/inverted - Must be a multiple of 6 as one third are single target, and 1/6th are left/right.
+nPracticeTrials = 24;    % Must be a multipe of 12 as one third are single target, 1/6th are left/right, and 1/12th are canonical/inverted.
+nSingleTargets = 3;        % Ratio to dual i.e. 1 single every 3 trials
+nBlocks = 4;            % 4
+nConditions = 2;
+nSessions = 1;          % 2
+praticeOrder = 1; % 1 = canonical practice first, 2 = inverted practice first.
+
+% Randomise
+randomSeed = sum(100*clock);
+thisStream = RandStream.create('mt19937ar', 'seed', randomSeed);
+RandStream.setGlobalStream(thisStream); 
+
+
+while true
+    participantID = upper(input('Enter two-letter participant initials: ', 's'));
+    if length(participantID)==2
+        break
+    end
+end
+
+oldDirectory = cd(userDirectory);
+userDataFile = [participantID '_RotatedDualRSVP_Exp2_UserData.mat'];
+
+if ~exist(userDataFile,'file')
+   % Make data file
+   % Check number of existing records
+   allUserDataFiles = what;
+   numUserDataFiles = size(allUserDataFiles.mat,1);
+   
+   % Assign next session
+   sessionOrder = randperm(2);
+   nextSession = 1;
+   totalTime = [];
+   
+   % Save
+   save(userDataFile,'sessionOrder','nextSession','totalTime');
+   skipInstructions = 0;
+   
+else
+    
+    load(userDataFile);
+    skipInstructions = 1;
+    
+end
+
+cd(oldDirectory);
+
+if nextSession > nSessions
+    fprintf('\n\nAll sessions completed!');
+    return;
+end
+
+% Initialise equipment parameters
[…]

### 2018-08-16 E2/PilotData/Data/wranglingAndPlotting.R

@@ -1,0 +1,87 @@
+library(R.matlab)
+library(ggplot2)
+library(reshape2)
+rm(list=ls())
+
+setwd('~/Google Drive/Backwards paper/secondPaper/E2/Data/')
+
+datafiles <- list.files(pattern = '^[A-Z][A-Z]')
+
+chars <- 'ABCDEFGJKLPQRTUVY'
+chars <- unlist(strsplit(chars,split=''))
+
+IDs <- c()
+
+nRows <- 200* length(datafiles)*2
+
+masterDF <- expand.grid(ID = character(nRows), condition = factor(x=character(1), levels = c(1,2), labels = c('Normal', 'Rotated')), responseOrientation = numeric(1),trial = numeric(1), stream = numeric(1), responsePosRelative = numeric(1), stringsAsFactors = F)
+
+start <- 1
+for(file in datafiles){
+  ID <- paste(unlist(strsplit(file, split =''))[1:2], collapse = '')
+  IDs <- c(IDs, ID)
+  
+  
+  matFile <- readMat(file) 
+  
+  
+  responsePosRelativeMatrix <- matrix(NA, nrow = 200, ncol = 5)
+  responsePosMatrix <- matrix(NA, nrow = 200, ncol = 5)
+  
+  for(stream in 1:2){
+    for(trial in 1:200){
+      thisResponse <- matFile$allResponses[trial, stream]
+      thisStream <- matFile$allLetterOrder[trial, stream,]
+      responsePos <- which(thisStream == thisResponse)
+      responsePosRelative <- responsePos - matFile$allTargets[trial, stream]
+      responsePosRelativeMatrix[trial,stream] <- responsePosRelative
+      responsePosMatrix[trial,stream] <- responsePos
+    }
+  }
+  responsePosRelativeMatrix[,3] <- matFile$allConditions[1,] 
+  responsePosRelativeMatrix[,4] <- matFile$allResponseOrientations[,1]
+  responsePosRelativeMatrix[,5] <- 1:200
+  #SPE left, SPE right, Target orientation, response orientation, trial
+  
+  responsePosMatrix[,3] <- matFile$allConditions[1,]
+  responsePosMatrix[,4] <- 1:200
+  
+  responsePosRelativeDF <- data.frame(responsePosRelativeMatrix)
+  responsePosRelativeDF <- cbind(rep(ID, times=200), responsePosRelativeDF)
+  
[…]

### 2018-08-16 E2/analysis/temp.R

@@ -1,0 +1,23 @@
+
+#Import raw data
+dataPath<- file.path("E2/Data")
+#Experiment was administered by MATLAB
+#.mat file been preprocessed into melted long dataframe by importE1data.Rmd
+file.exists( file.path(dataPath, "backwards2E2_rawDataFromMAT.rda") )
+getwd()
+data<- readRDS( file.path(dataPath, "backwards2E2_rawDataFromMAT.rda") ) 
+
+#tidy data
+library(dplyr)
+df<- data
+#to work with dplyr, can't have array field like letterSeq
+df$letterSeq<- NULL
+
+numItemsInStream<- length( data$letterSeq[1,] )  
+minSPE<- -17; maxSPE<- 17
+
+dg<- df %>% filter(subject < "AC")
+
+
+library(mixRSVP)
+analyzeOneCondition(df %>% filter(subject=="AA"),numItemsInStream,parameterBounds(),nReplicates=5)

### 2018-08-16 E2/analysis/tempAnalyzeOneCondition.R

@@ -1,0 +1,110 @@
+#Used by analyzeOneCondition and
+doFit <- function( SPE, minSPE, maxSPE, pseudoUniform, paramBounds ) {
+  #Calculate parameter guess
+  startingParams<- parametersGuess( paramBounds$lower, paramBounds$upper )
+  fit<- fitModel(SPE, minSPE, maxSPE, pseudoUniform, startingParams, paramBounds)
+  warns<- fit$warnings
+  fit<- fit$content
+  fitMetric<- fit$value
+  #Need to replace any NULL warnings with None, otherwise assigning this to other dataframes will delete that column
+  fitNoNames<-as.numeric(fit) #Otherwise p1, p2, p3 names preserved, which is confusing
+  return( list(efficacy=fitNoNames[1], latency=fitNoNames[2], precision=fitNoNames[3], val=fitMetric, warnings=warns) )
+}
+
+fitOneCondition<- function(df, numItemsInStream, paramBounds, nReplicates=3) {
+  #nReplicates. Number of times to repeat each fit with different starting values. Pat maybe used 100
+  # Calculate the domain of possible serial position errors.
+  possibleTargetSP<- sort(unique(df$targetSP))
+  minTargetSP <- min(possibleTargetSP)
+  maxTargetSP <- max(possibleTargetSP)
+  minSPE <- 1 - maxTargetSP
+  maxSPE <- numItemsInStream - minTargetSP
+  
+  print(minSPE)
+  print(maxSPE)
+  print(df$targetSP)
+  #calculate the guessing distribution, empirically (based on actual targetSP)
+  pseudoUniform <- createGuessingDistribution(minSPE,maxSPE,df$targetSP,numItemsInStream)
+  
+  # Set some model-fitting parameters.
+  #fitMaxIter <- 10^4# Maximum number of fit iterations
+  #fitMaxFunEvals <- 10^4# Maximum number of model evaluations
+  
+  for (n in 1:nReplicates) { #fit the model many times (with different starting parameters)
+    
+    paramsPlusNegLogLikelihood<- doFit( df$SPE, minSPE, maxSPE, pseudoUniform, paramBounds )
+    #print(paramsPlusNegLogLikelihood)
+    #Save the best estimate
+    if (n==1) {
+      bestEstimate <- paramsPlusNegLogLikelihood
+    } else {
+      if (paramsPlusNegLogLikelihood$val < bestEstimate$val)
[…]

### 2018-08-16 E2/analysis/turnArrayColumnsIntoLists.R

@@ -1,0 +1,38 @@
+#Turn dataframe with array columns into list columns, so dplyr can work
+
+#d<- data.frame(x=c(1,2,3,4), y= matrix(c(1,2,3, 11,12,13, 20,21,22, 23,24,25) ,nrow=4,ncol=3) )
+#Above gives different results, variables auto-turned into y.1, y.2, etc
+d<- data.frame(x=c(1,2,3,4))
+y= matrix(c(1,2,3, 11,12,13, 20,21,22, 23,24,25) ,nrow=4,ncol=3)
+d$y = y
+d$z= y
+#d now looks like my dataframe
+
+#non-vectorized
+d$l <- NaN
+for (i in 1:nrow(d)) {
+  d[i,]$l <- list(d$y[i,])
+}
+
+#vectorized https://stackoverflow.com/questions/51622017/change-all-array-columns-of-a-data-frame-into-lists
+library(purrr)
+e<- map_if(d,is.matrix,~split(.,seq(nrow(.)))) %>% as_tibble
+#map_if(d,is.matrix,~split(.,seq(nrow(.)))) %>% as_tibble %>% print.data.frame
+
+f<- e %>% mutate(a=x)  #works
+
+###
+d<- data.frame(x=c(1,2,3,4))
+y= matrix(c(1,2,3, 11,12,13, 20,21,22, 23,24,25) ,nrow=4,ncol=3)
+d$y = y
+d$z= y
+e<- map_if(d,is.matrix,~split(.,seq(nrow(.)))) 
+
+
+#turn the weird array things into a list column
+#As explained here, tibbles can have list-columns https://cran.r-project.org/web/packages/tibble/vignettes/tibble.html
+# Here is how to build one:
+tt<-tibble(x=1:3)
+tt$y=list(1:5,1:10,1:20)
+tt$z = NaN
+tt$z[1] <- list((letterSeq[1,]))

### 2018-08-16 E2/analysis/turnMATintoMeltedDataframe.R

@@ -1,0 +1,102 @@
+#Loads raw data from MATLAB files for second backwards paper first experiment
+require(R.matlab)
+
+#define function to be used below
+turnMATintoMeltedDataframe<- function(fromMAT) {
+  #It was imported as a list of lists
+  mydata<-fromMAT
+  numTrials<- length( unlist( mydata$allConditions ) )
+  #Some of the list members are one-off fields pertaining to entire experiment. Remove them, put in miscInfo
+  miscInfo <- list()
+  membersToDelete<- list()
+  #loop through MATLAB .mat file data structure
+  for (i in 1:length(mydata)) {
+    thisListMember <- mydata[i]
+    name<- names(thisListMember)
+    len<- length(unlist(thisListMember))
+    #cat(name, len,'\n')           
+    if (len < numTrials) { #one-off information, not whole experiment
+      miscInfo[name] <- thisListMember
+      membersToDelete<-c(membersToDelete,name)
+    } 
+  }
+  for (m in membersToDelete) {
+    mydata[m]<-NULL #remove from list to leave only every-trial fields
+  }
+  
+  #Remaining fields are matrices, one dimension of which is the trial number. Other dimension is whichTarget in dual-stream case
+  #To melt, need to unpack the matrix, going through the second dimension and taking it out, putting at end to create flat dataframe
+  # with one row for each trial.
+  meltedDf <- data.frame()
+  for (target in 1:2) {
+    dfThis <- data.frame(trial=seq(1,numTrials))
+    dfThis$target<- target
+    for ( i in 1:length(mydata) ) {  #Go through each field, turn into column
+      thisMember <- mydata[i]
+      name <- names( thisMember )
+      thisMember <- thisMember[[name]] #for some reason it's always a list of 1 (with the name name)
+      #if only one dimension then it is a condition name applying to both streams
+      thisMember <- drop(thisMember) #remove any singleton dimensions
+      ndim <- length( dim( thisMember ) ) #number of dimensions
+      #if (name=="allLetterOrientations" | name=="allTargets") {
+      #  print(paste0("name=",name," thisMember=",thisMember," ndim=",ndim))
+      #}
+      if ((ndim==0)) {  #it is a condition name applying to both streams
+        dfThis[[name]] <- thisMember
+      } else if (ndim==2) { #else it has separate values for each target
[…]

### 2018-08-16 E2/analysis/importE2data.Rmd
gettting betterall the timee

@@
 
-Break out into target position -3, -2 etc. the targetStreamOrient
+Break out into target position -3, -2 etc.
+the targetStreamOrient variable that contains the orientaiton of each item.
+This will allow seeing whether having several orientations in a row triggers the lateral bias.
 
@@
   if (is.nan(E[i,]$targetSP)) {
-    E[i,]$orientMinus3<- NaN
+    E[i,]$orientMinus3 <- NaN
   } else {
-    E[i,]$orientMinus3<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -3    ]
-    E[i,]$orientMinus2<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -2    ]
-    E[i,]$orientMinus1<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -1    ]
-    E[i,]$orient0<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP    ]
-    E[i,]$orientPlus1<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP +1    ]
-    E[i,]$orientPlus2<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP +2    ]
+    E[i,]$orientMinus4 <- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -4    ]
+    E[i,]$orientMinus3 <- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -3    ]
+    E[i,]$orientMinus2 <- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -2    ]
+    E[i,]$orientMinus1 <- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -1    ]
+    E[i,]$orient0 <- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP    ]
+    E[i,]$orientPlus1 <- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP +1    ]
+    E[i,]$orientPlus2 <- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP +2    ]
   }

### 2018-08-16 E2/analysis/validateE2data.Rmd
gettting betterall the timee

@@
 
-Modified from importE1data.R
+Checking some stuff out
 
@@
 
-GoogleDrivePath<-".." # to main E2 folder
+mainPath<-".." # to main E2 folder
 
 #Second experiment of second backwards-letters paper. random orientation each frame.
-rawDataPath<- file.path(GoogleDrivePath,
+rawDataPath<- file.path(mainPath,
                             "Data/RawData/Data")
@@
 
-```{r read raw data}
+```{r readRawData}
 
@@
 
-Why is it that letterSeq is still in with all its values, but allLetterOrientations etc only has one? As you can see below, you literally can't slice
+The below works now that turnMATintoMeltedDataframe takes care of the list elements
 
 ```{r}
-CUall <- E[ ,E$subject=="CU" ]
+#CUall <- E[ ,E$subject=="CU" ]
+CUall <- E %>% filter(subject=="CU")
 i<- E$subject == "CU"
-E[i,] #doens't work. You can't slice a dataframe that has a list (letterSeq) as a member, yet no error message!
+E[i,] #previously didn't work. You can't slice a dataframe that has a list (letterSeq) as a member.
 subset(E, subject=="CU")
@@
 
-ALEX TO-DO
-Disaggregating and thus preserving the all variables, like allLetterOrientations and allTargetORientations
+I Disaggregated allLetterOrientations - done by importE2data.Rmd which renames it targetStreamOrientations
+
 Also maybe give targetSide values that make more common sense, like "dual", "left", and "right"
@@
 
-table(one$stream,one$targetSide,one$allResponseOrientations,useNA="always")
+#table(one$stream,one$targetSide,one$allResponseOrientations,useNA="always")
 ```

### 2018-08-16 E2/analysis/turnMATintoMeltedDataframe.R

@@ -1,5 +1,6 @@
 #Loads raw data from MATLAB files for second backwards paper first experiment
 require(R.matlab)
+library(tibble)
 
 #define function to be used below
 turnMATintoMeltedDataframe<- function(fromMAT) {
@@ -79,7 +80,7 @@ turnMATintoMeltedDataframe<- function(fromMAT) {
   library(purrr) #for map_if
   #data.frames and tibbles must have all atomic or list entries, not arrays, therefore
   #turn any remaining arrays (like letterSequence) into list so tidyverse operations can be done
-  E<- map_if(E,is.matrix,~split(.,seq(nrow(.)))) %>% as_tibble #https://stackoverflow.com/questions/51622017/change-all-array-columns-of-a-data-frame-into-lists
+  E<- map_if(E,is.matrix,~split(.,seq(nrow(.)))) %>% tibble::as_tibble() #https://stackoverflow.com/questions/51622017/change-all-array-columns-of-a-data-frame-into-lists
   
   names(E)[names(E) == 'allTargets'] <- 'targetSP'
   # allLetterOrder         : num [1:216, 1:2, 1:16] 13 6 9 1 10 16 11 1 12 8 ...

### 2018-08-16 E2/analysis/mixModelE2data.Rmd
starting analysis

@@
 library(dplyr)
-df<- data
-#to work with dplyr, can't have array field like letterSeq
-df$letterSeq<- NULL
+#to work with dplyr, can't have array field like letterSeq, so before I fixed it in turnMatIntoMeltedDataframe, 
+#I dropped the letterSeq thusly:
+#df$letterSeq<- NULL
 
-numItemsInStream<- length( data$letterSeq[1,] )  
+numItemsInStream<- length( data[1,"letterSeq"]$letterSeq$`1` )  
 minSPE<- -17; maxSPE<- 17
@@
 
-Why are there NaNs for targetSP?
+Hm, seems like funky StackOverflow way of turning matrix elements into list gave it this weird name "1" with backticks around it
+
+data[1,"letterSeq"]$letterSeq$`1`
+
+Make condition names more sensible - stream (Left/Right) and orientation (Canonical/Inverted)
 ```{r}
-table(is.nan(df$targetSP))
+E2<- data
+names(E2)[names(E2) == 'target'] <-'stream'
+names(E2)[names(E2) == 'condition'] <-'orientation'
+E2$stream[ E2$stream==1 ] <- "Left"
+E2$stream[ E2$stream==2 ] <- "Right"
+E2$targetSide[ E2$targetSide==0 ] <- "Both"
+E2$targetSide[ E2$targetSide==1 ] <- "Left"
+E2$targetSide[ E2$targetSide==2 ] <- "Right"
+
+#mutate condition to Orientation
+E2$orientation[ E2$orientation==1 ] <- "Canonical"
+E2$orientation[ E2$orientation==2 ] <- "Inverted"
 
 ```
+
+NaNs for targetSP indicate one-target trials, which are also indicated by targetSide==0
+```{r}
+
+table(E2$targetSP, E2$targetSide, dnn=c("targetSP","targetSide"), useNA="ifany")
+
+```
+Proportion of one-target trials seems to be `r table(is.nan(df$targetSP))["TRUE"] / sum( table(is.nan(df$targetSP)) )
+`
 
@@
 ```{r}
-dg<- df %>% filter(subject < "AC")
+dg<- E2 %>% filter(subject < "AC")
 
@@
 Try to analyze one bit of data.
-WHY ARE THERE NANS IN TARGETSP?
 ```{r}
 library(mixRSVP)
-dg<-df %>% filter(subject=="AA")
-analyzeOneCondition(dg,numItemsInStream,parameterBounds(),nReplicates=5)
+dg<-E2 %>% filter(subject=="AA")
+dg<-dg %>% filter( !is.nan(targetSP) )
 
-analyzeOneConditionDF(df %>% filter(subject=="AA"),numItemsInStream,parameterBounds(),nReplicates=5)
+analyzeOneCondition(dg,numItemsInStream,parameterBounds(),nReplicates=2)
+
+analyzeOneConditionDF(dg,numItemsInStream,parameterBounds(),nReplicates=2)
 
@@
 
-Fit mixture model with R (if fitting not previously done and cached). 
+Maybe do the one-target trials first, might be faster.
+```{r, echo=FALSE}
+one<- dg %>% filter(targetSide != "Both")
+
+condtnVariableNames <- c("subject","targetSide", "stream", "orientation")  
+#It seems there are no rows for one-target (targetSide!="Both") and stream!=targetSide,
+# which would mean that I can't test for stream swaps in the one-target condition
+
+estimates<- dg  %>%  #  filter(subject=="AA") %>%
+    group_by_(.dots = condtnVariableNames) %>%  #.dots needed when you have a variable containing multiple factor names
+    do(  analyzeOneConditionDF(.,numItemsInStream,parameterBounds(),nReplicates=5) )
+
+```
+
+Fit mixture model (if fitting not previously done and cached). 
 ```{r, echo=FALSE}
@@
 #Bizarrely, when knitting this often quits after only 6 parameter estimates. Don't know if there's some function fitting error catch problem
-
-condtnVariableNames <- c("subject","stream", "orientation")
 

### 2018-08-16 E2/analysis/mixModelE2data.Rmd
right

@@
 ```
-Proportion of one-target trials seems to be `r table(is.nan(df$targetSP))["TRUE"] / sum( table(is.nan(df$targetSP)) )
-`
+Proportion of one-target trials seems to be 
+`nrow(E2 %>% filter(targetSide!="Both")) / nrow(E2)`
 
@@
 dg<-E2 %>% filter(subject=="AA")
-dg<-dg %>% filter( !is.nan(targetSP) )
+dg<-dg %>% filter( !is.nan(targetSP) ) #excluding 1-target trials rows reporting stream on opposite side. Eventually, use those to evaluate swaps.
 
@@
 
-Maybe do the one-target trials first, might be faster.
+Analyse one subject, all conditions.
 ```{r, echo=FALSE}
-one<- dg %>% filter(targetSide != "Both")
+#one<- dg %>% filter(targetSide != "Both")
 
 condtnVariableNames <- c("subject","targetSide", "stream", "orientation")
-#It seems there are no rows for one-target (targetSide!="Both") and stream!=targetSide,
-# which would mean that I can't test for stream swaps in the one-target condition
 
-estimates<- dg  %>%  #  filter(subject=="AA") %>%
+estimatesAA<- dg  %>%  
     group_by_(.dots = condtnVariableNames) %>%  #.dots needed when you have a variable containing multiple factor names
     do(  analyzeOneConditionDF(.,numItemsInStream,parameterBounds(),nReplicates=5) )
-
 ```
@@
 
-estimates<- df  %>% filter(subject=="AA") %>%
+estimates<- df  %>%  filter( !is.nan(targetSP) ) %>%
     group_by_(.dots = condtnVariableNames) %>%  #.dots needed when you have a variable containing multiple factor names

### 2018-08-17 E2/analysis/inspectHists.Rmd
about to build renaming columns into import

@@
 BJ has essentially zero efficacy and may have tended to report letters from beginning of stream in one condition.
-
+```{r }
 SsWhoLookWeird<-c("CY","AP","BJ")
-
+```
 We'll see whether the statistical comparison of mixture modeling to guessing distribution flags these as not being fit significantly better by mixture model.

### 2018-08-17 E2/analysis/mixModelE2data.Rmd
about to build renaming columns into import

@@
 ```{r}
-
 table(E2$targetSP, E2$targetSide, dnn=c("targetSP","targetSide"), useNA="ifany")
-
 ```
@@
 ```{r, echo=FALSE}
-
+#devtools::install_github('alexholcombe/mixRSVP',build_vignettes=TRUE)
 library(mixRSVP)
@@
 
-estimates<- df  %>%  filter( !is.nan(targetSP) ) %>%
+estimates<- E2  %>%  filter( !is.nan(targetSP) ) %>% filter(subject>"AF") %>%
     group_by_(.dots = condtnVariableNames) %>%  #.dots needed when you have a variable containing multiple factor names
@@
 Save the param estimates to a file.
-```{Save param estimates, echo=FALSE}
+```{r SaveParamEstimates, echo=FALSE}
 
-write.csv(estimates, file = "Results/backwards2E2_paramEstimates.csv")
-#saveRDS(estimates, file = "Data/backwards2E1_paramEstimates.rda", compress="bzip2")
+resultsPath<- file.path("Results")
+
+write.csv(estimates, file = file.path(resultsPath, "backwards2E2_paramEstimates.csv"))
+#saveRDS(estimates, file = file.path(resultsPath, "Results/backwards2E1_paramEstimates.rda"), compress="bzip2")
 

### 2018-08-17 E2/analysis/validateE2data.Rmd
about to build renaming columns into import

@@
 ---
-title: "importE2data"
+title: "validateE2data"
 author: "Alex Holcombe"
@@
 ---
+Optional validation (don't need to run this), loads raw data on its own and validates a few things
 
@@
 
+Sanity-check some data by graphing.
+
+Investigate whether pathological identicalness of inverted/canonical happens, that happened in E1 . Seems it doesn't
+
+```{r}
+dg<- E2 %>% filter(subject < "AC")
+
+library(ggplot2)
+g=ggplot(dg, aes(x=SPE)) + facet_grid(subject+orientation~stream) #,  scales="free_y")
+g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
+g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
+show(g)
+```
+
 Save data

### 2018-08-19 E2/analysis/importE2data.Rmd
progress

@@
 
+Rename some columns
+```{r }
+E<- data
+names(E)[names(E) == 'target'] <-'stream'
+names(E)[names(E) == 'condition'] <-'orientation'
+E$stream[ E$stream==1 ] <- "Left"
+E$stream[ E$stream==2 ] <- "Right"
+E$targetSide[ E$targetSide==0 ] <- "Both"
+E$targetSide[ E$targetSide==1 ] <- "Left"
+E$targetSide[ E$targetSide==2 ] <- "Right"
+
+#mutate condition to Orientation
+E$orientation[ E$orientation==1 ] <- "Canonical"
+E$orientation[ E$orientation==2 ] <- "Inverted"
+```
+
 Saving data.

### 2018-08-19 E2/analysis/mixModelE2data.Rmd
progress

@@
 numItemsInStream<- length( data[1,"letterSeq"]$letterSeq$`1` )  
-minSPE<- -17; maxSPE<- 17
+minSPE<- -17; maxSPE<- 17  #But it's 16 items in stream, so shouldn't it be -15, +15?
 
@@
 ```{r}
+#devtools::install_github('alexholcombe/mixRSVP',build_vignettes=TRUE)
 library(mixRSVP)
@@
 
-estimates<- E2  %>%  filter( !is.nan(targetSP) ) %>% filter(subject>"AF") %>%
-    group_by_(.dots = condtnVariableNames) %>%  #.dots needed when you have a variable containing multiple factor names
+estimates<- E2  %>%  filter( !is.na(targetSP) ) %>%
+    group_by_at(.vars = condtnVariableNames) %>%  
     do(  analyzeOneConditionDF(.,numItemsInStream,parameterBounds(),nReplicates=5) )
@@
 head(estimates)
+```
+
+Calc numObservations for each condition
+```{r }
+#Calc numObservations for each condition. This is needed  for scaling the fine-grained Gaussian
+#gaussianScaledforData needs to know.
+dfGroups<-  E2  %>%  filter( !is.na(targetSP) ) %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
+#add nPerCond back to parameter estimates
+estimates<- merge(estimates,dfGroups)
 ```

### 2018-08-19 E2/analysis/doEverything.Rmd
working on oneTarget plotting

---
title: "inspect histograms"
author: "Alex Holcombe"
date: "2/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Optional validation (don't need to run this), loads raw data on its own and validates a few things

```{r validate data, echo=FALSE, message=FALSE}
source('validateE2data.Rmd')
```

Load data

```{r load data, echo=FALSE, message=FALSE}
source('importE2data.Rmd')
```

inspect Hists for weird data, non-performing Ss

```{r inspectHists}

source('inspectHists.Rmd')
```

Mixture model data

```{r}
source('mixModelE2data.Rmd')
```

Plot hists with psychophysical curves added,
check that and likelihood ratio test for whether to exclude, 
and infer
```{r}
source('inspectHistsWithFits.Rmd')
```

### 2018-08-19 E2/analysis/histsWithFitsTwoTarget.Rmd
working on oneTarget plotting

---
title: "inspect histograms"
author: "Alex Holcombe"
date: "2/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)

```

Load data and estimates

```{r load data, echo=FALSE, message=FALSE}

#Import raw data
dataPath<- file.path("../Data/")
#Experiment was administered by MATLAB
#.mat file been preprocessed into melted long dataframe by importE1data.Rmd
data<- readRDS( file.path(dataPath, "backwards2E2_rawDataFromMAT.rda") )

resultsPath<- file.path("Results")
estimates<- read.csv(file = file.path(resultsPath,"/backwards2E2_paramEstimates.csv"))

```

Fix the variable names if not done

```{r, echo=FALSE, message=FALSE}
#Below should now be done by import
if ('target' %in% names(data)) {
  names(data)[names(data) == 'target'] <-'stream'
  names(data)[names(data) == 'condition'] <-'orientation'
  data$stream[ data$stream==1 ] <- "Left"
  data$stream[ data$stream==2 ] <- "Right"
}
if (is.numeric(data$targetSide)) {
  data$targetSide[ data$targetSide==0 ] <- "Both"
  data$targetSide[ data$targetSide==1 ] <- "Left"
  data$targetSide[ data$targetSide==2 ] <- "Right"
}
if (is.numeric(data$orientation)) {
  #mutate condition to Orientation
  data$orientation[ data$orientation==1 ] <- "Canonical"
  data$orientation[ data$orientation==2 ] <- "Inverted"
}
```

Add the parameter estimates to the tibble

```{r, echo=FALSE, message=FALSE, fig.height=360, fig.width=10}
#want fig.height of 10 per subject

condtnVariableNames <- c("subject","stream", "orientation","targetSide")

#Add R parameter estimates to data tibble
#To merge them, unfortunately have to convert to dataframe first because dplyr has no binding of tibbles.
# https://stackoverflow.com/questions/45643886/how-to-combine-tibbles-with-common-but-non-identical-columns#45644319
#Not sure what is lost as a result.
dd<- as_data_frame(data)
es<- as_data_frame(estimates)
#Looks like each subject has only 8 efficacy values somehow, it's because there was an extra
#variable, X
es$X <- NULL
dat<- merge(dd,es)
dg<-dat %>% filter( !is.na(targetSP) ) #excluding 1-target trials rows reporting stream on opposite side. Eventually, use those to evaluate swaps.
```

Set up the experiment-specific parameters

```{r detailsNeeded}

numItemsInStream<- length( dg$letterSeq[[1]] )  
minSPE<- -17; maxSPE<- 17
annotateIt<-TRUE

numSs<-length(unique(dg$subject))
cat( paste0('Total num Ss=',numSs) )
```

Calc numObservations for each condition
```{r }
#Calc numObservations for each condition. This is needed  for scaling the fine-grained Gaussian
#gaussianScaledforData needs to know.
if (!("nPerCond" %in% colnames(dg))) {
  dfGroups<-  dg  %>%  filter( !is.na(targetSP) ) %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
  #add nPerCond back to parameter estimates
  estimates<- merge(estimates,dfGroups)
}
```

Calculate curves for parameters, to plot on histograms.
```{r, echo=TRUE, message=FALSE, fig.height=360, fig.width=10}
#devtools::install_github('alexholcombe/mixRSVP',build_vignettes=TRUE)
library(mixRSVP)

curves<- dg %>% #filter(subject >= "AA" ) %>%
  group_by_at(.vars = condtnVariableNames) %>% 
  do(calc_curves_dataframe(.,minSPE,maxSPE,numItemsInStream))
#fails with on-missing arguments to min; returning Inf no non-missing arguments to max; returning -Inf 
#Error in createGuessingDistribution(minSPE, maxSPE, df$targetSP, numItemsInStream) :  minSPE must be less than or equal to maxSPE

#DEBUG
# dd<- dg %>% filter(subject=="AA", stream=="Left",targetSide=="Both",orientation=="Canonical")
# 
# createGuessingDistribution(minSPE,maxSPE,dd$targetSP,numItemsInStream)
# targetSP<- dd$targetSP
# maxTargetSP <- max(targetSP)
# minSPEthisData<- 1 - max(targetSP)
# maxSPEthisData<- numItemsInStream - min(targetSP)
# if (maxSPEthisData > maxSPE) stop("maxSPE must be at least",maxSPEthisData,"based on the values you passed me")
# if (minSPEthisData < minSPE) stop("minSPE must be no greater than",minSPEthisData,"based on the values you passed me")
# calc_curves_dataframe(dd,minSPE,maxSPE,numItemsInStream)

```

```{r }
# #Calc numObservations to each condition. This is needed only for scaling the fine-grained Gaussian
# #Calc the number of observations for each condition, because gaussianScaledforData needs to know.
# dfGroups<- dg %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
# #add nPerCond back to parameter estimates
# estimates<- merge(estimates,dfGroups)

grain<-.05
gaussFine<- estimates %>% group_by_at(.vars = condtnVariableNames) %>% do(
  gaussian_scaled_from_df(.,minSPE,maxSPE,grain) )
```

Plot sample histograms, make sure plotting working.

```{r singleHistPlot}
library(ggplot2)

dAC<- dplyr::filter(dg,subject=="AC")  
plotContinuousGaussian<- TRUE 
annotateIt<-TRUE
g<- plot_hist_with_fit(dAC,minSPE,maxSPE,d915$targetSP,numItemsInStream,
                        plotContinuousGaussian,annotateIt, FALSE)
g <- g + geom_vline(xintercept = 0)
g + annotate("text", x = 12, y = 25, label = "AC")
```

Show single plot with curves using scalable method to facet_grid
```{r singleHistPlotScalable, echo=FALSE}

dAC<- dplyr::filter(dg,subject=="AC")  
bothSides<- dAC %>% filter(targetSide=="Both")
bothSidesCurves <- curves %>% filter(subject=="AC") %>% filter(targetSide=="Both")

g=ggplot(bothSides, aes(x=SPE)) + 
        facet_grid(orientation~stream) #,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
g<-g +theme_apa() #+theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank())# hide all gridlines.
g<-g +xlim(minSPE,maxSPE)
sz=.3
#show(g)
g<-g+ geom_point(data=bothSidesCurves,aes(x=x,y=combinedFitFreq),color="chartreuse3",size=sz)

#g<-g+ geom_line(data=curves,aes(x=x,y=guessingFreq),color="yellow",size=sz)
#Discretized Gaussian
#g<-g+ geom_line(data=curves,aes(x=x,y=gaussianFreq),color="lightblue",size=sz)

#mixSig - whether mixture model statistically significantly better than guessing
bothSidesCurves <- dplyr::mutate(bothSidesCurves, mixSig = ifelse(pLRtest <= .05, TRUE, FALSE)) #annotate_fit uses this to color the p-value
g<- annotate_fit(g,bothSidesCurves) #assumes curvesDf includes efficacy,latency,precision
#Somehow the which mixSig (TRUE or FALSE) is red and which green is flipped relative to plot_hist_with_fit even though
#identical commands are used. I haven't been able to work out why.
#g<- g + scale_color_manual(values=c("red","forestgreen")) #already done in annotate_fit
show(g)
```

Make sure can plot without plot_hist_with_fit

Plot all Ss
```{r, fig.height=200, fig.width=10}

bothSides<- dg %>% filter(targetSide=="Both")
bothSidesCurves <- curves %>% filter(targetSide=="Both")

g=ggplot(bothSides, aes(x=SPE)) + 
        facet_grid(subject+orientation~stream) #,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
g<-g +theme_apa() #+theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank())# hide all gridlines.
g<-g +xlim(minSPE,maxSPE)
sz=.3
#show(g)
g<-g+ geom_point(data=bothSidesCurves,aes(x=x,y=combinedFitFreq),color="chartreuse3",size=sz)
#g<-g+ geom_line(data=curves,aes(x=x,y=guessingFreq),color="yellow",size=sz)
#Discretized Gaussian
#g<-g+ geom_line(data=curves,aes(x=x,y=gaussianFreq),color="lightblue",size=sz)

#mixSig - whether mixture model statistically significantly better than guessing
bothSidesCurves <- dplyr::mutate(bothSidesCurves, mixSig = ifelse(pLRtest <= .05, TRUE, FALSE)) #annotate_fit uses this to color the p-value
g<- annotate_fit(g,bothSidesCurves) #assumes curvesDf includes efficacy,latency,precision
#Somehow the which mixSig (TRUE or FALSE) is red and which green is flipped relative to plot_hist_with_fit even though
#identical commands are used. I haven't been able to work out why.
#g<- g + scale_color_manual(values=c("red","forestgreen")) #already done in annotate_fit
show(g)

```

CY is bimodal but not outrageously so.

Subject AP looks to both have close to zero efficacy in some conditions and disregards instructions, reporting mostly after the cue.

BJ has essentially zero efficacy and may have tended to report letters from beginning of stream in one condition.
```{r }
SsWhoLookWeird<-c("CY","AP","BJ")
```
We'll see whether the statistical comparison of mixture modeling to guessing distribution flags these as not being fit significantly better by mixture model.

AP, BJ does flunk likelihood test! CY passes but data not horrifyingly bad.
Exclude by likelihood ratio test

Participant AI, BB suspicion of  attending to one side, let's see which people fail single-target test.

```{r singleTarget}
oneSide<- dg %>% filter(targetSide!="Both")
oneSideCurves <- curves %>% filter(targetSide!="Both")

g=ggplot(oneSide %>% filter(subject=="AA"), aes(x=SPE)) + 
        facet_grid(subject+orientation~stream) #,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
show(g)
g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
g<-g +theme_apa() #+theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank())# hide all gridlines.
g<-g +xlim(minSPE,maxSPE)
sz=.3
show(g)
g<-g+ geom_point(data=oneSideCurves,aes(x=x,y=combinedFitFreq),color="chartreuse3",size=sz)
#g<-g+ geom_line(data=curves,aes(x=x,y=guessingFreq),color="yellow",size=sz)
#Discretized Gaussian
#g<-g+ geom_line(data=curves,aes(x=x,y=gaussianFreq),color="lightblue",size=sz)

#mixSig - whether mixture model statistically significantly better than guessing
bothSidesCurves <- dplyr::mutate(oneSideCurves, mixSig = ifelse(pLRtest <= .05, TRUE, FALSE)) #annotate_fit uses this to color the p-value
g<- annotate_fit(g,oneSideCurves) #assumes curvesDf includes efficacy,latency,precision
#Somehow the which mixSig (TRUE or FALSE) is red and which green is flipped relative to plot_hist_with_fit even though
#identical commands are used. I haven't been able to work out why.
#g<- g + scale_color_manual(values=c("red","forestgreen")) #already done in annotate_fit
show(g)
```

### 2018-08-23 E2/analysis/histsWithFitsOneTarget.Rmd
excluding

---
title: "one target"
author: "Alex Holcombe"
date: "2/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)

```

Load data and estimates

```{r load data, echo=FALSE, message=FALSE}

#Import raw data
dataPath<- file.path("../Data/")
#Experiment was administered by MATLAB
#.mat file been preprocessed into melted long dataframe by importE1data.Rmd
data<- readRDS( file.path(dataPath, "backwards2E2_rawDataFromMAT.rda") )

resultsPath<- file.path("Results")
estimates<- read.csv(file = file.path(resultsPath,"/backwards2E2_paramEstimates.csv"))

```

Fix the variable names if not done

```{r, echo=FALSE, message=FALSE}
#Below should now be done by import
if ('target' %in% names(data)) {
  names(data)[names(data) == 'target'] <-'stream'
  names(data)[names(data) == 'condition'] <-'orientation'
  data$stream[ data$stream==1 ] <- "Left"
  data$stream[ data$stream==2 ] <- "Right"
}
if (is.numeric(data$targetSide)) {
  data$targetSide[ data$targetSide==0 ] <- "Both"
  data$targetSide[ data$targetSide==1 ] <- "Left"
  data$targetSide[ data$targetSide==2 ] <- "Right"
}
if (is.numeric(data$orientation)) {
  #mutate condition to Orientation
  data$orientation[ data$orientation==1 ] <- "Canonical"
  data$orientation[ data$orientation==2 ] <- "Inverted"
}
```

Add the parameter estimates to the tibble

```{r, echo=FALSE, message=FALSE, fig.height=360, fig.width=10}
#want fig.height of 10 per subject

condtnVariableNames <- c("subject","stream", "orientation","targetSide")

#Add R parameter estimates to data tibble
#To merge them, unfortunately have to convert to dataframe first because dplyr has no binding of tibbles.
# https://stackoverflow.com/questions/45643886/how-to-combine-tibbles-with-common-but-non-identical-columns#45644319
#Not sure what is lost as a result.
dd<- as_data_frame(data)
es<- as_data_frame(estimates)
#Looks like each subject has only 8 efficacy values somehow, it's because there was an extra
#variable, X
es$X <- NULL
dat<- merge(dd,es)
dg<-dat %>% filter( !is.na(targetSP) ) #excluding 1-target trials rows reporting stream on opposite side. Eventually, use those to evaluate swaps.
```

Set up the experiment-specific parameters

```{r detailsNeeded}

numItemsInStream<- length( dg$letterSeq[[1]] )  
minSPE<- -17; maxSPE<- 17
annotateIt<-TRUE

numSs<-length(unique(dg$subject))
cat( paste0('Total num Ss=',numSs) )
```

Calc numObservations for each condition
```{r }
#Calc numObservations for each condition. This is needed  for scaling the fine-grained Gaussian
#gaussianScaledforData needs to know.
if (!("nPerCond" %in% colnames(dg))) {
  dfGroups<-  dg  %>%  filter( !is.na(targetSP) ) %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
  #add nPerCond back to parameter estimates
  estimates<- merge(estimates,dfGroups)
}
```

Calculate curves for parameters, to plot on histograms.
```{r, echo=TRUE, message=FALSE, fig.height=360, fig.width=10}
#devtools::install_github('alexholcombe/mixRSVP',build_vignettes=TRUE)
library(mixRSVP)
oneTarget<- dg %>% filter(targetSide!="Both")
curvesOneTarget<- dg %>% filter(targetSide!="Both") %>%
  group_by_at(.vars = condtnVariableNames) %>% 
  do(calc_curves_dataframe(.,minSPE,maxSPE,numItemsInStream))
#fails with on-missing arguments to min; returning Inf no non-missing arguments to max; returning -Inf 
#Error in createGuessingDistribution(minSPE, maxSPE, df$targetSP, numItemsInStream) :  minSPE must be less than or equal to maxSPE

#DEBUG
# dd<- dg %>% filter(subject=="AA", stream=="Left",targetSide=="Both",orientation=="Canonical")
# 
# createGuessingDistribution(minSPE,maxSPE,dd$targetSP,numItemsInStream)
# targetSP<- dd$targetSP
# maxTargetSP <- max(targetSP)
# minSPEthisData<- 1 - max(targetSP)
# maxSPEthisData<- numItemsInStream - min(targetSP)
# if (maxSPEthisData > maxSPE) stop("maxSPE must be at least",maxSPEthisData,"based on the values you passed me")
# if (minSPEthisData < minSPE) stop("minSPE must be no greater than",minSPEthisData,"based on the values you passed me")
# calc_curves_dataframe(dd,minSPE,maxSPE,numItemsInStream)

```

```{r }
# #Calc numObservations to each condition. This is needed only for scaling the fine-grained Gaussian
# #Calc the number of observations for each condition, because gaussianScaledforData needs to know.
# dfGroups<- dg %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
# #add nPerCond back to parameter estimates
# estimates<- merge(estimates,dfGroups)

grain<-.05
gaussFine<- estimates %>%  filter(targetSide!="Both") %>% group_by_at(.vars = condtnVariableNames) %>% do(
  gaussian_scaled_from_df(.,minSPE,maxSPE,grain) )
```

Plot sample histograms, make sure plotting working.

```{r singleHistPlot}
library(ggplot2)

dAC<- dplyr::filter(oneTarget,subject=="AC")  
plotContinuousGaussian<- TRUE 
annotateIt<-TRUE
g<- plot_hist_with_fit(dAC,minSPE,maxSPE,d915$targetSP,numItemsInStream,
                        plotContinuousGaussian,annotateIt, FALSE)
g <- g + geom_vline(xintercept = 0)
g + annotate("text", x = 12, y = 25, label = "AC")
```

Show single plot with curves using scalable method to facet_grid.
In other words, make sure can plot without plot_hist_with_fit

```{r singleHistPlotScalable, echo=FALSE}

dAC<- dplyr::filter(oneTarget,subject=="AC")  
curvesThis <- curvesOneTarget %>% filter(subject=="AC")

g=ggplot(dAC, aes(x=SPE)) + 
        facet_grid(orientation~stream) #,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
g<-g +theme_apa() #+theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank())# hide all gridlines.
g<-g +xlim(minSPE,maxSPE)
sz=.3
#show(g)
g<-g+ geom_point(data=curvesThis,aes(x=x,y=combinedFitFreq),color="chartreuse3",size=sz)

#g<-g+ geom_line(data=curves,aes(x=x,y=guessingFreq),color="yellow",size=sz)
#Discretized Gaussian
#g<-g+ geom_line(data=curves,aes(x=x,y=gaussianFreq),color="lightblue",size=sz)

#mixSig - whether mixture model statistically significantly better than guessing
curvesThis <- dplyr::mutate(curvesThis, mixSig = ifelse(pLRtest <= .05, TRUE, FALSE)) #annotate_fit uses this to color the p-value
g<- annotate_fit(g,curvesThis) #assumes curvesDf includes efficacy,latency,precision
#Somehow the which mixSig (TRUE or FALSE) is red and which green is flipped relative to plot_hist_with_fit even though
#identical commands are used. I haven't been able to work out why.
#g<- g + scale_color_manual(values=c("red","forestgreen")) #already done in annotate_fit
show(g)
```

Plot all Ss
```{r, fig.height=200, fig.width=10}

g=ggplot(oneTarget, aes(x=SPE)) + 
        facet_grid(subject+orientation~stream) #,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
g<-g +theme_apa() #+theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank())# hide all gridlines.
g<-g +xlim(minSPE,maxSPE)
sz=.3
#show(g)
g<-g+ geom_point(data=curvesOneTarget,aes(x=x,y=combinedFitFreq),color="chartreuse3",size=sz)
#g<-g+ geom_line(data=curves,aes(x=x,y=guessingFreq),color="yellow",size=sz)
#Discretized Gaussian
#g<-g+ geom_line(data=curves,aes(x=x,y=gaussianFreq),color="lightblue",size=sz)

#mixSig - whether mixture model statistically significantly better than guessing
curvesOneTarget <- dplyr::mutate(curvesOneTarget, mixSig = ifelse(pLRtest <= .05, TRUE, FALSE)) #annotate_fit uses this to color the p-value
g<- annotate_fit(g,curvesOneTarget) #assumes curvesDf includes efficacy,latency,precision
#Somehow the which mixSig (TRUE or FALSE) is red and which green is flipped relative to plot_hist_with_fit even though
#identical commands are used. I haven't been able to work out why.
#g<- g + scale_color_manual(values=c("red","forestgreen")) #already done in annotate_fit
show(g)

```

```{r investigateDifficultyCollapsing}
#see preservingNonNumericWhenSummarizing.R
tw<- as_tibble(twoTarget)
tw<- tw %>% select(-targetStreamOrientations,-warnings,-letterSeq)
ty<- tw %>% group_by(subject,stream,orientation) %>% summarise_all(funs(if_else(is.numeric(.), mean(.), first(.))))

tt<- as_tibble(twoTarget) 
zb<- tt %>% group_by(subject,stream,orientation) %>% select(-targetStreamOrientations,-letterSeq) %>%
  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })

#The below works, but remember I've eliminated the list and string columns
#zz<- tw %>% group_by(subject,stream,orientation) %>%
#  summarise_all(function(x){ifelse(is.numeric(x),mean(x),last(x))})

tt<- as_tibble(twoTarget)
za<- tt %>% group_by(subject,stream,orientation) %>%
  summarise_all(function(x){ifelse(is.numeric(x),mean(x),last(x))})

tz<-tw
tz$targetSide <- as.factor(tz$targetSide)
tz %>% group_by(subject,stream,orientation) %>% summarise_all(funs(if_else(is.numeric(.), mean(.), first(.))))

tw %>% group_by(subject,stream,orientation) %>% summarise_at("targetSide",first)

ty<- tw %>% group_by(subject,stream,orientation) %>% summarise_all(funs(if_else(is.numeric(.), mean(.), -999.9)))
```

BE has e=.47 on left, .62 on right
BJ is just near chance and fails likelihood ratio test for 1-target, like did for 2-targets.
BO did badly for canonical on left, but passed lr test.
CF did badly for inverted on left, but passed lr test.
CL did pretty badly but got lucky, passed.
AP flunked one of the lr tests which is good because flunked some in two-target too.
So, lr test seems to be working well enough.

## Exclusions

Exclude by likelihood ratio test and efficacy.

For each subject, calculate how many two-target streams they failed. If 3 or 4 of 4, exclude (didn't think to mention 1-target condition in prereg).

Also check whether average efficacy is below 10% in both streams in either the upright or the reversed orientation.

```{r exclusions}

twoTarget<- dg %>% filter(targetSide=="Both")

AA<- twoTarget %>% filter(subject=="AA")
#Function for one subject, calculating how many stream conditions failed the test
calcExclude<- function(oneSub) {
  oneSub %>% group_by(stream,orientation)
}
#first, collapse over trials
#https://stackoverflow.com/questions/51978519/summarise-numeric-columns-return-last-value-of-non-numeric
tw<- twoTarget %>% group_by(subject,stream,orientation) %>% select(-targetStreamOrientations,-letterSeq) %>% #can't include list columns
  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })

#for each subject, count pLRtest instances >.05 and efficay <.1
B<-tw %>% ungroup %>% group_by(subject) %>% summarise(LRtestFails = sum(pLRtest>.05), effFails = sum(efficacy<.1))                                                              
#B %>% print(n=Inf)
exclude <- B %>% filter(LRtestFails>=3, effFails >

# efficacy is below 10% in both streams in either the upright or the reversed orientation.
# To calculate that, need to group by subject and orientation
C<-tw %>% ungroup %>% group_by(subject,orientation) %>% summarise(bothEffsFailed = sum(efficacy<.1)==2)
#Find which subjects, if any, has a bothEffsFailed==TRUE
lowEff<- C %>% filter(bothEffsFailed==TRUE)
lowEff <- unique(lowEff$subject)

A<- tw %>% filter(subject=="AA")
B<-A %>% ungroup %>% summarise(LRtestFails = sum(pLRtest<.05), effFails = sum(efficacy<.1))

twoTarget %>% select(-targetStreamOrientations,-warnings,-letterSeq) %>% group_by(stream,orientation) %>% summarise_all(funs(if_else(is.numeric(.), mean(.), 3))

two<- twoTarget %>% group_by(stream,orientation) %>% summarise_all(funs(if_else(!is.list(.), mean(.), last(.))))

  summarise_all(funs(mean,sd))
A <- two %>% filter(subject=="AA")

curvesOneTarget<- twoTarget %>%
  group_by(subject) %>% 
  do(calc_curves_dataframe(.,minSPE,maxSPE,numItemsInStream))

```

Participant AI, BB suspicion of  attending to one side from twoTargets condition.

### 2018-08-23 manuscript/manuscript.Rmd
excluding

---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

### Exclusions

In the prereg, we wrote: "
Participants will be excluded if their average efficacy is below 10% in both streams in either the upright or the reversed orientation), or if the likelihood ratio test testing whether the data is fit better by the mixture model than by the guessing distribution fails (is not less than p<.05) for three or four of the four conditions. Because the Gaussian component of the mixture model is constrained to only allow latencies between -4 and 4, one effect of this is to eliminate participants who consistently do not report a letter near the time of the ring, instead for example reporting one of the last few items in the stream, which the data of past experiments indicates that rare participants do."

## Material

## Procedure

Preregistration: https://osf.io/r327j/
http://htmlpreview.github.io/?https://github.com/alexholcombe/backwardsLtrs2ndPaper/blob/master/E2/analysis/bayesianAnalysisForPrereg.html , which used our custom prior (details in statistical models section) and found that 75 participants would yield a decent Bayes Factor (about 10 or more) for the effect size we are expecting.

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

### 2018-08-23 E2/analysis/dev/preservingNonNumericWhenSummarising.R

@@ -1,0 +1,47 @@
+#Original idea from
+#https://stackoverflow.com/a/45802176/302378
+
+#My question
+#https://stackoverflow.com/questions/51978519/summarise-numeric-columns-return-last-value-of-non-numeric/51980086#51980086
+
+# dataset('iris')
+# iris %>% group_by(Sepal.Length) %>% summarise_all(funs(if_else(is.numeric(.), mean(.), first(.))))
+# iris %>% select(-Species) %>% group_by(Sepal.Length) %>% 
+#   summarise_all(funs(if_else(is.numeric(.), mean(.), first(.))))
+
+
+set.seed(1234)
+category <- (c('A','A','E','E','B','B','C'))
+date <- seq(as.Date("2017-01-01"), by = "month", length.out = 7)
+value1 <- sample(seq(from = 91, to = 97, by = 1))
+dt <- data.frame(category, date, value1)
+dt<- as_tibble(dt)
+#works
+dt2<- dt %>% select(-marsupial) %>%
+  group_by(category) %>%
+  summarise_all(funs(if_else(is.numeric(.), mean(.), last(.))))
+print(dt2)
+
+marsupial <- c("quoll","phascogale",'triok','opossum','antechinus','bandicoot','Fat-tailed dunnart')
+dt$marsupial <- marsupial
+dt3<- dt %>% #doesn't work
+  group_by(category) %>%
+  summarise_all(funs(if_else(is.numeric(.), mean(.), last(.))))
+
+#WORKS except fucks up date, turns it into double. Also doesn't preserve columns that are lists
+dt4<- dt %>% group_by(category) %>%
+  summarise_all(function(x){ifelse(is.numeric(x),mean(x),last(x))}) 
+
+#Below works and the author said works on list columns, but doesn't, gives errors like
+# Error in summarise_impl(.data, dots) : 
+# Column `targetStreamOrientations` must be length 1 (a summary value), not 16
+dt5<- dt %>% group_by(category) %>%
+  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })
+#Below gives error related to date
+dt5<- dt %>% group_by(category) %>%
+  summarise_all(function(x){ if (is.numeric(.)) mean(.) else last(.) })
+
+dt6<- dt %>% group_by(category) %>%
+  summarise_if(is.numeric,mean)
+
+y = if (3==5) 4 else 5

### 2018-08-24 E2/analysis/histsWithFitsOneTarget.Rmd
exclusions

@@
 
-AA<- twoTarget %>% filter(subject=="AA")
-#Function for one subject, calculating how many stream conditions failed the test
-calcExclude<- function(oneSub) {
-  oneSub %>% group_by(stream,orientation)
-}
 #first, collapse over trials
@@
 #B %>% print(n=Inf)
-exclude <- B %>% filter(LRtestFails>=3, effFails >
-
+LRtestLosers <- B %>% filter(LRtestFails>=3)
+LRtestLosers<- unique(LRtestLosers$subject)
 # efficacy is below 10% in both streams in either the upright or the reversed orientation.
@@
 lowEff<- C %>% filter(bothEffsFailed==TRUE)
-lowEff <- unique(lowEff$subject)
+lowEffs <- unique(lowEff$subject)
+```
+
+I suppose there are 4 one-target conditions also, upright/reversed x left/right
+```{r }
+#Calculate exclusions based on one-target performance
+#oneTarget<- dg %>% filter(targetSide!="Both")
+one<- oneTarget %>% group_by(subject,targetSide,orientation) %>% 
+  select(-targetStreamOrientations,-letterSeq) %>% #can't include list columns
+  select(-trial,-targetSP,-respSP,-respOrder) %>% #clear detritus so easier to see
+  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })
+
+#Should probably look at the difference in efficacy between the left and the right
+one %>% group_by(subject,targetSide) %>%   function(x) { x$effDiff = x %>% filter(stream=="Left")   -    x %>% filter(stream=="Right") }
+
+one %>% group_by(subject,targetSide) %>% summarise_at("efficacy",
+       function(x) { x %>% filter(stream=="Left")   -    x %>% filter(stream=="Right") }  )
+
+  mutate(effDiff = 
+       function(x) { x %>% filter(stream=="Left")$efficacy   -    x %>% filter(stream=="Right")$efficacy }  )
+
+  summarise(LRtestFails = sum(pLRtest>.05), effFails = sum(efficacy<.1))
+
+```
+
+Make master list of Ss to exclude also from one-target condition??
+
+Actually exclude the losers
+```{r exclude, echo=FALSE}
+
+afterExclusions<- tw %>% filter(  !(subject %in% lowEffs)   )
+
+```
 

### 2018-08-24 E2/analysis/histsWithFitsOneTarget.Rmd
right

@@
 
-```{r investigateDifficultyCollapsing}
-#see preservingNonNumericWhenSummarizing.R
-tw<- as_tibble(twoTarget)
-tw<- tw %>% select(-targetStreamOrientations,-warnings,-letterSeq)
-ty<- tw %>% group_by(subject,stream,orientation) %>% summarise_all(funs(if_else(is.numeric(.), mean(.), first(.))))
-
-tt<- as_tibble(twoTarget) 
-zb<- tt %>% group_by(subject,stream,orientation) %>% select(-targetStreamOrientations,-letterSeq) %>%
-  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })
-
-#The below works, but remember I've eliminated the list and string columns
-#zz<- tw %>% group_by(subject,stream,orientation) %>%
-#  summarise_all(function(x){ifelse(is.numeric(x),mean(x),last(x))})
-
-tt<- as_tibble(twoTarget)
-za<- tt %>% group_by(subject,stream,orientation) %>%
-  summarise_all(function(x){ifelse(is.numeric(x),mean(x),last(x))})
-
-tz<-tw
-tz$targetSide <- as.factor(tz$targetSide)
-tz %>% group_by(subject,stream,orientation) %>% summarise_all(funs(if_else(is.numeric(.), mean(.), first(.))))
-
-tw %>% group_by(subject,stream,orientation) %>% summarise_at("targetSide",first)
-
-ty<- tw %>% group_by(subject,stream,orientation) %>% summarise_all(funs(if_else(is.numeric(.), mean(.), -999.9)))
-```
-
 BE has e=.47 on left, .62 on right
@@
 
-I suppose there are 4 one-target conditions also, upright/reversed x left/right
+There are 4 one-target conditions also, upright/reversed x left/right.
+Will exclude subject if failed likelihood test for both left and right in either condition.
+
+Will also exclude if difference in efficacy left-right exceeds?
 ```{r }
@@
 
-#Should probably look at the difference in efficacy between the left and the right
-one %>% group_by(subject,targetSide) %>%   function(x) { x$effDiff = x %>% filter(stream=="Left")   -    x %>% filter(stream=="Right") }
+#Look at the difference in efficacy between the left and the right
+effDiff <- function(df) {
+  #dgi<<-df
+  #STOP
+  effLeft<- (df %>% filter(targetSide=="Left"))$efficacy 
+  effRight<- (df %>% filter(targetSide=="Right"))$efficacy
+  df$effDiff<- effLeft - effRight
+  df$efficacy<- mean(c(effLeft,effRight))
+  df$effDiff<- effLeft - effRight
 
-one %>% group_by(subject,targetSide) %>% summarise_at("efficacy",
-       function(x) { x %>% filter(stream=="Left")   -    x %>% filter(stream=="Right") }  )
+  failedBothLR <- ((df %>% filter(targetSide=="Left"))$pLRtest >= .05) &
+    ((df %>% filter(targetSide=="Right"))$pLRtest >= .05)
+  df$failedBothLR <- failedBothLR
 
-  mutate(effDiff = 
-       function(x) { x %>% filter(stream=="Left")$efficacy   -    x %>% filter(stream=="Right")$efficacy }  )
+  return(df[1,]) #return first row, so it does get collapsed
+}
 
-  summarise(LRtestFails = sum(pLRtest>.05), effFails = sum(efficacy<.1))
+oneDiff<- one %>% group_by(subject,orientation) %>% do( effDiff(.) )
+oneDiff<- oneDiff %>% select(subject,targetSide,orientation,efficacy,effDiff,pLRtest,failedBothLR)
+oneDiff
+
+oneTargetLosers <- oneDiff %>% filter(failedBothLR == TRUE, effDiff > .4)
+
+```
+
+Make scatterplot of efficacy for left vs. right single target
+
+```{r}
+#To do that, do I need to unmelt?
+ggplot(one, aes(x=
+```
+
+Make scatterplot of efficacy difference between left and right single target on horizontal axis and dual-target difference on vertical axis.
+
+```{r}
 

### 2018-08-29 E2/analysis/doEverything.Rmd
this is going well

@@
 ---
-title: "inspect histograms"
+title: "Do everything"
 author: "Alex Holcombe"
-date: "2/26/2018"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output: html_document
@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
+The idea of this document is to have a single file that when knit does everything, [calling subsidiary Rmd documents](https://stackoverflow.com/questions/41962434/source-code-from-rmd-file-within-another-rmd)
+
+But the problem is, the importE2data.Rmd throws a bizarre error when knit, but not when run!
+Moreover, importE2data takes a very long time to run.
+So in practice, the better thing to do is to run the below files one by one.
 
@@
 
-```{r validate data, echo=FALSE, message=FALSE}
-source('validateE2data.Rmd')
+```{r validate, child = 'validateE2data.Rmd'}
 ```
@@
 
-```{r load data, echo=FALSE, message=FALSE}
-source('importE2data.Rmd')
+```{r import, child = 'importE2data.Rmd'}
 ```
 
-inspect Hists for weird data, non-performing Ss
+Inspect Hists for weird data, non-performing Ss.  This is prior to mixture modeling. Will do again after mixture modeling.
 
-```{r inspectHists}
-
-source('inspectHists.Rmd')
+```{r inspectHists, child = 'inspectHists.Rmd'}
 ```
@@
 
-```{r}
-source('mixModelE2data.Rmd')
+```{r inspectHists, child = 'mixModelE2data.Rmd'}
 ```
 
-Plot hists with psychophysical curves added,
+Plot two-target hists with psychophysical curves added,
 check that and likelihood ratio test for whether to exclude, 
-and infer
-```{r}
-source('inspectHistsWithFits.Rmd')
+```{r, child= 'histsWithFitsTwoTarget.Rmd'}
 ```
+
+Plot one-target hists with psychophysical curves added,
+
+```{r, child= 'histsWithFitsOneTarget.Rmd'}
+```
+
+Do exclusions, add column indicating whether subject excluded to dataframe, save that.
+Should also add handedness quotient from handednessQuotientsP2E2 but there are so few left-handers it wouldn't make a difference. Need to pool across papers to test handedness issue.
+```{r, child ='exclusions.Rmd')}
+```
+
+Look at incidence of location swaps.
+
+What will I do about orientation? I can see how often they get orientation wrong, but that by itself is not so interesting.
+I can look at the serial position correlogram of orientation and letter identity and I expect to see quite a strong correlation, contrary to what was found for two letters but something it would make sense to report at same time as Cheryl's finding.

### 2018-08-29 E2/analysis/exclusions.Rmd
this is going well

--- 2018-08-24 E2/analysis/histsWithFitsOneTarget.Rmd
@@
 ---
-title: "one target"
+title: "Exclusions"
 author: "Alex Holcombe"
@@
 
-Set up the experiment-specific parameters
-
-```{r detailsNeeded}
-
-numItemsInStream<- length( dg$letterSeq[[1]] )  
-minSPE<- -17; maxSPE<- 17
-annotateIt<-TRUE
-
-numSs<-length(unique(dg$subject))
-cat( paste0('Total num Ss=',numSs) )
-```
-
-Calc numObservations for each condition
-```{r }
-#Calc numObservations for each condition. This is needed  for scaling the fine-grained Gaussian
-#gaussianScaledforData needs to know.
-if (!("nPerCond" %in% colnames(dg))) {
-  dfGroups<-  dg  %>%  filter( !is.na(targetSP) ) %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
-  #add nPerCond back to parameter estimates
-  estimates<- merge(estimates,dfGroups)
-}
-```
-
-Calculate curves for parameters, to plot on histograms.
-```{r, echo=TRUE, message=FALSE, fig.height=360, fig.width=10}
-#devtools::install_github('alexholcombe/mixRSVP',build_vignettes=TRUE)
-library(mixRSVP)
-oneTarget<- dg %>% filter(targetSide!="Both")
-curvesOneTarget<- dg %>% filter(targetSide!="Both") %>%
-  group_by_at(.vars = condtnVariableNames) %>% 
-  do(calc_curves_dataframe(.,minSPE,maxSPE,numItemsInStream))
-#fails with on-missing arguments to min; returning Inf no non-missing arguments to max; returning -Inf 
-#Error in createGuessingDistribution(minSPE, maxSPE, df$targetSP, numItemsInStream) :  minSPE must be less than or equal to maxSPE
-
-#DEBUG
-# dd<- dg %>% filter(subject=="AA", stream=="Left",targetSide=="Both",orientation=="Canonical")
-# 
-# createGuessingDistribution(minSPE,maxSPE,dd$targetSP,numItemsInStream)
-# targetSP<- dd$targetSP
-# maxTargetSP <- max(targetSP)
-# minSPEthisData<- 1 - max(targetSP)
-# maxSPEthisData<- numItemsInStream - min(targetSP)
-# if (maxSPEthisData > maxSPE) stop("maxSPE must be at least",maxSPEthisData,"based on the values you passed me")
-# if (minSPEthisData < minSPE) stop("minSPE must be no greater than",minSPEthisData,"based on the values you passed me")
-# calc_curves_dataframe(dd,minSPE,maxSPE,numItemsInStream)
-
-```
-
-```{r }
-# #Calc numObservations to each condition. This is needed only for scaling the fine-grained Gaussian
-# #Calc the number of observations for each condition, because gaussianScaledforData needs to know.
-# dfGroups<- dg %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
-# #add nPerCond back to parameter estimates
-# estimates<- merge(estimates,dfGroups)
-
-grain<-.05
-gaussFine<- estimates %>%  filter(targetSide!="Both") %>% group_by_at(.vars = condtnVariableNames) %>% do(
-  gaussian_scaled_from_df(.,minSPE,maxSPE,grain) )
-```
-
-Plot sample histograms, make sure plotting working.
-
-```{r singleHistPlot}
-library(ggplot2)
-
-dAC<- dplyr::filter(oneTarget,subject=="AC")  
-plotContinuousGaussian<- TRUE 
-annotateIt<-TRUE
-g<- plot_hist_with_fit(dAC,minSPE,maxSPE,d915$targetSP,numItemsInStream,
-                        plotContinuousGaussian,annotateIt, FALSE)
-g <- g + geom_vline(xintercept = 0)
-g + annotate("text", x = 12, y = 25, label = "AC")
-```
-
-Show single plot with curves using scalable method to facet_grid.
-In other words, make sure can plot without plot_hist_with_fit
-
-```{r singleHistPlotScalable, echo=FALSE}
-
-dAC<- dplyr::filter(oneTarget,subject=="AC")  
-curvesThis <- curvesOneTarget %>% filter(subject=="AC")
-
-g=ggplot(dAC, aes(x=SPE)) + 
-        facet_grid(orientation~stream) #,  scales="free_y")
-g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
-g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
-g<-g +theme_apa() #+theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank())# hide all gridlines.
-g<-g +xlim(minSPE,maxSPE)
-sz=.3
-#show(g)
-g<-g+ geom_point(data=curvesThis,aes(x=x,y=combinedFitFreq),color="chartreuse3",size=sz)
-
-#g<-g+ geom_line(data=curves,aes(x=x,y=guessingFreq),color="yellow",size=sz)
-#Discretized Gaussian
-#g<-g+ geom_line(data=curves,aes(x=x,y=gaussianFreq),color="lightblue",size=sz)
-
-#mixSig - whether mixture model statistically significantly better than guessing
-curvesThis <- dplyr::mutate(curvesThis, mixSig = ifelse(pLRtest <= .05, TRUE, FALSE)) #annotate_fit uses this to color the p-value
-g<- annotate_fit(g,curvesThis) #assumes curvesDf includes efficacy,latency,precision
-#Somehow the which mixSig (TRUE or FALSE) is red and which green is flipped relative to plot_hist_with_fit even though
-#identical commands are used. I haven't been able to work out why.
-#g<- g + scale_color_manual(values=c("red","forestgreen")) #already done in annotate_fit
-show(g)
-```
-
-Plot all Ss
-```{r, fig.height=200, fig.width=10}
-
-g=ggplot(oneTarget, aes(x=SPE)) + 
-        facet_grid(subject+orientation~stream) #,  scales="free_y")
-g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
-g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
-g<-g +theme_apa() #+theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank())# hide all gridlines.
-g<-g +xlim(minSPE,maxSPE)
-sz=.3
-#show(g)
-g<-g+ geom_point(data=curvesOneTarget,aes(x=x,y=combinedFitFreq),color="chartreuse3",size=sz)
-#g<-g+ geom_line(data=curves,aes(x=x,y=guessingFreq),color="yellow",size=sz)
-#Discretized Gaussian
-#g<-g+ geom_line(data=curves,aes(x=x,y=gaussianFreq),color="lightblue",size=sz)
-
-#mixSig - whether mixture model statistically significantly better than guessing
-curvesOneTarget <- dplyr::mutate(curvesOneTarget, mixSig = ifelse(pLRtest <= .05, TRUE, FALSE)) #annotate_fit uses this to color the p-value
-g<- annotate_fit(g,curvesOneTarget) #assumes curvesDf includes efficacy,latency,precision
-#Somehow the which mixSig (TRUE or FALSE) is red and which green is flipped relative to plot_hist_with_fit even though
-#identical commands are used. I haven't been able to work out why.
-#g<- g + scale_color_manual(values=c("red","forestgreen")) #already done in annotate_fit
-show(g)
-
-```
+One-target:
 
@@
 
-## Exclusions
+BG, BJ, and AP have sigma near 4, clear outliers.
+
+## Exclusions based on two-target conditions
+
+Notes from histsWithFitsTwoTarget.Rmd:
+AP, especially BJ does flunk likelihood test! CY passes but data not horrifyingly bad.
+Exclude by likelihood-ratio test.
+Participant AI, BB suspicion of  attending to one side from twoTargets condition.
 
@@
 lowEff<- C %>% filter(bothEffsFailed==TRUE)
-lowEffs <- unique(lowEff$subject)
+lowEffLosers <- unique(lowEff$subject)
 ```
+
+## Exclusions based on one-target conditions
 
@@
 #Calculate exclusions based on one-target performance
-#oneTarget<- dg %>% filter(targetSide!="Both")
+oneTarget<- dg %>% filter(targetSide!="Both")
 one<- oneTarget %>% group_by(subject,targetSide,orientation) %>% 
@@
   df$effDiff<- effLeft - effRight
+  df$effLeft <- effLeft; df$effRight <- effRight
+
   df$efficacy<- mean(c(effLeft,effRight))
@@
 oneDiff<- one %>% group_by(subject,orientation) %>% do( effDiff(.) )
-oneDiff<- oneDiff %>% select(subject,targetSide,orientation,efficacy,effDiff,pLRtest,failedBothLR)
-oneDiff
+oneDiff %>% select(subject,targetSide,orientation,efficacy,effDiff,pLRtest,failedBothLR)
 
 oneTargetLosers <- oneDiff %>% filter(failedBothLR == TRUE, effDiff > .4)
-
+oneTargetLosers<- unique(oneTargetLosers$subject)
 ```
@@
 ```{r}
-#To do that, do I need to unmelt?
-ggplot(one, aes(x=
+#To do that, do I'd either need to unmelt (spread) but rather than do that, I programmed putting effLeft and effRight on same row into effDiff
+g<- ggplot(oneDiff, aes(x=effLeft, y=effRight, color=orientation)) + geom_point()
+g<-g +  geom_abline(intercept = 0, slope = 1) + theme_apa() + xlim(0,1) + ylim(0,1)
+g +    geom_rug(col=rgb(.5,0,0,alpha=.2))
+
 ```
@@
 
-Make master list of Ss to exclude also from one-target condition??
+I now have LRtestLosers, lowEffLosers, and oneTargetLosers
 
-Actually exclude the losers
+Actually exclude the losers? Probably better if flag them so can run inferential stats both with and without them.
+
+```{r exclude, echo=FALSE}
+allLosers <- unique( c(LRtestLosers, lowEffLosers, oneTargetLosers) )
+#afterExclusions<- dat %>% filter(  !(subject %in% lowEffs)   )
+dat <- dat %>% mutate(exclude = (subject %in% allLosers))
+#table(dat$subject,dat$exclude)
+```
+
+Next stop, inferential stats?  Save data with estimates and exclude field added.
+
 ```{r exclude, echo=FALSE}
 
-afterExclusions<- tw %>% filter(  !(subject %in% lowEffs)   )
-
+pathToSave <- file.path( "..","Data")
+saveRDS(dat, file.path(pathToSave,"backwards2E2_dataWithEstimatesAndExcludeColumn.rda"), compress = "bzip2")
 ```
-
-A<- tw %>% filter(subject=="AA")
-B<-A %>% ungroup %>% summarise(LRtestFails = sum(pLRtest<.05), effFails = sum(efficacy<.1))
-
-twoTarget %>% select(-targetStreamOrientations,-warnings,-letterSeq) %>% group_by(stream,orientation) %>% summarise_all(funs(if_else(is.numeric(.), mean(.), 3))
-
-two<- twoTarget %>% group_by(stream,orientation) %>% summarise_all(funs(if_else(!is.list(.), mean(.), last(.))))
-
-  summarise_all(funs(mean,sd))
-A <- two %>% filter(subject=="AA")
-
-curvesOneTarget<- twoTarget %>%
-  group_by(subject) %>% 
-  do(calc_curves_dataframe(.,minSPE,maxSPE,numItemsInStream))
-
-```
-
-Participant AI, BB suspicion of  attending to one side from twoTargets condition.

### 2018-08-29 E2/analysis/histsWithFitsOneTarget.Rmd
this is going well

--- 2018-08-23 E2/analysis/histsWithFitsOneTarget.Rmd
@@
 author: "Alex Holcombe"
-date: "2/26/2018"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output: html_document
@@
 
-## Exclusions
-
-Exclude by likelihood ratio test and efficacy.
-
-For each subject, calculate how many two-target streams they failed. If 3 or 4 of 4, exclude (didn't think to mention 1-target condition in prereg).
-
-Also check whether average efficacy is below 10% in both streams in either the upright or the reversed orientation.
-
-```{r exclusions}
-
-twoTarget<- dg %>% filter(targetSide=="Both")
-
-AA<- twoTarget %>% filter(subject=="AA")
-#Function for one subject, calculating how many stream conditions failed the test
-calcExclude<- function(oneSub) {
-  oneSub %>% group_by(stream,orientation)
-}
-#first, collapse over trials
-#https://stackoverflow.com/questions/51978519/summarise-numeric-columns-return-last-value-of-non-numeric
-tw<- twoTarget %>% group_by(subject,stream,orientation) %>% select(-targetStreamOrientations,-letterSeq) %>% #can't include list columns
-  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })
-
-#for each subject, count pLRtest instances >.05 and efficay <.1
-B<-tw %>% ungroup %>% group_by(subject) %>% summarise(LRtestFails = sum(pLRtest>.05), effFails = sum(efficacy<.1))                                                              
-#B %>% print(n=Inf)
-exclude <- B %>% filter(LRtestFails>=3, effFails >
-
-# efficacy is below 10% in both streams in either the upright or the reversed orientation.
-# To calculate that, need to group by subject and orientation
-C<-tw %>% ungroup %>% group_by(subject,orientation) %>% summarise(bothEffsFailed = sum(efficacy<.1)==2)
-#Find which subjects, if any, has a bothEffsFailed==TRUE
-lowEff<- C %>% filter(bothEffsFailed==TRUE)
-lowEff <- unique(lowEff$subject)
-
-A<- tw %>% filter(subject=="AA")
-B<-A %>% ungroup %>% summarise(LRtestFails = sum(pLRtest<.05), effFails = sum(efficacy<.1))
-
-twoTarget %>% select(-targetStreamOrientations,-warnings,-letterSeq) %>% group_by(stream,orientation) %>% summarise_all(funs(if_else(is.numeric(.), mean(.), 3))
-
-two<- twoTarget %>% group_by(stream,orientation) %>% summarise_all(funs(if_else(!is.list(.), mean(.), last(.))))
-
-  summarise_all(funs(mean,sd))
-A <- two %>% filter(subject=="AA")
-
-curvesOneTarget<- twoTarget %>%
-  group_by(subject) %>% 
-  do(calc_curves_dataframe(.,minSPE,maxSPE,numItemsInStream))
-
-```
-
 Participant AI, BB suspicion of  attending to one side from twoTargets condition.

### 2018-08-29 E2/analysis/importE2data.Rmd
this is going well

--- 2018-08-16 E2/analysis/importE2data.Rmd
@@
 
-E<- readInAllFiles(rawDataPath)
+Ee<- readInAllFiles(rawDataPath)
 #TT<- as_tibble(E).  Seems to have no effect because weird hybrid type
@@
 #Calculate the serial position error
-E$SPE<- E$respSP - E$targetSP
+Ee$SPE<- Ee$respSP - Ee$targetSP
 
 #give targetSide values that make  sense, like "dual", "left", and "right"
-E<- E %>% mutate(targetSide = case_when(targetSide==0 ~ "both", 
+Ee<- Ee %>% mutate(targetSide = case_when(targetSide==0 ~ "both", 
                                        targetSide==1 ~"left", targetSide==2 ~"right",
@@
 #Very time-consuming!
-E$orientMinus3<- -99; E$orientMinus2<- -99; E$orientMinus1<- -99; E$orient0<- -99; E$orientPlus1<- -99; E$orientPlus2<- -99;
+Ee$orientMinus3<- -99; Ee$orientMinus2<- -99; Ee$orientMinus1<- -99; Ee$orient0<- -99; Ee$orientPlus1<- -99; Ee$orientPlus2<- -99;
 
-for (i in 1:nrow(E)) {
+for (i in 1:nrow(Ee)) {
   #row<- E[i,]
@@
   #If targetSP == NaN, it's a one-target trial and wrong side
-  if (is.nan(E[i,]$targetSP)) {
-    E[i,]$orientMinus3 <- NaN
+  if (is.nan(Ee[i,]$targetSP)) {
+    Ee[i,]$orientMinus3 <- NaN
   } else {
-    E[i,]$orientMinus4 <- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -4    ]
-    E[i,]$orientMinus3 <- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -3    ]
-    E[i,]$orientMinus2 <- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -2    ]
-    E[i,]$orientMinus1 <- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -1    ]
-    E[i,]$orient0 <- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP    ]
-    E[i,]$orientPlus1 <- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP +1    ]
-    E[i,]$orientPlus2 <- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP +2    ]
+    Ee[i,]$orientMinus4 <- Ee[i,]$targetStreamOrientations[[1]][    Ee[i,]$targetSP -4    ]
+    Ee[i,]$orientMinus3 <- Ee[i,]$targetStreamOrientations[[1]][    Ee[i,]$targetSP -3    ]
+    Ee[i,]$orientMinus2 <- Ee[i,]$targetStreamOrientations[[1]][    Ee[i,]$targetSP -2    ]
+    Ee[i,]$orientMinus1 <- Ee[i,]$targetStreamOrientations[[1]][    Ee[i,]$targetSP -1    ]
+    Ee[i,]$orient0 <- Ee[i,]$targetStreamOrientations[[1]][    Ee[i,]$targetSP    ]
+    Ee[i,]$orientPlus1 <- Ee[i,]$targetStreamOrientations[[1]][    Ee[i,]$targetSP +1    ]
+    Ee[i,]$orientPlus2 <- Ee[i,]$targetStreamOrientations[[1]][    Ee[i,]$targetSP +2    ]
   }
@@
 # for (i in 1:nrow(E)) {
-#   row<- E[i,]
+#   row<- Ee[i,]
 #   orientsThis <- row$targetStreamOrientations[[1]]
@@
 #   if (is.nan(targetPos)) {
-#     E[i,]$orientMinus3<- NaN
+#     Ee[i,]$orientMinus3<- NaN
 #   }
-#   E[i,]$orientMinus3<- orientsThis[targetPos-3]
-#   #E[i,]$orientMinus3<- E[i,]$targetStreamOrientations[[1]][targetPos-3]
-#   E[i,]$orientMinus2<- orientsThis[targetPos-2]
-#   E[i,]$orientMinus1<- orientsThis[targetPos-1]
-#   E[i,]$orient0<- orientsThis[targetPos] #should be same as $targetOrient
-#   E[i,]$orientPlus1<- orientsThis[targetPos+1]
-#   E[i,]$orientPlus2<- orientsThis[targetPos+2]
+#   Ee[i,]$orientMinus3<- orientsThis[targetPos-3]
+#   #Ee[i,]$orientMinus3<- Ee[i,]$targetStreamOrientations[[1]][targetPos-3]
+#   Ee[i,]$orientMinus2<- orientsThis[targetPos-2]
+#   Ee[i,]$orientMinus1<- orientsThis[targetPos-1]
+#   Ee[i,]$orient0<- orientsThis[targetPos] #should be same as $targetOrient
+#   Ee[i,]$orientPlus1<- orientsThis[targetPos+1]
+#   Ee[i,]$orientPlus2<- orientsThis[targetPos+2]
 # }
+```
+
+Rename some columns
+```{r renameColumns}
+#Something bizarre happens when knitting this, the below command throws error. But no problem when running without knitting
+#Error in names(E)[names(E) == "target"] <- "stream" :   names() applied to a non-vector
+names(Ee)[names(Ee) == 'target'] <-'stream'
+names(Ee)[names(Ee) == 'condition'] <-'orientation'
+Ee$stream[ Ee$stream==1 ] <- "Left"
+Ee$stream[ Ee$stream==2 ] <- "Right"
+Ee$targetSide[ Ee$targetSide==0 ] <- "Both"
+Ee$targetSide[ Ee$targetSide==1 ] <- "Left"
+Ee$targetSide[ Ee$targetSide==2 ] <- "Right"
+
+#Give orientations some better names.
+Ee$orientation[ Ee$orientation==1 ] <- "Canonical"
+Ee$orientation[ Ee$orientation==2 ] <- "Inverted"
 ```
@@
 ```{r saveData}
-backwards2_E2 <- E
+backwards2_E2 <- Ee
 pathToSave <- file.path( "..","Data")
-saveRDS(backwards2_E2, file = "../Data/backwards2E1_rawDataFromMAT.rda", compress="bzip2")
+saveRDS(backwards2_E2, file = "../Data/backwards2E2_rawDataFromMAT.rda", compress="bzip2")
 ```

### 2018-08-29 E2/analysis/validateE2data.Rmd
this is going well

--- 2018-08-16 E2/analysis/validateE2data.Rmd
@@
 ---
-title: "importE2data"
+title: "validateE2data"
 author: "Alex Holcombe"
-date: "2/27/2018"
+date:  "`r format(Sys.time(), '%d %B, %Y')`"
 output: html_document
@@
 
+Optional validation (don't have to run this to go to next step), loads raw data on its own and validates a few things
+
 1/3 of trials have 1 target (numTargets==1)
 targetSide ==0  means targets on both sides
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
 
@@
 
+Sanity-check some data by graphing.
+
+Investigate whether pathological identicalness of inverted/canonical happens, that happened in E1 . Seems it doesn't
+
+```{r}
+dg<- E2 %>% filter(subject < "AC")
+
+library(ggplot2)
+g=ggplot(dg, aes(x=SPE)) + facet_grid(subject+orientation~stream) #,  scales="free_y")
+minSPE<- -17; maxSPE<- 17  #But it's 16 items in stream, so shouldn't it be -15, +15?
+g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
+g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
+show(g)
+```
+
 Save data

### 2018-08-29 E2/analysis/dev/Chapter1.Rmd
right

--- 2018-08-16 E2/analysis/validateE2data.Rmd
@@
 ---
-title: "importE2data"
+title: "Chapter1"
 author: "Alex Holcombe"
-date: "2/27/2018"
+date: "8/28/2018"
 output: html_document
@@
 
-```{r}
-readInAllFiles<- function(rawDataPath) {
+# Chapter 1
 
-  files <- dir(path=rawDataPath,pattern='.mat')  #find all data files in this directory
-  dfAll<-data.frame()
-  for (i in 1:length(files)) { #read in each file
-    fileThis<- file.path(rawDataPath,files[i])
-    rawDataLoad=tryCatch(
-      readMat(fileThis),
-      error=function(e) {
-        stop( paste0("ERROR reading the file ",fileThis," :",e) )
-      } )
-    apparentSubjectName <- strsplit(files[i],split="_")[[1]][1]
-    subjectName<- rawDataLoad$participantID[[1]]
-    if (apparentSubjectName != subjectName) {
-      stop( paste0("WARNING apparentSubjectName",apparentSubjectName," from filename does not match subjectName in data structure",subjectName) )
-    }
-    else { cat(paste0(' Subject ',apparentSubjectName,' read in')) }
-    rawDataLoad$file <- files[i]
-
-    dfThis<- turnMATintoMeltedDataframe(rawDataLoad)
-
-    tryCatch(
-      dfAll<-rbind(dfAll,dfThis), #if fail to bind new with old,
-      error=function(e) { #Give feedback about how the error happened
-        cat(paste0("Tried to merge but error:",e) )
-      } )
-  }
-  return(dfAll)
-}
-```
-
-```{r readRawData}
-
-E<- readInAllFiles(rawDataPath)
-
-#Calculate the serial position error
-E$SPE<- E$respSP - E$targetSP
-
-```
-
-Why are there NaNs for targetSP? Because that's when only one target (targetSide==1 or ==2)
+This is chapter 1.
 
 ```{r}
-table(E$subject, E$targetSP,useNA="always")
+1
 ```
-
-Show how targetSP works
-```{r}
-library(dplyr)
-En <- E
-En$letterSeq<- NULL #otherwise can't use dplyr
-CU <- En %>% filter(subject=="CU")
-table(CU$target,CU$targetSide, useNA="always")
-#targetSide ==0  means targets on both sides 
-table(CU$targetSide, CU$targetSP, useNA="always")
-```
-
-The below works now that turnMATintoMeltedDataframe takes care of the list elements
-
-```{r}
-#CUall <- E[ ,E$subject=="CU" ]
-CUall <- E %>% filter(subject=="CU")
-i<- E$subject == "CU"
-E[i,] #previously didn't work. You can't slice a dataframe that has a list (letterSeq) as a member.
-subset(E, subject=="CU")
-j<- En$subject == "CU"
-EnCU<- En[j,]
-```
-
-Load in one file and inspect it
-```{r}
-
-source(file.path("turnMATintoMeltedDataframe.R"))
-
-fileThis<- file.path(rawDataPath, "CU_18-05-10_1.mat")
-
-rawDat=tryCatch(
-  readMat(fileThis),
-  error=function(e) {
-    stop( paste0("ERROR reading the file ",fileThis," :",e) )
-  } )
-str(rawDat)
-CUmelted<- turnMATintoMeltedDataframe(rawDat)
-
-```
-Somehow, what would seem to be the second trial has NaN for lots of stuff
-
- $ allLetterOrder         : num [1:216, 1:2, 1:16] 13 6 9 1 10 16 11 1 12 8 ...
- $ allLetterOrientations  : num [1:216, 1:16] 1 0 1 0 0 0 1 0 1 0 ...
- $ allTargets             : num [1:216, 1:2] 5 NaN 6 8 7 7 6 6 6 8 ...
- $ allTargetOrientations  : num [1:216, 1:2] 0 NaN 1 1 0 1 1 1 0 0 ...
- $ allResponses           : num [1:216, 1:2] 9 NaN 1 3 5 11 5 3 8 11 ...
- $ allResponseOrientations: num [1:216, 1:2] 0 NaN 0 0 0 1 0 0 0 0 ...
- $ allResponseOrder       : num [1:216, 1] 2 2 1 2 2 1 2 1 1 1 ...
- $ allRTs                 : num [1:216, 1:2] 12.78 NaN 1.81 6.19 10.19 ...
-
-Inspect another one 
-```{r}
-fileThis<- file.path(rawDataPath, "BC_18-01-25_1.mat")
-
-rawDat=tryCatch(
-  readMat(fileThis),
-  error=function(e) {
-    stop( paste0("ERROR reading the file ",fileThis," :",e) )
-  } )
-str(rawDat)
-
-```
-
-Sanity visual check that see histogram.
-```{r sanity visual check}
-
-sanityVisualCheck<-TRUE
-if (sanityVisualCheck) {
-  library(ggplot2)
-  g=ggplot(E, aes(x=SPE))
-  #plot data
-  g<-g+geom_histogram(binwidth=1)
-  show(g)
-
-  h<-ggplot(E,aes(x=SPE)) + geom_histogram(binwidth=1)
-  show(h)
-}
-
-```
-
-Meaning of condition variable
-* 1 - upright
-* 2 - inverted
-
-Make condition names more sensible - stream (Left/Right) and orientation (Canonical/Inverted)
-```{r}
-E2<- E
-names(E2)[names(E2) == 'target'] <-'stream'
-names(E2)[names(E2) == 'condition'] <-'orientation'
-E2$stream[ E2$stream==1 ] <- "Left"
-E2$stream[ E2$stream==2 ] <- "Right"
-#mutate condition to Orientation
-E2$orientation[ E2$orientation==1 ] <- "Canonical"
-E2$orientation[ E2$orientation==2 ] <- "Inverted"
-
-```
-
-Save data
-```{r}
-#save dataframe into proper library data directory
-backwards2_E2<-E2
-saveRDS(backwards2_E2, file = "../Data/backwards2E2_rawDataFromMAT.rda", compress="bzip2")
-```
-
-Look at 1-target performance, understand that variable better because didn't exist in previous experiment I think.
-
-The first table printed below could easily fool you into thinking that stream (Left, Right) and targetSide (1,2) are not consistent.
-We think that's because it records something for both sides (stream variable) even when there is only 1 target (targetSide == 1 or 2 rather than 0, which means two targets).
-
-I Disaggregated allLetterOrientations - done by importE2data.Rmd which renames it targetStreamOrientations
-
-Also maybe give targetSide values that make more common sense, like "dual", "left", and "right"
-
-But the second table below shows that they make sense, as they are only inconsistent when allResponseOrientations==NaN.  But what the hell are those trials where  allResponseOrientations==NaN?
-```{r}
-library(dplyr)
-E2$letterSeq <- NULL
-one <- E2 %>% filter(numTargets==1)
-
-table(one$stream,one$targetSide,useNA="always")
-
-table(E2$stream,E2$targetSide,E2$numTargets, useNA="always")
-
-table(one$stream,one$targetSide,one$orientation,useNA="always")
-
-#table(one$stream,one$targetSide,one$allResponseOrientations,useNA="always")
-```

### 2018-08-29 E2/analysis/inferentialStats.Rmd
right

--- 2018-03-07 E1/doInferentialStats.Rmd
@@
 ```{r}
-
-estimates<- read.csv(file="Results/backwards2E1_paramEstimates.csv", header=TRUE, sep=",")
-estimates$X<- NULL #Not sure where X comes from but it's just a row number
+dataPath <- file.path( "..","Data")
+#Created by exclusions.Rmd
+data<- readRDS( file.path(dataPath, "backwards2E2_dataWithEstimatesAndExcludeColumn.rda") ) 
+data<- as_tibble(data)
 ```
 
-Exclude Ss who failed likelihood ratio test in all four cases.
-Create new column nPassed for in how many conditions they passed (out of 4).
+Why is there no targetSP==NA in data? That's the rows reporting stream on opposite side. Eventually, use those to evaluate swaps. I don't know what happened to it, because it's created in 
+exclusions.Rmd, which on line 67 creates the thing merged with estimates that is saved. 
+filter( !is.na(targetSP) )
+
+For inferential stats, don't care about individual trials, so extract parameter estimates from full tibble.
 
 ```{r}
-library(dplyr)
 
-estimates<- estimates %>% mutate(mixModelFitsBetter = pLRtest<.05)
+estimates<- data %>% select(-targetStreamOrientations,-letterSeq) %>% #can't include list columns
+  group_by(subject,targetSide,orientation,stream) %>% #in other words, average over trials 
+  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })
 
-condtnVariableNames <- c("subject","stream", "orientation")
+library(readr)
+resultsPath = "Results/"
+write_tsv(estimates, file.path(resultsPath,"backwards2E2_paramEstimatesWithExcludeColumn.tsv"))
 
-e<- estimates %>%  group_by(subject) %>%
-  summarise(nPassed = sum(mixModelFitsBetter)) 
-estimates<- merge(e,estimates)
-
-estimates<- estimates %>% mutate( exclude = (nPassed==0)  )
 ```
 
-Assess the size of the effect to inform Bayesian test of next experiment (probably use raw units)
+Plot data
 
@@
 require(ggplot2)
-gg<-ggplot(estimates, aes(x=stream, y=efficacy)) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
+gg<-ggplot(estimates %>% filter(targetSide=="Both"), 
+           aes(x=stream, y=efficacy, color=orientation)) + geom_point() + stat_summary(fun.y="mean",geom="point",size=8, alpha=.5) + theme_apa() +ggtitle('nobody excluded')
 show(gg)
 
+hh<-ggplot(estimates %>% filter(targetSide=="Both", exclude==FALSE), aes(x=stream, y=efficacy, color=orientation)) + 
+  geom_point(position = position_dodge(width = 0.3)) + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5, position = position_dodge(width = 0.3)) + 
+  theme_apa() +ggtitle('after exclusions') + 
+  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = 0.3))
+
+show(hh)
+
+```
+
+Graph the latency and sigma as well
+BG, BJ, and AP have sigma near 4, clear outliers.
+
+```{r, echo=FALSE, message=FALSE}
+require(ggplot2)
+
+hh<-ggplot(estimates %>% filter(targetSide=="Both", exclude==FALSE), aes(x=stream, y=latency, color=orientation)) + 
+  geom_point(position = position_dodge(width = 0.3)) +  stat_summary(fun.y="mean",geom="point",size=4, alpha=.5, position = position_dodge(width = 0.3)) + 
+  theme_apa() +ggtitle('after exclusions') + 
+  stat_summary(fun.data =mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = 0.3))
+show(hh)
+
+ii<-ggplot(estimates %>% filter(targetSide=="Both", exclude==FALSE), aes(x=stream, y=precision, color=orientation)) + 
+  geom_point(position = position_dodge(width = 0.3)) +  stat_summary(fun.y="mean",geom="point",size=4, alpha=.5, position = position_dodge(width = 0.3)) + 
+  theme_apa() +ggtitle('after exclusions') + 
+  stat_summary(fun.data =mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = 0.3))
+show(ii)
+
+```
+
+Convert from long to wide so that both left and right stream estimates are in the same row, while going from wide to long for efficacy,latency,precision so can use facet_grid with them.
+
+```{r, echo=FALSE, message=FALSE}
+
 library(data.table)
-
 #The below works but drops many variables, so think about merging later
-eff<- data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="efficacy")
+#I guess what comes after the "~" is what goes from long to wide
+eff<- data.table::dcast(estimates, orientation+subject+exclude+targetSide~stream,
+                        value.var="efficacy")
 eff$dv<-"efficacy"
-lat<-data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="latency") 
+lat<-data.table::dcast(estimates, orientation+subject+exclude+targetSide~stream,
+                       value.var="latency") 
 lat$dv<-"latency"
-pre<-data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="precision") 
+pre<-data.table::dcast(estimates, orientation+subject+exclude+targetSide~stream,
+                       value.var="precision") 
 pre$dv<-"precision"
@@
 # data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var=c("efficacy","latency"))
-head(eff)
+head(all)
 ```
 
-Calculate bias
+Calculate and plot bias. 
 ```{r , fig.height=20, fig.width=5}
 all$leftBias<- all$Left - all$Right
-hh<-ggplot(all, aes(x=orientation, y=leftBias)) + facet_grid(.~dv) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
-hh<-hh+stat_summary(fun.data="mean_cl_boot",geom="errorbar",width=.3,conf.int=.95,
-                  width=5,size=1) 
+hh<-ggplot(all %>% filter(targetSide=="Both"), 
+           aes(x=orientation, y=leftBias)) + facet_grid(dv~., scales="free") +
+    stat_summary(fun.y="mean",geom="point",size=4, alpha=.5) + geom_point() +
+  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.2) + theme_apa() +
+
 show(hh)

### 2018-08-30 E2/analysis/histsWithFitsTwoTarget.Rmd
this is going well
about to remove loss of one-target opposite stream

@@
 ---
-title: "inspect histograms"
+title: "Histograms with fits two targets"
 author: "Alex Holcombe"
-date: "2/26/2018"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output: html_document
@@
 
-```
-
-Fix the variable names if not done
-
-```{r, echo=FALSE, message=FALSE}
-#Below should now be done by import
-if ('target' %in% names(data)) {
-  names(data)[names(data) == 'target'] <-'stream'
-  names(data)[names(data) == 'condition'] <-'orientation'
-  data$stream[ data$stream==1 ] <- "Left"
-  data$stream[ data$stream==2 ] <- "Right"
-}
-if (is.numeric(data$targetSide)) {
-  data$targetSide[ data$targetSide==0 ] <- "Both"
-  data$targetSide[ data$targetSide==1 ] <- "Left"
-  data$targetSide[ data$targetSide==2 ] <- "Right"
-}
-if (is.numeric(data$orientation)) {
-  #mutate condition to Orientation
-  data$orientation[ data$orientation==1 ] <- "Canonical"
-  data$orientation[ data$orientation==2 ] <- "Inverted"
-}
 ```
@@
 
-Calc numObservations for each condition
-```{r }
-#Calc numObservations for each condition. This is needed  for scaling the fine-grained Gaussian
-#gaussianScaledforData needs to know.
-if (!("nPerCond" %in% colnames(dg))) {
-  dfGroups<-  dg  %>%  filter( !is.na(targetSP) ) %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
-  #add nPerCond back to parameter estimates
-  estimates<- merge(estimates,dfGroups)
-}
-```
-
 Calculate curves for parameters, to plot on histograms.
@@
 ```{r }
-# #Calc numObservations to each condition. This is needed only for scaling the fine-grained Gaussian
-# #Calc the number of observations for each condition, because gaussianScaledforData needs to know.
-# dfGroups<- dg %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
-# #add nPerCond back to parameter estimates
-# estimates<- merge(estimates,dfGroups)
+# The fine-grained Gaussian to potentially plot
 
@@
 dAC<- dplyr::filter(dg,subject=="AC")  
-bothSides<- dAC %>% filter(targetSide=="Both")
-bothSidesCurves <- curves %>% filter(subject=="AC") %>% filter(targetSide=="Both")
+bothSides<- dAC %>% filter(targetSide=="both")
+bothSidesCurves <- curves %>% filter(subject=="AC") %>% filter(targetSide=="both")
 
@@
 ```{r singleTarget}
-oneSide<- dg %>% filter(targetSide!="Both")
-oneSideCurves <- curves %>% filter(targetSide!="Both")
+oneSide<- dg %>% filter(targetSide!="both")
+oneSideCurves <- curves %>% filter(targetSide!="both")
 

### 2018-08-30 E2/analysis/importE2data.Rmd
about to remove loss of one-target opposite stream

@@
 ```{r set up functions}
-
 #Loads raw data from MATLAB files for second backwards paper second experiment
@@
 
+Rename some columns
+```{r renameColumns}
+#Something bizarre happens when knitting this, the below command throws error. But no problem when running without knitting
+#Error in names(E)[names(E) == "target"] <- "stream" :   names() applied to a non-vector
+names(Ee)[names(Ee) == 'target'] <-'stream'
+names(Ee)[names(Ee) == 'condition'] <-'orientation'
+Ee$stream[ Ee$stream==1 ] <- "Left"
+Ee$stream[ Ee$stream==2 ] <- "Right"
+Ee$targetSide[ Ee$targetSide==0 ] <- "Both"
+Ee$targetSide[ Ee$targetSide==1 ] <- "Left"
+Ee$targetSide[ Ee$targetSide==2 ] <- "Right"
+
+#Give orientations some better names.
+Ee$orientation[ Ee$orientation==1 ] <- "Canonical"
+Ee$orientation[ Ee$orientation==2 ] <- "Inverted"
+```
+
 Next is very TIME CONSUMING.
 
-Break out into target position -3, -2 etc.
-the targetStreamOrient variable that contains the orientaiton of each item.
+Break out into target position -4, -3, -2 etc.
+the targetStreamOrientation variable that contains the orientation of each item.
 This will allow seeing whether having several orientations in a row triggers the lateral bias.
@@
 #Very time-consuming!
-Ee$orientMinus3<- -99; Ee$orientMinus2<- -99; Ee$orientMinus1<- -99; Ee$orient0<- -99; Ee$orientPlus1<- -99; Ee$orientPlus2<- -99;
+Ee$orientMinus4<- -99; Ee$orientMinus3<- -99; Ee$orientMinus2<- -99; Ee$orientMinus1<- -99; Ee$orient0<- -99; Ee$orientPlus1<- -99; Ee$orientPlus2<- -99;
 
@@
 
-Rename some columns
-```{r renameColumns}
-#Something bizarre happens when knitting this, the below command throws error. But no problem when running without knitting
-#Error in names(E)[names(E) == "target"] <- "stream" :   names() applied to a non-vector
-names(Ee)[names(Ee) == 'target'] <-'stream'
-names(Ee)[names(Ee) == 'condition'] <-'orientation'
-Ee$stream[ Ee$stream==1 ] <- "Left"
-Ee$stream[ Ee$stream==2 ] <- "Right"
-Ee$targetSide[ Ee$targetSide==0 ] <- "Both"
-Ee$targetSide[ Ee$targetSide==1 ] <- "Left"
-Ee$targetSide[ Ee$targetSide==2 ] <- "Right"
+Sanity check the orientMinus4, orientMinus3, etc.
+What I don't understand is why the anti-diagonal (the frequency of orientations on two different frames being different rather than being the same orientation)
+Maybe it's related to half of those being the same stream and half the opposite stream?
 
-#Give orientations some better names.
-Ee$orientation[ Ee$orientation==1 ] <- "Canonical"
-Ee$orientation[ Ee$orientation==2 ] <- "Inverted"
+I've checked some individual streams and they don't have any obvious constraints, look random.
+
+Because allLetterOrientations didn't have second dimension, unlike allLetterOrder, I assumed only the target stream was saved (after all, both have same orientation) and I called it targetStreamOrientations.
+I have now done code review of the MATLAB file, DualRSVP_randomOrientationWithinStreams.m
+
+```{r }
+
+table(Ee$orientMinus2,Ee$orientPlus1)
+
+```
+
+Yes, nan targetSP are in there, which are when is one-target trial and reporting other stream
+```{r}
+Ee %>% filter(is.nan(targetSP) )
 ```

### 2018-08-30 E2/DualRSVP_randomOrientationWithinStreams.m

@@ -423,11 +423,18 @@ if ~skipInstructions
 
             letterStream = mod([randperm(nItems); randperm(nItems)],nLetters)+1;
             letterOrientaion =  repmat(0:1,1,nLetters/2); % Must be an even number of letters in letterArray
+            %AH Now shuffle that list of 0's and 1's, by creating randomly shuffled indices
             letterOrientaion = letterOrientaion(randperm(length(letterOrientaion)));
             
             %Check target has correct orientation
             if (thisCondition==1)&&(letterOrientaion(thisTarget(1,1))==1)
-                uprightLetters = find(letterOrientaion==0);
+            	%AH I presume the above being true means target has the wrong orientation
+                uprightLetters = find(letterOrientaion==0); %AH uprightLetters now is a list of indices of where the stream has the correct orientation 
+                %AH Now, shuffle the 
+                %AH What does the ",1" do in randperm
+                %AH Seems like its setting a random upright letter to upside down,
+                %AH that way can then set the actual target to rightside up
+                %AH randperm(n,k) returns a row vector containing k unique integers selected randomly from 1 to n inclusive.  In other words, picks one of the upright letter positions
                 letterOrientaion(uprightLetters(randperm(length(uprightLetters),1))) = 1;
                 letterOrientaion(thisTarget(1,1)) = 0;
             elseif (thisCondition==2)&&(letterOrientaion(thisTarget(1,1))==0)                

### 2018-08-30 E2/analysis/turnMATintoMeltedDataframe.R

@@ -89,7 +89,8 @@ turnMATintoMeltedDataframe<- function(fromMAT) {
   
   #The only other variable that is a list even after splitting by target is allLetterOrientations
   # allLetterOrientations  : num [1:216, 1:16] 1 0 1 0 0 0 1 0 1 0 ...
-  # Because doesn't have second dimension, unlike allLetterOrder, seems only the target stream was saved?
+  # Because doesn't have second dimension, unlike allLetterOrder, seems only the target stream was saved,
+  # which would make sense because both streams have the same orientation.
   # We will have to set it to NULL to avoid list, but want to preserve the orientations at
   # least for a few items around the target so can assess if takes time to build up.
   # Need to ask Chris to verify it's the target stream it's recording.

### 2018-08-30 E2/analysis/exclusions.Rmd
Better fixed

@@
 es$X <- NULL
-dat<- merge(dd,es)
+#oldNoNa<- merge(dd,es) #loses the is.na(targetSP) rows
+dat<- merge(dd,es, all = TRUE)
 dg<-dat %>% filter( !is.na(targetSP) ) #excluding 1-target trials rows reporting stream on opposite side. Eventually, use those to evaluate swaps.
+
+```
+
+Merge operation lost it  targetSP==NA in data? That's the rows reporting stream on opposite side. Eventually, use those to evaluate swaps. I don't know what happened to it, because it's created in 
+exclusions.Rmd, which on line 67 creates the thing merged with estimates that is saved. 
+```{r, echo=FALSE}
+table(dat$targetSP,useNA="ifany")
+
+#dat %>% filter( is.na(targetSP) )
 ```
@@
 
-twoTarget<- dg %>% filter(targetSide=="Both")
+twoTarget<- dg %>% filter(targetSide=="both")
 
@@
 #Calculate exclusions based on one-target performance
-oneTarget<- dg %>% filter(targetSide!="Both")
+oneTarget<- dg %>% filter(targetSide!="both")
 one<- oneTarget %>% group_by(subject,targetSide,orientation) %>% 
@@
   #STOP
-  effLeft<- (df %>% filter(targetSide=="Left"))$efficacy 
-  effRight<- (df %>% filter(targetSide=="Right"))$efficacy
+  effLeft<- (df %>% filter(targetSide=="left"))$efficacy 
+  effRight<- (df %>% filter(targetSide=="right"))$efficacy
   df$effDiff<- effLeft - effRight
@@
 
-  failedBothLR <- ((df %>% filter(targetSide=="Left"))$pLRtest >= .05) &
-    ((df %>% filter(targetSide=="Right"))$pLRtest >= .05)
+  failedBothLR <- ((df %>% filter(targetSide=="left"))$pLRtest >= .05) &
+    ((df %>% filter(targetSide=="right"))$pLRtest >= .05)
   df$failedBothLR <- failedBothLR
@@
 #To do that, do I'd either need to unmelt (spread) but rather than do that, I programmed putting effLeft and effRight on same row into effDiff
+library(ggplot2)
 g<- ggplot(oneDiff, aes(x=effLeft, y=effRight, color=orientation)) + geom_point()
-g<-g +  geom_abline(intercept = 0, slope = 1) + theme_apa() + xlim(0,1) + ylim(0,1)
+g<-g +  geom_abline(intercept = 0, slope = 1) + theme_bw() + xlim(0,1) + ylim(0,1)
 g +    geom_rug(col=rgb(.5,0,0,alpha=.2))
@@
 
-```{r exclude, echo=FALSE}
+```{r excludeAll, echo=FALSE}
 allLosers <- unique( c(LRtestLosers, lowEffLosers, oneTargetLosers) )
@@
 ```{r exclude, echo=FALSE}
-
 pathToSave <- file.path( "..","Data")

### 2018-08-30 E2/analysis/exclusions.Rmd
cool

@@
 
-Merge operation lost it  targetSP==NA in data? That's the rows reporting stream on opposite side. Eventually, use those to evaluate swaps. I don't know what happened to it, because it's created in 
-exclusions.Rmd, which on line 67 creates the thing merged with estimates that is saved. 
+Merge operation lost   targetSP==NA in data, but now fixed
 ```{r, echo=FALSE}

### 2018-08-30 E2/analysis/inferentialStats.Rmd
about to remove loss of one-target opposite stream
cool

@@
 
-Why is there no targetSP==NA in data? That's the rows reporting stream on opposite side. Eventually, use those to evaluate swaps. I don't know what happened to it, because it's created in 
-exclusions.Rmd, which on line 67 creates the thing merged with estimates that is saved. 
-filter( !is.na(targetSP) )
-
 For inferential stats, don't care about individual trials, so extract parameter estimates from full tibble.
@@
 require(ggplot2)
-gg<-ggplot(estimates %>% filter(targetSide=="Both"), 
+gg<-ggplot(estimates %>% filter(targetSide=="both"), 
            aes(x=stream, y=efficacy, color=orientation)) + geom_point() + stat_summary(fun.y="mean",geom="point",size=8, alpha=.5) + theme_apa() +ggtitle('nobody excluded')
@@
 
-hh<-ggplot(estimates %>% filter(targetSide=="Both", exclude==FALSE), aes(x=stream, y=efficacy, color=orientation)) + 
+hh<-ggplot(estimates %>% filter(targetSide=="both", exclude==FALSE), aes(x=stream, y=efficacy, color=orientation)) + 
   geom_point(position = position_dodge(width = 0.3)) + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5, position = position_dodge(width = 0.3)) + 
@@
 
-hh<-ggplot(estimates %>% filter(targetSide=="Both", exclude==FALSE), aes(x=stream, y=latency, color=orientation)) + 
+hh<-ggplot(estimates %>% filter(targetSide=="both", exclude==FALSE), aes(x=stream, y=latency, color=orientation)) + 
   geom_point(position = position_dodge(width = 0.3)) +  stat_summary(fun.y="mean",geom="point",size=4, alpha=.5, position = position_dodge(width = 0.3)) + 
@@
 
-ii<-ggplot(estimates %>% filter(targetSide=="Both", exclude==FALSE), aes(x=stream, y=precision, color=orientation)) + 
+ii<-ggplot(estimates %>% filter(targetSide=="both", exclude==FALSE), aes(x=stream, y=precision, color=orientation)) + 
   geom_point(position = position_dodge(width = 0.3)) +  stat_summary(fun.y="mean",geom="point",size=4, alpha=.5, position = position_dodge(width = 0.3)) + 
@@
 all$leftBias<- all$Left - all$Right
-hh<-ggplot(all %>% filter(targetSide=="Both"), 
+hh<-ggplot(all %>% filter(targetSide=="both"), 
            aes(x=orientation, y=leftBias)) + facet_grid(dv~., scales="free") +

### 2018-08-30 E2/analysis/analyzeStreakOfOrientations.Rmd
try

---
title: "orientationStreaks"
author: "Alex Holcombe"
date: "2/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in data, parameter estimates

```{r}
dataPath <- file.path( "..","Data")
#Created by exclusions.Rmd
ff<- readRDS( file.path(dataPath, "backwards2E2_dataWithEstimatesAndExcludeColumn.rda") ) 
ff<- as_tibble(ff)
```

Yes, nan targetSP are in there, which are when is one-target trial and reporting other stream
```{r}
table(ff$targetSP,useNA="ifany")
#ff %>% filter(is.nan(targetSP) )

```

Look at proportion correct for SPE -1,0,1 as function of number of orientations in a row

Calculate approx correct, then mean for each condition.
First do it for all trials, not for streaks
```{r}
ff <- ff %>% mutate(approxCorr = as.numeric( abs(SPE)<=1  ))

hh<- ff %>% select(-targetStreamOrientations,-letterSeq) %>% #can't include list columns
  filter(!is.nan(targetSP), numTargets>1) %>%
  group_by(subject,targetSide,orientation,stream) %>% #in other words, average over trials 
  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })
```

Plot data

```{r, echo=FALSE, message=FALSE}
require(ggplot2)
ppp<-ggplot(hh %>% filter(targetSide=="both", exclude==FALSE), 
           aes(x=stream, y=approxCorr, color=orientation)) + geom_point() +
    stat_summary(fun.y="mean",geom="point",size=8, alpha=.5) + 
    stat_summary(fun.y="mean",geom="line",aes(group=orientation)) + 
    stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.2) +
    theme_apa() +ggtitle('after exclusions')
show(ppp)

```

Now filter by streakiness

```{r, echo=FALSE}

ff<- ff %>% mutate( streak4 = (orientMinus3==orientMinus2) & (orientMinus2==orientMinus1) & (orientMinus1==orient0) )

jj<- ff %>% select(-targetStreamOrientations,-letterSeq) %>% #can't include list columns
  filter(!is.nan(targetSP), numTargets>1) %>%
  filter(streak4==TRUE) %>%
  group_by(subject,targetSide,orientation,stream) %>% #in other words, average over trials 
  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })

pp<-ggplot(jj %>% filter(targetSide=="both", exclude==FALSE), 
           aes(x=stream, y=approxCorr, color=orientation)) + geom_point() +
      stat_summary(fun.y="mean",geom="point",size=8, alpha=.5) + 
      stat_summary(fun.y="mean",geom="line",aes(group=orientation)) + 
      stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.2) +
      theme_apa() +ggtitle('after exclusions, and streak of 3 trials only')
show(pp)
```

Combine streak trials and non-streak trials in single plot

```{r, echo=FALSE}
jj<- ff %>% select(-targetStreamOrientations,-letterSeq) %>% #can't include list columns
  filter(!is.nan(targetSP), numTargets>1) %>%
  group_by(subject,streak4,targetSide,orientation,stream) %>% #in other words, average over trials 
  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })

qq<-ggplot(jj %>% filter(targetSide=="both", exclude==FALSE), 
           aes(x=stream, y=approxCorr, color=orientation)) + geom_point() +
      stat_summary(fun.y="mean",geom="point",size=8, alpha=.5) + 
      stat_summary(fun.y="mean",geom="line",aes(group=orientation)) + 
      stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.2) + facet_grid(.~streak4) +
      theme_apa() +ggtitle('after exclusions. Not streak of 4 (left) and streak of 4 (right) ')
show(qq)
```

No effect of the streak on the left bias. But it does look like it might have helped overall performance to have the streak, which would be cool, because it'd show the streak was registered in a sense by the brain.

### 2018-08-30 manuscript/manuscript.Rmd
good

@@
 ---
-title             : "The title"
+title             : "Backwards letters paper 2"
 shorttitle        : "Title"
@@
   - id            : "1"
-    institution   : "Wilhelm-Wundt-University"
+    institution   : "The University of Sydney"
   - id            : "2"
@@
 abstract: |
-  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
+  During natural reading of canonically-oriented English text, attention shifts from left to right.
 

### 2018-08-31 E1/inferentialStatsE1.Rmd
writing

--- 2018-03-07 E1/doInferentialStats.Rmd
@@
 
-estimates<- read.csv(file="Results/backwards2E1_paramEstimates.csv", header=TRUE, sep=",")
+resultsPath <- file.path("Results")
+estimates<- read.csv(file=file.path(resultsPath,"backwards2E1_paramEstimates.csv"), header=TRUE, sep=",")
 estimates$X<- NULL #Not sure where X comes from but it's just a row number
+library(tidyverse)
+estimates <- as_tibble(estimates)
 ```
+
+## Exclusions
+
+Exclusion based on low efficacy - "participants whose efficacy in either RSVP stream in any condition  lower than 10% (as indicated by the mixture modeling code)  removed from data analysis to reduce floor effects."
 
@@
 
+There was no mention of the LRtest in the prereg, forgot to mention that probably. But I looked at it and was fairly consistent with low efficacy.
+
 ```{r}
-library(dplyr)
 
-estimates<- estimates %>% mutate(mixModelFitsBetter = pLRtest<.05)
+#lowEff <- estimates %>% filter(efficacy <= .1)
 
-condtnVariableNames <- c("subject","stream", "orientation")
+lowE<- estimates %>%  group_by(subject) %>%
+  summarise(exclude = any(efficacy<.1) )
 
-e<- estimates %>%  group_by(subject) %>%
-  summarise(nPassed = sum(mixModelFitsBetter)) 
-estimates<- merge(e,estimates)
+#estimates<- estimates %>% mutate(mixModelFitsBetter = pLRtest<.05)
 
-estimates<- estimates %>% mutate( exclude = (nPassed==0)  )
+estimates<- merge(lowE,estimates) 
+estimates <- as_tibble(estimates)
+
+toExclude <- unique((estimates %>% filter(exclude==TRUE))$subject)
 ```
 
-Assess the size of the effect to inform Bayesian test of next experiment (probably use raw units)
+Num to exclude is `r length(toExclude)`
+```{r}
+write_tsv(estimates,path=file.path(resultsPath,'backwards2E1_paramEstimatesWithExcludeColumn.tsv'))
+
+```
+
+Assess the size of the effect
 

### 2018-08-31 E2/analysis/correlograms.Rmd
writing

---
title: "correlograms"
author: "Alex Holcombe"
date: "2/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Need to calculate the SPE of the orientation error. Unfortunately, it's ambiguous. Therefore, I will have multiple variables: orient0reported, orientPlus1reported, etc.   
And want to make a correlogram wherein

Read in data, parameter estimates

```{r}
dataPath <- file.path( "..","Data")
#Created by exclusions.Rmd
ff<- readRDS( file.path(dataPath, "backwards2E2_dataWithEstimatesAndExcludeColumn.rda") ) 
ff<- as_tibble(ff)
```

First look before collapsing wihtin subjects
```{r}
#I think I need to put the data in wide form so that it counts each combinatino of SPEleft and SPE right
#SPE<-data.table::dcast(ff %>% filter(exclude==FALSE), orientation+subject+targetSide~stream,
#                       value.var="SPE") 
SPE<-data.table::dcast(ff %>% filter(exclude==FALSE, numTargets==2), subject+trial~stream,
                       value.var="SPE") 
#Now get the counts so can plot heatmap
SPEcount<- SPE %>% count(Left,Right)
#plot that heatmap
correl<- ggplot(SPEcount, aes(x = Left, y = Right)) + geom_tile(aes(fill = n)) # +  facet_grid(separation~.)
correl

#Zero in on center because most large SPEs are guesses
#Plot with limits to zoom in
correl + xlim(-2,2) + ylim(-2,2)

```

What was it for the first paper? For E1,`r P1E1canonical`  Canonical and 
`r P1E1reversed` Mirror-reversed. Difference is `r P1E1canonical-P1E1reversed`
. For E2, bias difference very similar at `r P1E2leftBiasDiff`.
For E2 vertical arrangement, `r P1E2vertBiasDiff`

Probably want to calculate a correlation within each subject.

For correlogram

Could use below to plot average SPE for each condition, but should probably make it from scratch
```{r}
estimates <- read_tsv("backwards2E2_paramEstimatesWithExcludeColumn.tsv")
SPE<-data.table::dcast(estimates, orientation+subject+exclude+targetSide~stream,
                       value.var="SPE") 
```

The below is copied from code I wrote for Cheryl

#Calculate correlation
cor.test(df$responsePosRelativeleft, df$responsePosRelativeright,
         alternative = c("two.sided"),
         method = c("pearson"),
         exact = NULL, conf.level = 0.95, continuity = FALSE)

smallSPEs <- df %>% filter(abs(responsePosRelativeleft) < maxSPE ) %>% filter( abs(responsePosRelativeright) < maxSPE )

#ddply(smallSPEs, .(wordEcc), 
near<- smallSPEs %>% filter( wordEcc==1 )
far <- smallSPEs %>% filter( wordEcc > 1 )

cor.test(smallSPEs$responsePosRelativeleft, smallSPEs$responsePosRelativeright,
         alternative = c("two.sided"),
         method = c("pearson"),
         exact = NULL, conf.level = 0.95, continuity = FALSE)

nearCorr<- cor.test(near$responsePosRelativeleft, near$responsePosRelativeright,
                    alternative = c("two.sided"),
                    method = c("pearson"),
                    exact = NULL, conf.level = 0.95, continuity = FALSE)

farCorr<- cor.test(far$responsePosRelativeleft, far$responsePosRelativeright,
                   alternative = c("two.sided"),
                   method = c("pearson"),
                   exact = NULL, conf.level = 0.95, continuity = FALSE)

subjectCorr <- dlply(df,.(filename), function(x) cor.test(x$responsePosRelativeleft, x$responsePosRelativeright,
                                                          alternative = c("two.sided"),
                                                          method = c("pearson"),
                                                          exact = NULL, conf.level = 0.95, continuity = FALSE) )
farSubjectCorr <- dlply(far,.(filename), function(x) cor.test(x$responsePosRelativeleft, x$responsePosRelativeright,
                                                          alternative = c("two.sided"),
                                                          method = c("pearson"),
                                                          exact = NULL, conf.level = 0.95, continuity = FALSE) )
nearSubjectCorr <- dlply(near,.(filename), function(x) cor.test(x$responsePosRelativeleft, x$responsePosRelativeright,
                                                              alternative = c("two.sided"),
                                                              method = c("pearson"),
                                                              exact = NULL, conf.level = 0.95, continuity = FALSE) )

#nearCorrs <- c(.0711,.5278,.2991,.1807,.4680,.3630,-.0264,.3778,.4048,.1536,.1868,.0261,.1062,.1441,.0160)
#farCorrs <- c(-.2472,0,.2684,-.0428,.3736,-.0826,-.3119,-.0456,.1658,-.0603,-.2156,.0542,.1151,.1744,-.1492)
#Participants <- c(12,13,14,15,16,17,22,23,24,25,26,27,28,32,34)
#as.data.frame(allCorrData)
#allCorrData <- cbind(Participants,nearCorrs,farCorrs)
#t.test(nearCorrs,farCorrs, alternative = c("two.sided"), mu = 0, paired = FALSE, var.equal=FALSE, conf.level = 0.95)

#To-do
#Eliminate the five bad Ss [DONE!]
#Do the correlation analysis on each S individually, and then do paired t-tests on the subject means

#ADVANCED
#Address the argument that it wasn't correlated in the far just because of more guessing
#  - Do simulation of same correlation but with more guesses to make sure it is still significant despite the differing guessing rates
#Calculate for each subject and then do a t-test for near and far
#Analyse the frequent bigrams and actual words separately.

### 2018-08-31 E2/analysis/inferentialStats.Rmd
good
writing

@@
 data<- readRDS( file.path(dataPath, "backwards2E2_dataWithEstimatesAndExcludeColumn.rda") ) 
+library(tidyverse)
 data<- as_tibble(data)
@@
 
-Plot data
+Plot data 1 target
+
+```{r, echo=FALSE, message=FALSE}
+require(ggplot2)
+require(papaja)
+require(Hmisc) #for mean_cl_boot
+
+hh<-ggplot(estimates %>% filter(targetSide!="both", exclude==FALSE), aes(x=stream, y=efficacy, color=orientation)) + 
+  geom_point(position = position_dodge(width = 0.1)) + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5, position = position_dodge(width = 0.1)) + 
+  theme_apa() +ggtitle('after exclusions') + 
+  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = 0.1)) +
+  stat_summary(fun.y="mean",geom="line",aes(group=orientation))
+
+show(hh)
+
+```
+
+Plot data 2 targets
 
@@
 
-What was it for the first paper? For E1,`r P1E1canonical`  Canonical and 
-`r P1E1reversed` Mirror-reversed. Difference is `r P1E1canonical-P1E1reversed`
-. For E2, bias difference very similar at `r P1E2leftBiasDiff`.
-For E2 vertical arrangement, `r P1E2vertBiasDiff`
-
 Do conventional stats.

### 2018-08-31 manuscript/manuscript.Rmd
writing

@@
 author: 
-  - name          : "First Author"
+  - name          : "Alex O. Holcombe"
     affiliation   : "1"
     corresponding : yes    # Define only one corresponding author
-    address       : "Postal address"
+    address       : "School of Psychology, Brennan MacCallum Bldg (A19), The University of Sydney, Sydney, NSW 2006, Australia"
     email         : "my@email.com"
-  - name          : "Ernst-August Doelle"
+  - name          : "Christopher Bush"
     affiliation   : "1,2"
@@
 authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
+  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line. This manuscript was written in RMarkdown using the *papaja* package [@R-papaja].
 
@@
 abstract: |
-  During natural reading of canonically-oriented English text, attention shifts from left to right.
+  When one reads, early visual stages process the letters and words in parallel, but at a later stage, letters may need to be selected by attention to be fully processed.
+This transition to limited-capacity processing has recently been studied with a task requiring identification of simultaneously-presented letters embedded in a longer stream.
+Performance is markedly worse for identifying two letters rather than one.
+Although the task bears little resemblance to natural reading, implicit reading direction based on the letters' orientation nevertheless can determine which letter is poorly reported.
+This might be a result of word recognition processes rapidly and automatically extracting allocentric or object-relative reference frames.
+Alternatively, prioritisation based on implicit reading order (a left bias, for canonically-oriented letters) may not occur without a repeated regularity, such as the same orientation being used on multiple trials.
+In previous investigations, the trials were blocked, with all letters in the stream having the same orientation.
+Here, we first showed strong reading-order prioritization when trial order was random.
+Second, we found that when each frame within the stream had a random orientation, the bias was no longer affected by letter orientation, and defaulted to a left bias.
+Analysis of the errors supports the theory that the two streams of letters are sampled simultaneously, and a high-level process prioritizes one stream for consolidation into memory.
+  <!-- 250 word max for Memory & Cognition, https://www.springer.com/psychology/cognitive+psychology/journal/13421  -->
 
-  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
-
-  One sentence clearly stating the **general problem** being addressed by  this particular study.
-
-  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
-  One or two sentences to put the results into a more **general context**.
-
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "keywords"
+keywords          : "letters, words, attention, reading, allocentric"
 wordcount         : "X"
 
-bibliography      : ["r-references.bib"]
+bibliography      : ["referencesInfo/r-references.bib","referencesInfo/references.bib"]
 
@@
 
-# Methods
+# Introduction
+
+As you read these words, early visual stages process the letters and words in parallel.
+But at a later stage, letters may need to be selected by attention to be fully processed [@reichle_serial-attention_2011; @treisman_feature_1980].
+This transition to limited-capacity processing has recently been studied with a task requiring identification of simultaneously-presented letters embedded in a longer stream.
+Performance is markedly worse for identifying two letters rather than one.
+Although the task bears little resemblance to natural reading, implicit reading direction based on the letters' orientation nevertheless can determine which letter is poorly reported.
+This might be a result of word recognition processes rapidly and automatically extracting allocentric or object-relative reference frames.
+Alternatively, prioritisation based on implicit reading order (a left bias, for canonically-oriented letters) may not occur without a repeated regularity, such as the same orientation being used on multiple trials.
+In previous investigations, the trials were blocked, with all letters in the stream having the same orientation.
+Here, we first showed strong reading-order prioritization when trial order was random.
+Second, we found that when each frame within the stream had a random orientation, the bias was no longer affected by letter orientation, and defaulted to a left bias.
+Analysis of the errors supports the theory that the two streams of letters are sampled simultaneously, and a high-level process prioritizes one stream for consolidation into memory.
+
+The hypotheses, stopping rule, and other details of both experiments was preregistered (E1: https://osf.io/jdy8p/register/565fb3678c5e4a66b5582f67; E2: )
+
+# Experiment 1
+
+For the first experiment, 
+ https://osf.io/jdy8p/register/565fb3678c5e4a66b5582f67
+
+The experiment compares two within-subject levels of letter orientation (canonical or inverted), participants will experience both conditions within a single experimental session with the conditions presented in a mixed order.
+The method is almost exactly the same as that of Holcombe, Nguyen, & Goodbourn (in press).
+The only non-trivial difference is that in that experiment the two conditions were presented in separate blocks, whereas here they are mixed.
+This will allow us to determine whether simply seeing the backwards letters on an individual trial is sufficient to stop left-side prioritization, or whether having repeated successive trials of a particular condition is necessary.
+This speaks to how automatic the left-side prioritization observed in Holcombe et al.
+(in press) is.
+
+If the left-side prioritisation is caused by implied reading direction then the left side bias will be stronger for letters presented in canonical orientation compared to inverted letters.
+Based on the findings of Holcombe, Nguyen, & Goodbourn (in press), we predict that the left bias will be significantly stronger in canonical orientation than inverted.
+We predict that latency and precision will not differ significantly between letter orientations
+
+## Methods
+
 We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
@@
 
-## Participants
+### Participants
+
+Our original study (Holcombe, Nguyen, & Goodbourn, in press) of the effect of rotating the letters so they faced to the left yielded a large effect size of Cohen’s d = .94, which requires a sample size of only 14 to detect with 95% power.
+However, we expected that mixing the conditions within a block, as done here, might result in an effect size somewhat smaller, which led us to plan a sample size of 45 before exclusions (more details at the preregistration).
+
+Forty-five participants were run (XXX females).
 
@@
 
-In the prereg, we wrote: "
-Participants will be excluded if their average efficacy is below 10% in both streams in either the upright or the reversed orientation), or if the likelihood ratio test testing whether the data is fit better by the mixture model than by the guessing distribution fails (is not less than p<.05) for three or four of the four conditions.
-Because the Gaussian component of the mixture model is constrained to only allow latencies between -4 and 4, one effect of this is to eliminate participants who consistently do not report a letter near the time of the ring, instead for example reporting one of the last few items in the stream, which the data of past experiments indicates that rare participants do."
+As planned, six participants whose efficacy in either RSVP stream in any condition was lower than 10% (as indicated by the mixture modeling code) were removed from data analysis to reduce floor effects.
 
@@
 
+# Experiment 2
+
+## Method
+
+### Exclusions
+
+In the prereg, we wrote: "
+Participants will be excluded if their average efficacy is below 10% in both streams in either the upright or the reversed orientation), or if the likelihood ratio test testing whether the data is fit better by the mixture model than by the guessing distribution fails (is not less than p<.05) for three or four of the four conditions.
+Because the Gaussian component of the mixture model is constrained to only allow latencies between -4 and 4, one effect of this is to eliminate participants who consistently do not report a letter near the time of the ring, instead for example reporting one of the last few items in the stream, which the data of past experiments indicates that rare participants do."
+
 \newpage

### 2018-08-31 E1/inferentialStatsE1.Rmd
cool

@@
 show(gg)
+
+require(papaja)
+require(Hmisc) #for mean_cl_boot
+dodg = .3
+gg<-ggplot(estimates %>% filter(exclude==FALSE), 
+           aes(x=orientation, y=efficacy, color=stream)) + 
+          geom_point(position = position_dodge(width = dodg)) + 
+  stat_summary(fun.y="mean",geom="point",size=8, alpha=.5, position = position_dodge(width = dodg)) +
+  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = dodg)) +
+  theme_apa() +ggtitle('good')
+show(gg)
+
+```
 

### 2018-08-31 manuscript/manuscript.Rmd
cool

@@
 This will allow us to determine whether simply seeing the backwards letters on an individual trial is sufficient to stop left-side prioritization, or whether having repeated successive trials of a particular condition is necessary.
-This speaks to how automatic the left-side prioritization observed in Holcombe et al.
-(in press) is.
+This speaks to how automatic the left-side prioritization observed in @holcombe_implied_2017 is.
 
 If the left-side prioritisation is caused by implied reading direction then the left side bias will be stronger for letters presented in canonical orientation compared to inverted letters.
-Based on the findings of Holcombe, Nguyen, & Goodbourn (in press), we predict that the left bias will be significantly stronger in canonical orientation than inverted.
+Based on the findings of @holcombe_implied_2017, we predict that the left bias will be significantly stronger in canonical orientation than inverted.
 We predict that latency and precision will not differ significantly between letter orientations
@@
 
+Visual stimuli were generated using MATLAB R2012b with Psychtoolbox-3 (Brainard, 1997; Kleiner, Brainard, & Pelli, 2007; Pelli, 1997) on a MacBook Pro and displayed on a Mitsubishi DiamondView touchscreen LCD monitor with a resolution of 1,024 x 768 pixels and nominal refresh rate of 56 Hz (Experiment 2)CHRISCHECK.
+The code was adapted from the code used in a previous publication [@holcombe_implied_2017] that was written by [Patrick Goodbourn](https://orcid.org/0000-0001-7899-7355).
+
 ### Participants
 
-Our original study (Holcombe, Nguyen, & Goodbourn, in press) of the effect of rotating the letters so they faced to the left yielded a large effect size of Cohen’s d = .94, which requires a sample size of only 14 to detect with 95% power.
+Our previous study [@holcombe_implied_2017] of the effect of rotating the letters so they faced to the left yielded a large effect size of Cohen’s d = .94, which requires a sample size of only 14 to detect with 95% power.
 However, we expected that mixing the conditions within a block, as done here, might result in an effect size somewhat smaller, which led us to plan a sample size of 45 before exclusions (more details at the preregistration).
@@
 
-## Material
-
 ## Procedure
@@
 ## Data analysis
-We used `r cite_r("r-references.bib")` for all our analyses.
+The analysis code and results are all online (https://github.com/alexholcombe/backwardsLtrs2ndPaper). `r cite_r("referencesInfo/r-references.bib")` was used for the analyses.
 
 # Results
+
+The results of Experiment 1 were very similar to those of @holcombe_implied_2017.
+
+```{r, echo=FALSE, message=FALSE}
+require(ggplot2)
+require(papaja)
+require(Hmisc) #for mean_cl_boot
+dodg = .3
+gg<-ggplot(estimates %>% filter(exclude==FALSE), 
+           aes(x=orientation, y=efficacy, color=stream)) + 
+          geom_point(position = position_dodge(width = dodg)) + 
+  stat_summary(fun.y="mean",geom="point",size=8, alpha=.5, position = position_dodge(width = dodg)) +
+  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = dodg)) +
+  theme_apa() +ggtitle('good')
+show(gg)
+```
 

### 2018-08-31 manuscript/manuscript.Rmd
right

@@
 figurelist        : no
+figsintext        : yes
 tablelist         : no
@@
 ```{r, echo=FALSE, message=FALSE}
+require(readr); require(dplyr)
+estimates <- read_tsv(file.path("..","E1","Results",'backwards2E1_paramEstimatesWithExcludeColumn.tsv'))
 require(ggplot2)
@@
   stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = dodg)) +
-  theme_apa() +ggtitle('good')
+  theme_apa() 
 show(gg)
 ```
+
+apa_print() and apa_table() facilitate reporting results of your analyses. Take a look at the .Rmd of the example manuscript in the folder example and the resulting .pdf.
+
+Drop a supported analysis result, such as an htest- or lm-object, into apa_print() and receive a list of possible character strings that you can use to report the results of your analysis.
+
+my_lm <- lm(Sepal.Width ~ Sepal.Length + Petal.Width + Petal.Length, data = iris)
+apa_lm <- apa_print(my_lm)
 

### 2018-09-03 manuscript/manuscript.Rmd
right

@@
 
-## Procedure
+### Procedure
 
@@
 
+### Mixture modeling
+In a traditional analyses of performance in cued , the dependent variable would simply be accuracy at reporting the cued letter, for the left and for the right stream.
+However, that simple analysis can conceal differences in the dynamics of target selection from the streams [@vul_temporal_2008].
+To address this, we developed mixture modeling [@goodbourn_pseudoextinction:_2015], which accommodates the possibility that in different conditions or with different streams, the latency of selection or temporal variability of selection may vary together with or separately from the rate of success of selecting a letter.
+
+To perform mixture modeling, the data are coded as *serial position errors*.
+On each trial, for each stream, the participant chooses a letter from the choice array presented at the end of the trial.
+All the letters in the choice array were in fact presented, but they occurred in a random order.
+Thus, the letter presented can be expressed as a serial position error.
+The serial position error is 0 if the letter in the ring cue is reported.
+The serial position is -1 if the letter presented just before the cue is reported.
+It is +2 if the letter reported was that presented two after the cue And so on.
+
+A histogram of the temporal errors forms a prominent hill whose peak is around the time of the cue (it almost always peaks at the time of the cue or one letter after).
+In addition, there is a "floor" to the histogram which formed by both very large and small errors, which appears to conform to the shape predicted by the participant making a random guess or not perceiving the target letter.
+
+The mixture modeling code [@R-mixRSVP] fits the temporal error distribution with a combination of a Gaussian distribution (capturing the hill) and the guessing distribution.
+A maximum likelihood procedure finds the proportion of trials that can be attributed to the Gaussian, as opposed to the guessing distribution, and finds the best-fitting parameters ($\mu$, $\sigma$).
+Overall, then, we have the proportion of trials in which participants likely reported a letter around the time of the target, which we call *efficacy*, and the latency ($\mu$) and precision ($\sigma$).
+
 # Results
@@
 The results of Experiment 1 were very similar to those of @holcombe_implied_2017.
-
-```{r, echo=FALSE, message=FALSE}
+<!-- (ref:my-figure-caption)
+This is a figure caption for my figure.
+fig.cap = "(ref:my-figure-caption)"
+-->
+<!-- fig.cap style fig captions example here: https://osf.io/ujwvr/?show=revision-->
+```{r, fig.cap="Efficacy of ", echo=FALSE, message=FALSE}
 require(readr); require(dplyr)
@@
 ```
+
+```{r, ttestEffE1}
+if(!require(BayesFactor)){install.packages('BayesFactor')}
+library(BayesFactor)
+
+leftCanon <- estimates %>% filter(exclude==FALSE,orientation=="Canonical",stream=="Left")
+rightCanon <- estimates %>% filter(exclude==FALSE,orientation=="Canonical",stream=="Right")
+leftInverted<-estimates %>% filter(exclude==FALSE,orientation=="Inverted",stream=="Left")
+rightInverted<-estimates %>% filter(exclude==FALSE,orientation=="Inverted",stream=="Right")
+
+bb<- BayesFactor::ttestBF(leftCanon$efficacy,rightCanon$efficacy)
+
+```
+
+As seen in Figure 1, when the text was canonically oriented, efficacy for the left stream (`r round(mean(leftCanon$efficacy),2)` ± `r round(sd(leftCanon$efficacy),2)` was much higher than for the right stream (`r round(mean(rightCanon$efficacy),2)` ± `r round(sd(rightCanon$efficacy),2)`), but when the text was inverted, efficacy for the left stream (`r round(mean(leftInverted$efficacy),2)` ± `r round(sd(leftInverted$efficacy),2)` was lower than for the right stream (`r round(mean(rightInverted$efficacy),2)` ± `r round(sd(rightInverted$efficacy),2)`)
+
+efficacy was lower for the left stream () than for the right stream ().
+
+"A total of 3 paired samples t-test will be conducted on left bias score (calculations below), latency and precision.
+All 3 t-tests will enter letter orientation (canonical vs inverted) as the independent variable."
+
+Despite success at reporting the cued letter from the right stream occurring at a much lower rate
+
+PLOT LEFT BIAS, will plot only these from now on.
+
+```{r, fig.cap="Left bias", echo=FALSE, message=FALSE}
+
+gg<-ggplot(estimates %>% filter(exclude==FALSE), 
+           aes(x=orientation, y=efficacy, color=stream)) + 
+          geom_point(position = position_dodge(width = dodg)) + 
+  stat_summary(fun.y="mean",geom="point",size=8, alpha=.5, position = position_dodge(width = dodg)) +
+  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = dodg)) +
+  theme_apa() 
+show(gg)
+```
+
+apa_print()
+
+As declared in the preregistration subsequent analyses will use "left bias" as the dependent variable, which is efficacy (or other dependent variable, such as latency) for the left stream minus that for the right stream.
+
+A total of 3 paired samples t-test will be conducted on left bias score (calculations below), latency and precision.
+All 3 t-tests will enter letter orientation (canonical vs inverted) as the independent variable.
 

### 2018-09-04 manuscript/manuscript.Rmd
good

@@
 The experiment compares two within-subject levels of letter orientation (canonical or inverted), participants will experience both conditions within a single experimental session with the conditions presented in a mixed order.
-The method is almost exactly the same as that of Holcombe, Nguyen, & Goodbourn (in press).
+The method is almost exactly the same as that of [@holcombe_implied_2017].
 The only non-trivial difference is that in that experiment the two conditions were presented in separate blocks, whereas here they are mixed.
@@
 
-Preregistration: https://osf.io/r327j/
-http://htmlpreview.github.io/?https://github.com/alexholcombe/backwardsLtrs2ndPaper/blob/master/E2/analysis/bayesianAnalysisForPrereg.html , which used our custom prior (details in statistical models section) and found that 75 participants would yield a decent Bayes Factor (about 10 or more) for the effect size we are expecting.
+The letters in both streams were presented at a rate of 10 per second (66 ms presentation and 33 ms interstimulus interval).
+The two target letters, which were simultaneous, were surrounded by white cue rings of .1 deg thickness and 5 deg diameter.
+The screen was viewed from a distance of approximately 57 cm.
 
@@
 However, that simple analysis can conceal differences in the dynamics of target selection from the streams [@vul_temporal_2008].
-To address this, we developed mixture modeling [@goodbourn_pseudoextinction:_2015], which accommodates the possibility that in different conditions or with different streams, the latency of selection or temporal variability of selection may vary together with or separately from the rate of success of selecting a letter.
+To address this, we developed mixture modeling [@goodbourn_pseudoextinction2015], which accommodates the possibility that in different conditions or with different streams, the latency of selection or temporal variability of selection may vary together with or separately from the rate of success of selecting a letter.
 
@@
 
-bb<- BayesFactor::ttestBF(leftCanon$efficacy,rightCanon$efficacy)
-
+bb<- BayesFactor::ttestBF(leftCanon$efficacy-rightCanon$efficacy,
+                          leftInverted$efficacy-rightInverted$efficacy)
+latexReport<- (apa_print(bb)$statistic)[[1]]
 ```
 
-As seen in Figure 1, when the text was canonically oriented, efficacy for the left stream (`r round(mean(leftCanon$efficacy),2)` ± `r round(sd(leftCanon$efficacy),2)` was much higher than for the right stream (`r round(mean(rightCanon$efficacy),2)` ± `r round(sd(rightCanon$efficacy),2)`), but when the text was inverted, efficacy for the left stream (`r round(mean(leftInverted$efficacy),2)` ± `r round(sd(leftInverted$efficacy),2)` was lower than for the right stream (`r round(mean(rightInverted$efficacy),2)` ± `r round(sd(rightInverted$efficacy),2)`)
+As seen in Figure 1, when the text was canonically oriented, efficacy for the left stream (`r round(mean(leftCanon$efficacy),2)` ± `r round(sd(leftCanon$efficacy),2)`) was markedly higher than for the right stream (`r round(mean(rightCanon$efficacy),2)` ± `r round(sd(rightCanon$efficacy),2)`), but when the text was inverted, efficacy for the left stream (`r round(mean(leftInverted$efficacy),2)` ± `r round(sd(leftInverted$efficacy),2)`) was lower than for the right stream (`r round(mean(rightInverted$efficacy),2)` ± `r round(sd(rightInverted$efficacy),2)`).
 
-efficacy was lower for the left stream () than for the right stream ().
+For our primary statistical analyses (as preregistered), we use "left bias" as the dependent variable.
+Here, there is very strong evidence, `r latexReport`, that the left bias is greater in the canonical condition (`r mean(leftCanon$efficacy - rightCanon$efficacy)`  ± `r sd(leftCanon$efficacy - rightCanon$efficacy)`) than in the inverted condition (`r mean(leftInverted$efficacy - rightInverted$efficacy)`  ± `r sd(leftInverted$efficacy - rightInverted$efficacy)`).
+
+```{r latencyPrecision}
+bflLR<- ttestBF(leftCanon$latency,rightCanon$latency)
+latexReportBflLR<- (apa_print(bflLR)$statistic)[[1]]
+bflLRinvert<- ttestBF(leftInverted$latency,rightInverted$latency)
+latexReportBflLRinvert<- (apa_print(bflLRinvert)$statistic)[[1]]
+
+bfl<- ttestBF(leftCanon$latency - rightCanon$latency, leftInverted$latency - rightInverted$latency)
+latexReportBfl<- (apa_print(bfl)$statistic)[[1]]
+bfp<- ttestBF(leftCanon$precision - rightCanon$precision, leftInverted$precision - rightInverted$precision)
+latexReportBfp<- (apa_print(bfp)$statistic)[[1]]
+```
+
+The latency and precision were similar in all four conditions (canonical, left = `r round(mean(leftCanon$latency)*100)` ms, right = `r round(mean(rightCanon$latency)*100)` ms; inverted, left = `r round(mean(leftInverted$latency)*100)` ms, right = `r round(mean(rightInverted$latency)*100)` ms) such that the Bayes Factor t-test supported the null hypothesis of no difference between left and right in both the canonical (`r latexReportBflLR`) and inverted conditions (`r latexReportBflLRinvert`).
+
+canonical left stream: `r mean(leftCanon$latency)*100`
+`r mean(rightCanon$latency)*100`
+inverted left stream: `r mean(leftInverted$latency)*100`
+`r mean(rightInverted$latency)*100`
+
+`r latexReportBfl`
+
+leftCanon$latency
 
@@
 
-my_lm <- lm(Sepal.Width ~ Sepal.Length + Petal.Width + Petal.Length, data = iris)
-apa_lm <- apa_print(my_lm)
-
 # Discussion
@@
 ## Method
+
+Only differences from Experiment 1 are described below.
+
+### Procedure
+
+The letters in both streams were presented at a slightly slower rate than in Experiment 1, at 8 per second (80 ms presentation and 40 ms interstimulus interval).
+
+Preregistration: https://osf.io/r327j/
+http://htmlpreview.github.io/?https://github.com/alexholcombe/backwardsLtrs2ndPaper/blob/master/E2/analysis/bayesianAnalysisForPrereg.html , which used our custom prior (details in statistical models section) and found that 75 participants would yield a decent Bayes Factor (about 10 or more) for the effect size we are expecting.
 

### 2018-09-04 manuscript/manuscript.Rmd
mama

@@
 For our primary statistical analyses (as preregistered), we use "left bias" as the dependent variable.
-Here, there is very strong evidence, `r latexReport`, that the left bias is greater in the canonical condition (`r mean(leftCanon$efficacy - rightCanon$efficacy)`  ± `r sd(leftCanon$efficacy - rightCanon$efficacy)`) than in the inverted condition (`r mean(leftInverted$efficacy - rightInverted$efficacy)`  ± `r sd(leftInverted$efficacy - rightInverted$efficacy)`).
+The evidence here is very strong, `r latexReport`, that the left bias is greater in the canonical condition (`r mean(leftCanon$efficacy - rightCanon$efficacy)`  ± `r sd(leftCanon$efficacy - rightCanon$efficacy)`) than in the inverted condition (`r mean(leftInverted$efficacy - rightInverted$efficacy)`  ± `r sd(leftInverted$efficacy - rightInverted$efficacy)`).
 
@@
 latexReportBfp<- (apa_print(bfp)$statistic)[[1]]
+
 ```
 
-The latency and precision were similar in all four conditions (canonical, left = `r round(mean(leftCanon$latency)*100)` ms, right = `r round(mean(rightCanon$latency)*100)` ms; inverted, left = `r round(mean(leftInverted$latency)*100)` ms, right = `r round(mean(rightInverted$latency)*100)` ms) such that the Bayes Factor t-test supported the null hypothesis of no difference between left and right in both the canonical (`r latexReportBflLR`) and inverted conditions (`r latexReportBflLRinvert`).
-
-canonical left stream: `r mean(leftCanon$latency)*100`
-`r mean(rightCanon$latency)*100`
-inverted left stream: `r mean(leftInverted$latency)*100`
-`r mean(rightInverted$latency)*100`
-
-`r latexReportBfl`
-
-leftCanon$latency
-
-"A total of 3 paired samples t-test will be conducted on left bias score (calculations below), latency and precision.
-All 3 t-tests will enter letter orientation (canonical vs inverted) as the independent variable."
-
-Despite success at reporting the cued letter from the right stream occurring at a much lower rate
-
-PLOT LEFT BIAS, will plot only these from now on.
-
-```{r, fig.cap="Left bias", echo=FALSE, message=FALSE}
-
-gg<-ggplot(estimates %>% filter(exclude==FALSE), 
-           aes(x=orientation, y=efficacy, color=stream)) + 
-          geom_point(position = position_dodge(width = dodg)) + 
-  stat_summary(fun.y="mean",geom="point",size=8, alpha=.5, position = position_dodge(width = dodg)) +
-  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = dodg)) +
-  theme_apa() 
-show(gg)
-```
-
-apa_print()
-
-As declared in the preregistration subsequent analyses will use "left bias" as the dependent variable, which is efficacy (or other dependent variable, such as latency) for the left stream minus that for the right stream.
-
-A total of 3 paired samples t-test will be conducted on left bias score (calculations below), latency and precision.
-All 3 t-tests will enter letter orientation (canonical vs inverted) as the independent variable.
-
-apa_print() and apa_table() facilitate reporting results of your analyses. Take a look at the .Rmd of the example manuscript in the folder example and the resulting .pdf.
-
-Drop a supported analysis result, such as an htest- or lm-object, into apa_print() and receive a list of possible character strings that you can use to report the results of your analysis.
+The latency and precision were similar in all four conditions, such that the Bayesian t-test supported the null hypothesis of no difference between left and right in both conditions (canonical: `r latexReportBflLR`, inverted: `r latexReportBflLRinvert`).
+The latency values are:
+canonical, left = `r round(mean(leftCanon$latency)*100)` ms, right = `r round(mean(rightCanon$latency)*100)` ms; inverted, left = `r round(mean(leftInverted$latency)*100)` ms, right = `r round(mean(rightInverted$latency)*100)` ms, and the precision values are: canonical, left = `r round(mean(leftCanon$precision)*100)` ms, right = `r round(mean(rightCanon$precision)*100)`; inverted, left = `r round(mean(leftInverted$precision)*100)`, right = `r round(mean(rightInverted$precision)*100)`.
+Thus, there was no support for the hypothesis of a difference between the left biases for the canonical and the inverted conditions, for latency (`r latexReportBfl`) or for precision (`r latexReportBfp`.
 
 # Discussion
+
+Experiment 1 replicated the usual [@goodbourn_pseudoextinction2015; @holcombe_implied_2017; @ransley_reading_2018] dissociation between the efficacy and the dynamics of selection as indexed by latency and precision.
+That is, despite spatial biases in efficacy, there was no spatial bias in the latency and precision.
+This suggests that the reason for the poor performance on the right in the canonical condition, and on the left in the inverted condition, was not due to a delay in attentional selection or failure to allocate attention on one side.
+It may instead be due to a post-selection processing failure.
+Perhaps a post-selection process has a severe capacity limitation, resulting in strong competition between the two letters.
+The novelty here was that the particular letter orientation (canonical or inverted) was chosen randomly on each trial.
+In previous work, trials with the same orientation were blocked, and it was unknown whether this was necessary.
+It might have been the case that multiple successive trials with the same orientation are required before the reading direction bias was set up in the mind.
 

### 2018-09-05 manuscript/manuscript.Rmd
text

@@
 Analysis of the errors supports the theory that the two streams of letters are sampled simultaneously, and a high-level process prioritizes one stream for consolidation into memory.
+
+## Prioritization and coordinate frames
+
+What causes the letter on one side to be accurately identified more often than the letter on the other side?
+An answer to this should have two parts.
+One part should be an account of which processing stages the letter on one side gains an advantage, as well as the nature of those processing stages.
+The answer must also explain what causes the particular side that is prioritized to be the side that is prioritized.
+The present experiments were designed to provide insight into the second question.
+One possibility is that the prioritization emerges after repeated trials of a particular orientation. This is the "gradual emergence" possibility.
+
+could in principle reflect the slow emergence of a cognitive strategy, albeit one that does not
+
+the left side, say, to be prioritized rather than the right side.
+
+prioritization to occ
+
+A mental coordinate system must be used to allocate resources differentially to two locations, as a location can only be defined relative to a coordinate frame.
+To favor the first letter as defined by implicit reading direction, the mind must recognize the orientation of the letters and determine which of the two letters is on the left (for the canonical orientation) or which on the right (for the mirror-reversed or inverted condition).
+Studies of brain-injured patients have revealed aspects of how such relative spatial loca- tions are represented.
+Among patients with parietal neglect, some neglect the left visual field, some neglect the left half of each word, and some neglect the first part of each word, even if the letters are mirror- reversed so that the first part of the word is to the right (Haywood & Coltheart, 2001; Hillis et al., 2005).
+The reference frame thought to underpin the last phenomenon is sometimes referred to as object-centered or word-centered (Caramazza & Hillis, 1990).
+Conceivably, it could be critical to the implicit reading-direction bias documented in the present experiments: If the two letters acted like a single word, the word-centered frame would represent the location of the first letter that can then be prioritized for processing.
+Although some scheme for calculating the first letter is clearly needed, we deem it unlikely that it is the same as that used to represent a word.
+We give our reasons here.
+The present experiments used widely separated letters, which did not appear to form a single word or even a group.
+Moreover, the present finding of a very large advantage for the first letter relative to the last letter is not a typical characteristic of individual- word processing.
+Although the literature on individual-word rec- ognition does show that the first letters of a word are more informative than the last, the effect is not as large as the left advantage in our dual-RSVP paradigm (e.g., Blais et al., 2009).
+The RSVP literature does not seem to note a particularly large advantage either.
+In pilot experiments with a single RSVP stream of four-letter words (rather than two RSVP streams of individual letters), we have not observed a large advantage for the first two letters.
+Thus, it appears that the bias toward the first letter in our experiments does not reflect a word-centered processing frame and may be unrelated to the processes that yield word-centered neglect and normal word identification.
+Although two letters within a word may be processed in parallel (Ans, Carbonnel, & Valdois, 1998; Blais et al., 2009; Coltheart, Rastle, Perry, Langdon, & Ziegler, 2001; Davis, 2010; Perry, Ziegler, & Zorzi, 2007; cf.
+Whitney & Cornelissen, 2008), some evidence indicates that widely spaced letters are identified serially (Cohen, Dehaene, Vinckier, Jobert, & Montavont, 2008; Risko, Lanthier, & Besner, 2011).
+This serial processing may be a char- acteristic of the extraction of spatial relationships between arbi- trary objects as well as letters.
+In particular, attentional shifts between objects may be critical to extracting the spatial relation- ship between them (Franconeri, Scimeca, Roth, Helseth, & Kahn, 2012; Holcombe, Linares, & Vaziri-Pashkam, 2011).
+Our laresults suggest that the two target letters are sampled in parallel, but a subsequent stage may process them serially.
+Reading expe- rience may have led to this learned visual routine (Ullman, 1984) of beginning with the first letter according to letter orientation.
+Training monkeys on spatial-relationship judgments can result in parietal neurons tuned to relative location for the trained stimuli (Olson, 2003).
+Such neurons may be involved in coding the relative location of the letters, in combination with task-related (reading) top-down signals (Filimon, 2015).
+
+## The present studies
 
@@
 
+ANALYZE THE EFFECT OF RESPONSE ORDER
+
+ANALYZE THE EFFECT OF A RUN OF TRIALS, USING PERCENT CORRECT
+
 # Discussion
@@
 It might have been the case that multiple successive trials with the same orientation are required before the reading direction bias was set up in the mind.
+In Experiment 2 we further probed the necessary conditions for the reading direction bias.
 
 # Experiment 2
+
+LOOK UP IRINA'S AND CORBALLIS' PAPERS ON ORIENTAITON, SEE IF THEY SAY ANYTHING ABOUT REFERENCE FRAMES.
+One
+
+Kim in her thesis says "spatial bias on early versus late trials on each block) showed little effect,"
+COULD OPTIONALLY INCLUDE AN EXPERIMENT 3, JUST REPORTING THE PSYC1 CANONICAL VERSUS INVERTED TO SHOW IT DOESN'T NEED AN RSVP STREAM
 
@@
 
+## Results
+
 \newpage

### 2018-09-06 E1/trialStreaks.Rmd
whoops,first commit of trialStreaks

--- 2018-03-07 E1/doInferentialStats.Rmd
@@
 ---
-title: "InferentialStats"
+title: "trialStreaks"
 author: "Alex Holcombe"
-date: "2/27/2018"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output: html_document
@@
 
-Read in parameter estimates
+## Goal
 
-```{r}
+For each trial (row of the tibble), find the n-1, n-2, n-3 trial etc., put its orientation into a new column called neg1orient, neg2orient, etc.
+Then can filter by whether they equal the current orientation, and analyze left bias with percent correct the dependent variable.
 
-estimates<- read.csv(file="Results/backwards2E1_paramEstimates.csv", header=TRUE, sep=",")
-estimates$X<- NULL #Not sure where X comes from but it's just a row number
+## Implementation
+
+One could make a new grouping column that is trialnum / 4 that would group together every 4 trials.
+That would analyze only every 4 trials, which would avoid the problem of correlated trials, but would not provide a running streak assessment.
+
+rolling window calculations, autocorrelation
+from: http://www.business-science.io/timeseries-analysis/2017/08/30/tidy-timeseries-analysis-pt-4.html
+"apply tq_mutate() using the lag.xts function.
+We can provide column names for the new columns by prefixing “lag_” to the lag numbers, k, which the sequence from 1 to 28.
+The output is all of the lags for each package."
+
+## Statistical issues
+Did van der Burg or any of them address statistical independence? Alais' latest paper does not, so probably they never did.
+
+```{r load data, echo=FALSE, message=FALSE}
+library(tidyverse)
+
+#Import raw data
+dataPath<- file.path("Data/")
+#Experiment was administered by MATLAB
+#.mat file been preprocessed into melted long dataframe by importE1data.Rmd
+data<- readRDS( file.path(dataPath, "backwards2E1_rawDataFromMAT.rda") )
+
+#to work with dplyr, can't have array field like letterSeq
+data$letterSeq<- NULL
+
+#tidy data
+df<- as_tibble(data)
+
 ```
 
-Exclude Ss who failed likelihood ratio test in all four cases.
-Create new column nPassed for in how many conditions they passed (out of 4).
+I want only raw data, not estimates, but need to get exclude column from estimates to screen out excluded Ss
+```{r, echo=FALSE}
+library(readr)
+resultsPath <- file.path("Results")
+estimates<- readr::read_tsv(file=file.path(resultsPath,"backwards2E1_paramEstimatesWithExcludeColumn.tsv"))
+```
+Determine Ss to exclude and join to raw data
+```{r, echo=FALSE}
+#I don't remember what the excluded column is that's already there, but it excludes more Ss than we want
+excludeOrNot <- estimates %>% group_by(subject) %>% summarise(exclude = last(exclude))
 
-```{r}
-library(dplyr)
-
-estimates<- estimates %>% mutate(mixModelFitsBetter = pLRtest<.05)
-
-condtnVariableNames <- c("subject","stream", "orientation")
-
-e<- estimates %>%  group_by(subject) %>%
-  summarise(nPassed = sum(mixModelFitsBetter)) 
-estimates<- merge(e,estimates)
-
-estimates<- estimates %>% mutate( exclude = (nPassed==0)  )
+#Join with raw data
+df<- as_tibble( merge(excludeOrNot,df) )
+ #join
 ```
 
-Assess the size of the effect to inform Bayesian test of next experiment (probably use raw units)
+[This post](http://www.business-science.io/timeseries-analysis/2017/08/30/tidy-timeseries-analysis-pt-4.html) covers autocorrelation.
 
-```{r, echo=FALSE, message=FALSE}
-require(ggplot2)
-gg<-ggplot(estimates, aes(x=stream, y=efficacy)) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
-show(gg)
+Demonstrate I can calculate lags with tiny subset of data
+```{r}
+library(tidyquant)
+#tidyquant functions search for and key off a date or POSIXct column. 
+#So, need to convert trial to that.
+dfd <- df %>% mutate(trialDate = as_date(trial))
+#remove some crud so easier to view
+dfd<- dfd %>% select(-allRTs,-respSP,-resp,-targetSP,-excluded)
+#Critical column (orientation) must be numeric
+toyExn<- dfd %>% mutate(orientCI = ifelse(orientation=="Canonical",1,-1))
 
+k <- 1:4
+col_names <- paste0("lag_", k)
+
+lags <- toyExn %>% slice(1:10) %>%
+    tidyquant::tq_mutate(
+        select     = orientCI,
+        mutate_fun = lag.xts,
+        k          = 1:4,
+        col_rename = col_names
+    )
+print(lags %>% select(-exclude,-stream))
+```
+
+But need to do this for every condition. And want to do it after have brought together left stream and right stream correct, so can plot the autocorrelation for the left bias.
+
+# Use tq_mutate() to get lags 1:28 using lag.xts()
+k <- 1:4
+col_names <- paste0("lag_", k)
+
+toyEx <- dfd %>% select(trialDate,orientation)
+
+tidyverse_lags <- toyEx %>% slice(1:10) %>%
+    tidyquant::tq_mutate(
+        select     = orientation,
+        mutate_fun = lag.xts,
+        k          = 1:4,
+        col_rename = col_names
+    )
+
+toyExn<- toyEx %>% mutate(orientCI = ifelse(orientation=="Canonical",1,-1))
+
+lags <- toyExn %>% slice(1:10) %>%
+    tidyquant::tq_mutate(
+        select     = orientCI,
+        mutate_fun = lag.xts,
+        k          = 1:4,
+        col_rename = col_names
+    )
+#It looks like I need to do this separately for each subject
+#Try just one subject
+toyy <- toyExn %>% filter(subject==
+
+tidyverse_lags <- dfd %>%
+    tidyquant::tq_mutate(
+        select     = orientation,
+        mutate_fun = lag.xts,
+        k          = 1:28,
+        col_rename = col_names
+    )
+
+```
+
+Assess the size of the effect
+
+Get 
+```{r}
 library(data.table)
@@
 #The below works but drops many variables, so think about merging later
-eff<- data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="efficacy")
+eff<- data.table::dcast(estimates, orientation+subject~stream,value.var="efficacy")
 eff$dv<-"efficacy"
-lat<-data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="latency") 
+lat<-data.table::dcast(estimates, orientation+subject~stream,value.var="latency") 
 lat$dv<-"latency"
-pre<-data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="precision") 
+pre<-data.table::dcast(estimates, orientation+subject~stream,value.var="precision") 
 pre$dv<-"precision"

### 2018-09-06 E1/trialStreaksE1.Rmd
redo with all Ss

--- 2018-09-06 E1/trialStreaks.Rmd
@@
 But need to do this for every condition. And want to do it after have brought together left stream and right stream correct, so can plot the autocorrelation for the left bias.
+Look for where I did percent correct
 
-# Use tq_mutate() to get lags 1:28 using lag.xts()
+Try to bring left and right stream onto same row, so can calculate left bias on individual trials. 
+```{r}
+library(data.table)
+
+lr <- data.table::dcast(df, subject+trial+orientation~stream,value.var="SPE")
+as_tibble(lr)
+```
+
+```{r}
+lr<- lr %>% mutate(orientCI = ifelse(orientation=="Canonical",1,-1))
+
+#tidyquant functions search for and key off a date or POSIXct column. 
+#So, need to convert trial to that.
+lr <- lr %>% mutate(trialDate = as_date(trial))
+
 k <- 1:4
@@
 
-toyEx <- dfd %>% select(trialDate,orientation)
-
-tidyverse_lags <- toyEx %>% slice(1:10) %>%
-    tidyquant::tq_mutate(
-        select     = orientation,
-        mutate_fun = lag.xts,
-        k          = 1:4,
-        col_rename = col_names
-    )
-
-toyExn<- toyEx %>% mutate(orientCI = ifelse(orientation=="Canonical",1,-1))
-
-lags <- toyExn %>% slice(1:10) %>%
+lags <- lr %>% slice(1:10) %>%
     tidyquant::tq_mutate(
@@
     )
-#It looks like I need to do this separately for each subject
-#Try just one subject
-toyy <- toyExn %>% filter(subject==
-
-tidyverse_lags <- dfd %>%
-    tidyquant::tq_mutate(
-        select     = orientation,
-        mutate_fun = lag.xts,
-        k          = 1:28,
-        col_rename = col_names
-    )
+print(lags )
 
@@
 
-Assess the size of the effect
+Calculate left bias for each trial
+```{r}
+lags$leftBias <- lags$Left - lags$Right
 
-Get 
+```
+Calculate autocorrelation
+
+First, we need to correlate each of the lags to the “leftBias” column. This involves a few steps that can be strung together in a dplyr pipe (%>%):
+
+The goal is to get orientCI and each lag side-by-side so we can do a correlation.
+To do this we use gather() to pivot each of the lagged columns into a “tidy” (long format) data frame, and we exclude “package”, “date”, and “count” columns from the pivot.
+
 ```{r}
-library(data.table)
-
-#The below works but drops many variables, so think about merging later
-eff<- data.table::dcast(estimates, orientation+subject~stream,value.var="efficacy")
-eff$dv<-"efficacy"
-lat<-data.table::dcast(estimates, orientation+subject~stream,value.var="latency") 
-lat$dv<-"latency"
-pre<-data.table::dcast(estimates, orientation+subject~stream,value.var="precision") 
-pre$dv<-"precision"
-all<-rbind(eff,lat,pre)
-#I tried to do the above with a single command but makes an error even though StackOverflow example works
-#https://stackoverflow.com/questions/37332076/from-long-to-wide-data-with-multiple-columns/37332345
-# data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var=c("efficacy","latency"))
-head(eff)
+# Calculate the autocorrelations and 95% cutoffs
+lr_autocorr <- lags %>%
+    gather(key = "lag", value = "lag_value", -c(subject, trialDate,leftBias, orientation,trial)) %>%
+    mutate(lag = str_sub(lag, start = 5) %>% as.numeric) %>%
+    group_by(subject, lag) %>%
+    summarize(
+        cor = cor(x = leftBias, y = lag_value, use = "pairwise.complete.obs"),
+        cutoff_upper = 2/(n())^0.5,
+        cutoff_lower = -2/(n())^0.5
+        )
+lr_autocorr
 ```
 
-Calculate bias
+Now that we have the correlations calculated by package and lag number in a nice “tidy” format, we can visualize the autocorrelations
+
 ```{r , fig.height=20, fig.width=5}
-all$leftBias<- all$Left - all$Right
-hh<-ggplot(all, aes(x=orientation, y=leftBias)) + facet_grid(.~dv) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
-hh<-hh+stat_summary(fun.data="mean_cl_boot",geom="errorbar",width=.3,conf.int=.95,
-                  width=5,size=1) 
-show(hh)
+lr_autocorr %>%
+    ggplot(aes(x = lag, y = cor, color = subject, group = subject)) +
+    # Add horizontal line a y=0
+    geom_hline(yintercept = 0) +
+    # Plot autocorrelations
+    geom_point(size = 2) +
+    geom_segment(aes(xend = lag, yend = 0), size = 1) +
+    # Add cutoffs
+    geom_line(aes(y = cutoff_upper), color = "blue", linetype = 2) +
+    geom_line(aes(y = cutoff_lower), color = "blue", linetype = 2) +
+    # Add facets
+    facet_wrap(~ subject, ncol = 3) +
+    # Aesthetics
+    expand_limits(y = c(-1, 1)) +
+    scale_color_tq() +
+    theme_tq() +
+    labs(
+        title = paste0("Tidyverse ACF Plot: Lags ", rlang::expr_text(k)),
+        subtitle = "Appears to be a weekly pattern",
+        x = "Lags"
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)
+    )
 ```

### 2018-09-06 E1/trialStreaksE1.Rmd
wow i did it

@@
 lr <- lr %>% mutate(trialDate = as_date(trial))
-
+lr<- as_tibble(lr)
 k <- 1:4
@@
 
-lags <- lr %>% slice(1:10) %>%
+lags <- lr %>% group_by(subject) %>%
     tidyquant::tq_mutate(
@@
 ```
+
 Calculate autocorrelation
@@
 
-```{r , fig.height=20, fig.width=5}
+```{r , fig.height=5, fig.width=5}
 lr_autocorr %>%
@@
     # Add facets
-    facet_wrap(~ subject, ncol = 3) +
+    #facet_wrap(~ subject, ncol = 3) +
     # Aesthetics
-    expand_limits(y = c(-1, 1)) +
-    scale_color_tq() +
+    expand_limits(y = c(-.5, .5)) +
+    #scale_color_tq() +
+    stat_summary(fun.y="mean",geom="line") + #,aes(group=orientation))
     theme_tq() +
     labs(
-        title = paste0("Tidyverse ACF Plot: Lags ", rlang::expr_text(k)),
-        subtitle = "Appears to be a weekly pattern",
+        title = paste0("autocorrelation: Lags ", rlang::expr_text(k)),
+        subtitle = "Appears to have no pattern",
         x = "Lags"
@@
 
-Print effect size
+Collapse across subjects.
+
 ```{r}
-leftBiasesCanonical<- filter(all,dv=="efficacy" & orientation=="Canonical")$leftBias
-leftBiasCanonical <- mean( leftBiasesCanonical )
-leftBiasesInverted<- filter(all,dv=="efficacy" & orientation=="Inverted")$leftBias
-leftBiasInverted<- mean(   leftBiasesInverted  )
-leftBiasDiff<- leftBiasCanonical - leftBiasInverted
-leftBias_sd <- sd(   leftBiasesCanonical - leftBiasesInverted )
+lr_autocorr %>% group_by(lag) %>% summarise(cor=mean(cor))
 ```

### 2018-09-06 E1/trialStreaksE1.Rmd
good

@@
 ```{r}
+lr <- lr %>% filter(excludeOrNot==FALSE)
 lr<- lr %>% mutate(orientCI = ifelse(orientation=="Canonical",1,-1))
@@
     ) +
+    theme(
+        legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)
+    )
+```
+
+Use stat_summary to average across Ss in plot
+
+```{r , fig.height=5, fig.width=5}
+lr_autocorr %>% 
+    ggplot(aes(x = lag, y = cor)) +
+    # Add horizontal line a y=0
+    geom_hline(yintercept = 0) +
+    # Plot autocorrelations
+stat_summary(fun.y="mean",geom="point",size=4) + 
+  ggtitle('after exclusions') + 
+  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4) +
+    # Add facets
+    #facet_wrap(~ subject, ncol = 3) +
+    # Aesthetics
+    expand_limits(y = c(-.1, .1)) +
+    #scale_color_tq() +
+    stat_summary(fun.y="mean",geom="line") + #,aes(group=orientation))
+    theme_tq() +
+    xlab("Lag") + ylab('Correlation of leftBias with orientation of trial lag in past')
     theme(

### 2018-09-06 E1/trialStreaksE1.Rmd
right

--- 2018-09-06 E1/trialStreaksE1.Rmd
@@
 dfd <- df %>% mutate(trialDate = as_date(trial))
+dfd <- dfd %>% filter(exclude==FALSE)
+
 #remove some crud so easier to view
@@
 #Critical column (orientation) must be numeric
-toyExn<- dfd %>% mutate(orientCI = ifelse(orientation=="Canonical",1,-1))
+dfe<- dfd %>% mutate(orientCI = ifelse(orientation=="Canonical",1,-1))
 
@@
 
-lags <- toyExn %>% slice(1:10) %>%
+lags <- dfe %>% slice(1:10) %>%
     tidyquant::tq_mutate(
@@
 lr <- data.table::dcast(df, subject+trial+orientation~stream,value.var="SPE")
-as_tibble(lr)
+lr<-as_tibble(lr)
 ```
@@
 lr<- as_tibble(lr)
-k <- 1:4
+k <- 0:4
 col_names <- paste0("lag_", k)
@@
         mutate_fun = lag.xts,
-        k          = 1:4,
+        k          = k,
         col_rename = col_names
@@
 ```{r}
-lags$leftBias <- lags$Left - lags$Right
+lags<- lags %>% mutate(leftApproxCorr = abs(Left)<2, rightApproxCorr = abs(Right)<2 )
+lags$leftApproxCorr <- as.numeric(lags$leftApproxCorr)
+lags$rightApproxCorr <- as.numeric(lags$rightApproxCorr)
 
+lags$leftBias <- lags$leftApproxCorr - lags$rightApproxCorr
+```
+
+Correlation of leftBias with orientation on this trial is pretty small (`r cor(lags$orientCI,lags$leftBias)` )
+
+Calculate correlation with left performance instead of left bias
+```{r}
+cor(lags$orientCI, lags$leftApproxCorr)
 ```
@@
 lr_autocorr <- lags %>%
-    gather(key = "lag", value = "lag_value", -c(subject, trialDate,leftBias, orientation,trial)) %>%
+    gather(key = "lag", value = "lag_value", -c(subject, trialDate,leftBias, orientation,trial,Left,Right,leftApproxCorr,rightApproxCorr)) %>%
     mutate(lag = str_sub(lag, start = 5) %>% as.numeric) %>%
@@
     # Plot autocorrelations
-    geom_point(size = 2) +
-    geom_segment(aes(xend = lag, yend = 0), size = 1) +
-    # Add cutoffs
-    geom_line(aes(y = cutoff_upper), color = "blue", linetype = 2) +
-    geom_line(aes(y = cutoff_lower), color = "blue", linetype = 2) +
+    geom_point(size = .5) +
+      stat_summary(fun.y="mean",geom="line") + #,aes(group=orientation))
+   # geom_segment(aes(xend = lag, yend = 0), size = 1) +
+    # Plot summaries
+stat_summary(fun.y="mean",geom="point",size=4, aes(group="abc"),color="black") + 
+  stat_summary(fun.data = mean_cl_boot, geom="errorbar",  aes(group="abc"), width=0.4, color="black") +  
     # Add facets
@@
     #scale_color_tq() +
+    xlab("Lag") + ylab('Correlation of leftBias with orientation of previous trial')+
+    theme_tq() +
+    theme(
+        legend.position = "none",
+        axis.text.x = element_text(angle = 45, hjust = 1)
+    )
+```
+
+Use stat_summary to average across Ss in plot
+
+```{r }
+lr_autocorr %>% 
+    ggplot(aes(x = lag, y = cor)) +
+    # Add horizontal line a y=0
+    geom_hline(yintercept = 0) +
+    # Plot autocorrelations
+stat_summary(fun.y="mean",geom="point",size=4) + 
+  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4) +
+    # Add facets
+    #facet_wrap(~ subject, ncol = 3) +
+    # Aesthetics
+    expand_limits(y = c(-.1, .1)) +
+  ggtitle('Autocorrelation') + 
+    #scale_color_tq() +
     stat_summary(fun.y="mean",geom="line") + #,aes(group=orientation))
     theme_tq() +
-    labs(
-        title = paste0("autocorrelation: Lags ", rlang::expr_text(k)),
-        subtitle = "Appears to have no pattern",
-        x = "Lags"
-    ) +
+    xlab("Lag") + ylab('Correlation of leftBias with orientation of previous trial')
     theme(

### 2018-09-06 E1/trialStreaksE1.Rmd
right on

@@
 
-```{r , fig.height=5, fig.width=5}
+```{r , fig.height=5, fig.width=7}
+library('papaja')
 lr_autocorr %>%
@@
   stat_summary(fun.data = mean_cl_boot, geom="errorbar",  aes(group="abc"), width=0.4, color="black") +  
+      stat_summary(fun.y="mean",geom="line",aes(group="abc"),color="black",size=1) + 
     # Add facets
@@
     expand_limits(y = c(-.5, .5)) +
+   theme(legend.position = "none") +
     #scale_color_tq() +
-    xlab("Lag") + ylab('Correlation of leftBias with orientation of previous trial')+
-    theme_tq() +
-    theme(
-        legend.position = "none",
-        axis.text.x = element_text(angle = 45, hjust = 1)
-    )
+    xlab("Lag") + ylab('Correlation of leftBias with orientation')+
+  ggtitle('Correlation of left bias with this trial orientation\n and previous trial orientations') + theme_apa()
+
+ggsave(file.path("Results","Plots","autocorrE1.png"))
 ```
@@
 
-Collapse across subjects.
+Collapse across subjects so can potentially report exact values of correlations.
 
@@
 
-The leftBiasDiff for this experiment is `r leftBiasDiff` or in Cohen's d `leftBiasDiff / leftBias_sd`, very healthy.
+In case of nonlinearities, compare streaks of 3 in a row to everything else.
 
-```{r, echo=FALSE}
-E1n<- 16
-#Raw effect size plus standard error and sd
-P1E1canonical=.23;  P1E1c_se = .058; P1E1c_sd = P1E1c_se * sqrt(E1n)
-P1E1reversed= .062;  P1E1r_se = .07; P1E1r_sd = P1E1r_se * sqrt(E1n)
-#To calculate sd of the bias diff, consider that the variances sum,
-#therefore first sum the variances then convert back to standard deviation
-#Although only true for *independent* random variables and these unlikely to be totally independent
-P1E1diff_sdCalculated = sqrt( (P1E1c_sd^2 + P1E1r_sd^2) )  #.257
-
-E2n<- 24
-P1E2canonical=.218; P1E2c_se =.034; P1E2c_sd= P1E2c_se * sqrt(E2n)
-P1E2inverted=.017; P1E2i_se =.049; P1E2i_sd = P1E2i_se * sqrt(E2n)
-P1E2diff_sdCalculated = sqrt( (P1E2c_sd^2 + P1E2i_sd^2) ) #.2066
-
-P1E2vertUpright=.188; P1E2vu_se=.043; P1E2vu_sd = P1E2vu_se * sqrt(E2n)
-P1E2vertInverted= -.07; P1E2vi_se=.049; P1E2vi_sd = P1E2vi_se * sqrt(E2n)
-P1E2vdiff_sdCalculated = sqrt( (P1E2vi_sd^2 + P1E2vu_sd^2) )
-
-P1E1leftBiasDiff<- P1E1canonical-P1E1reversed
-P1E2leftBiasDiff<- P1E2canonical - P1E2inverted
-P1E2vertBiasDiff<- P1E2vertUpright - P1E2vertInverted
-#reported as  .258 +/- .047 in the paper
-
-#The diff standard errors are also available from the paper directly. But still need to 
-#convert to SD to get Cohen's d.
-
-P1E1diff_se<-.07; P1E1diff_sd<- P1E1diff_se*sqrt(E2n) 
-P1E2diff_se<-.038; P1E2diff_sd<- P1E2diff_se*sqrt(E2n) 
-P1E2vdiff_se<- .047; P1E2vdiff_sd<- P1E2vdiff_se*sqrt(E2n)
-
+```{r}
+lagss<- lags %>% mutate(streak5 = (lag_0==lag_1) & (lag_1==lag_2) & (lag_2==lag_3) & (lag_3==lag_4))
+lagss<- lagss %>% mutate(streak4 = (lag_0==lag_1) & (lag_1==lag_2) & (lag_2==lag_3))
+table(lagss$streak4)
+table(lagss$streak5)
 ```
 
-What was it for the first paper? For E1,`r P1E1canonical`  Canonical and 
-`r P1E1reversed` Mirror-reversed. Difference is `r P1E1canonical-P1E1reversed`
-. For E2, bias difference very similar at `r P1E2leftBiasDiff`.
-For E2 vertical arrangement, `r P1E2vertBiasDiff`
+Plot left bias for streak versus not, grouped by orientation.
 
-Do conventional stats.
-```{r}
-
-require(ez)
-
-aa <- ezANOVA(data=filter(all,dv=="efficacy"), dv=leftBias, within=orientation, wid=subject)
-cat("F=", aa$ANOVA$F, " p=", aa$ANOVA$p, "\n", sep="")
-
+First collapse trials to calculate average within subject.
+```{r, plotstreak}
+lagavg <- lagss %>% filter(!is.na(streak4)) %>% 
+                  group_by(subject,orientation,streak4) %>%
+                summarise(leftBias = mean(leftBias))
 ```
 
-Do Bayesian t-test. That is, take the likelihood of the left bias difference observed under the prior and divide it by the likelihood of the left bias difference observed under the null.
-
-Use the default Bayes test (but for E2 will use a custom prior)
-
-```{r, echo=FALSE}
-if(!require(BayesFactor)){install.packages('BayesFactor')}
-library(BayesFactor)
-
-leftBiasDifferences<- leftBiasesCanonical - leftBiasesInverted
-
-ttestResults<- ttestBF( leftBiasDifferences )
-
+```{r, plotstreak}
+ggplot(lagss, aes(x=orientation,y=leftBias,color=streak4)) +
+  stat_summary(fun.y="mean",geom="point") +
+  theme_apa()
 ```
-
-The Bayes factor from a one-sample default Bayes t-test for the leftBiasDifferences is `r ttestResults@bayesFactor$bf`
-
-The full results are 
-```r 
-print(ttestResults)
-```
-
-A custom prior would provide more power, as detailed in bayesianAnalysisForPrereg.Rmd in E2 folder
-
-Do the correlogram and correlation

### 2018-09-07 E1/trialStreaksE1.Rmd
good

@@
 
-rolling window calculations, autocorrelation
+autocorrelation
 from: http://www.business-science.io/timeseries-analysis/2017/08/30/tidy-timeseries-analysis-pt-4.html
@@
 ## Statistical issues
-Did van der Burg or any of them address statistical independence? Alais' latest paper does not, so probably they never did.
+Did van der Burg or any of them address statistical independence? Alais' latest paper does not, so probably they never did. John Cass says they do it all in Excel.
 
@@
 library('papaja')
+alph<-.5
 lr_autocorr %>%
@@
     # Plot autocorrelations
-    geom_point(size = .5) +
-      stat_summary(fun.y="mean",geom="line") + #,aes(group=orientation))
+    geom_point(size = .5, alpha=alph) +
+      stat_summary(fun.y="mean",geom="line",alpha=alph) + #,aes(group=orientation))
    # geom_segment(aes(xend = lag, yend = 0), size = 1) +
@@
     xlab("Lag") + ylab('Correlation of leftBias with orientation')+
-  ggtitle('Correlation of left bias with this trial orientation\n and previous trial orientations') + theme_apa()
+  #ggtitle('Correlation of left bias with this trial orientation\n and previous trial orientations') +
+  theme_apa()
 
@@
 
+Do the Bayesian t-test comparing the biases when it's a streak versus not.
 ```{r, plotstreak}
-ggplot(lagss, aes(x=orientation,y=leftBias,color=streak4)) +
-  stat_summary(fun.y="mean",geom="point") +
+require(BayesFactor)
+streak<- lagavg %>% filter(streak4==TRUE)
+notStreak<- lagavg %>% filter(streak4==FALSE)
+bb<- BayesFactor::ttestBF(streak$leftBias, notStreak$leftBias)
+latexReport<- (apa_print(bb)$statistic)[[1]]
+cat(latexReport, file=file.path("Results","streakBayesfactor.txt")) #to be read in by manuscript.Rmd
+```
+
+Then plot. Nothing happening.
+```{r, plotstreak}
+dodg = .3
+ggplot(lagavg, aes(x=orientation,y=leftBias,color=streak4)) +
+  #stat_summary(fun.y="mean",geom="point") +
+      geom_hline(yintercept = 0) +
+          geom_point(position = position_dodge(width = dodg)) + 
+  stat_summary(fun.y="mean",geom="point",size=8, alpha=.5, position = position_dodge(width = dodg)) +
+  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = dodg)) +
   theme_apa()

### 2018-09-07 manuscript/manuscript.Rmd
good

@@
 <!-- fig.cap style fig captions example here: https://osf.io/ujwvr/?show=revision-->
-```{r, fig.cap="Efficacy of ", echo=FALSE, message=FALSE}
+
+(ref:effE1) Efficacy of .
+
+```{r, fig.cap='(ref:effE1)', echo=FALSE, message=FALSE}
 require(readr); require(dplyr)
@@
 
-ANALYZE THE EFFECT OF A RUN OF TRIALS, USING PERCENT CORRECT
+## No effect of previous trials
+
+Many perceptual phenomena are subject to serial dependencies.
+To investigate if that is the case here, we assessed whether the left bias on a trial is affected only by the orientation of the letters on that trial, or is instead also affected by the orientation of the letters on previou trials.
+Mixture modeling could not be used to calculate a left bias for an individual trial, so instead of efficacy we used "approximately correct", which was also used for certain purposes in previous papers [e.g., @holcombe_implied_2017].
+"Approximately correct" is defined as a serial position error of -1, 0, or 1.
+For an individual trial, then, left bias was this approximately correct value for the left stream minus approximately correct for the right stream.
+For each participant, this trial-by-trial left bias was correlated with the orientation of the current and four previous trials. This is plotted in Figure \@ref(corrE1f) .
+
+(ref:corrE1) Correlation of left bias with this trial's orientation and with previous trial orientations.
+
+\@ref(corrE1f)
+
+```{r corrE1f, fig.cap='(ref:corrE1)'}
+require(knitr)
+knitr::include_graphics(file.path("..","E1","Results","Plots","autocorrE1.png"))
+```
+
+```{r readStreakBF, echo=FALSE, message=FALSE}
+ filePath<- file.path("..","E1","Results","streakBayesfactor.txt")
+ con = file(filePath, "r")
+ line = readLines(con, n = 1)
+```
+
+It is possible that the build-up of the bias is nonlinear in that the effect of a streak of trials with the same orientation is greater than the sum of the effect of the individual previous trials.
+We therefore compared the left bias after a streak of four trials with the same orientation to the left bias in all the trials that were not preceded by such a streak.
+However, there was no evidence for an effect, Bayesian t-test `r line`.
 

### 2018-09-08 manuscript/manuscript.Rmd
interactions as expected

@@
 
-ANALYZE THE EFFECT OF RESPONSE ORDER
+## No effect of response order
+One process that reading order can affect is memory retrieval. For readers of languages written left-to-right, readers 
+When people retrieve a list of items
+
+One way that reading order can exert its effect is by affecting the
+
+In studies of serial-order memory, readers of European lan- guages have been shown to associate sequences of items with positions from left to right.
+Items presented later in a list are associated with spatial positions further to the right, even when all stimuli are presented at the same position (Rinaldi et al., 2015; van Dijck, Abrahamse, Majerus, & Fias, 2013).
+These studies further suggest that when people retrieve a list of items in order, they scan a mental representation from left to right, which facilitates retrieval in the correct temporal order.
+So far, the effect of implied reading order has not been tested in this literature, but it has been suggested that it will have an effect (Abrahamse, van Dijck, Majerus, & Fias, 2014).
+
+In certain conditions, memory for serial order is spatialized by the human mind.
 
@@
  line = readLines(con, n = 1)
+ close(con)
 ```
